### 2022-08-03 3-reports/characterization_report.Rmd
markdown report

---
title: "Untitled"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r moisture}
tar_read(gg_moisture)
```

```{r pH}
tar_read(gg_pH)
```

```{r sp_cond}
tar_read(gg_sp_cond)
```

```{r tctnts}
tar_read(gg_tctnts)
```

```{r weoc}
tar_read(gg_weoc)
```

```{r din}
tar_read(gg_din)
```

---

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>

### 2022-08-03 2-code/3-tctnts.R
update data files

@@ -97,5 +97,5 @@ data_with_key %>%
 #
 
 # export ------------------------------------------------------------------
-tctnts_samples %>% write.csv("1-data/processed/total_CNS.csv", row.names = F, na = "")
+tctnts_samples %>% write.csv("1-data/processed/total_CNS_partial_2022-07-24.csv", row.names = F, na = "")
 

### 2022-10-25 2-code/a-functions_processing.R
loss on ignition

@@ -46,6 +46,17 @@ process_moisture = function(moisture_dat){
 }
 
 
+# loss on ignition
+process_loi = function(loi_data){
+  
+  #loi = 
+  loi_data %>%
+    mutate_at(vars(contains("wt_")), as.numeric) %>% 
+  mutate(percent_om = ((wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g))*100,
+         percent_om = round(percent_om, 2))  
+  
+}
+
 # pH
 process_pH = function(pH_data){
 #pH_samples = 

### 2022-10-25 2-code/b-functions_analysis.R

@@ -19,6 +19,23 @@ plot_moisture = function(moisture_processed, sample_key){
   
 }
 
+plot_loi = function(loi_processed, sample_key){
+  
+  loi = 
+    loi_processed %>% 
+    left_join(sample_key) %>% 
+    reorder_horizon() %>% reorder_transect()  
+  
+  loi %>% 
+    ggplot(aes(x = site, y = percent_om, color = transect, shape = horizon))+
+    geom_point(position = position_dodge(width = 0.4),
+               size = 2)+
+    facet_wrap(region ~ ., scales = "free_x")+
+    labs(y = "Organic Matter, %")+
+    scale_color_manual(values = pal_transect)
+  
+}
+
 plot_pH = function(pH_processed, sample_key){
   
   pH =

### 2022-10-25 _targets.R

@@ -36,6 +36,8 @@ list(
   # data files
   tar_target(moisture_data, import_gsheet("1nYzExPmpv01IYalqo2L3pmDsBDvhjdvHYMJo31N5_MA")),
   tar_target(moisture_processed, process_moisture(moisture_data)),
+  tar_target(loi_data, import_gsheet("1vvNKOtewhjR0PVYCLfvpRxY9QiOWzUvRV0WX5ou2qT4")),
+  tar_target(loi_processed, process_loi(loi_data)),
   tar_target(pH_data, import_gsheet("177ZR7O9JpsgVmFfLh72LfJ72oM-oTc6k_r1qE0XHL64")),
   tar_target(pH_processed, process_pH(pH_data)),
   tar_target(tctn_data, import_tctn_data(FILEPATH = "1-data/tctnts")),
@@ -49,6 +51,7 @@ list(
   
   # analysis - graphs
   tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),
+  tar_target(gg_loi, plot_loi(loi_processed, sample_key)),
   tar_target(gg_pH, plot_pH(pH_processed, sample_key)),
   tar_target(gg_sp_cond, plot_sp_cond(pH_processed, sample_key)),
   tar_target(gg_tctnts, plot_tctnts(tctnts_data_samples, sample_key)),

### 2022-10-25 2-code/a-functions_processing.R
ICP data

@@ -228,7 +228,7 @@ import_icp_data = function(FILEPATH){
   
 }
 
-process_icp = function(icp_data){
+process_icp = function(icp_data, moisture_processed, subsampling){
   
   icp_processed = 
     icp_data %>% 
@@ -242,25 +242,27 @@ process_icp = function(icp_data){
     arrange(analysis_ID) %>% 
     left_join(analysis_key %>% dplyr::select(analysis_ID, sample_label)) %>% 
     relocate(analysis_ID, sample_label) 
-    
-  icp = 
+  
+  icp_samples = 
     icp_processed %>% 
-    left_join(sample_key) %>% 
-    filter(!is.na(site))
-
-  icp_long = 
-    icp %>%
     pivot_longer(cols = c(Na:S), names_to = "species", values_to = "ppm") %>% 
-    mutate(ppm = as.numeric(ppm)) %>% 
-    reorder_transect() %>% 
-    reorder_horizon()
-    
-  
-  icp_long %>% 
-    ggplot(aes(x = site, y = ppm, color = transect, shape = horizon)) +
-    geom_point(position = position_dodge(width = 0.3))+
-    facet_wrap(species ~ region, scales = "free")
-  
+    left_join(moisture_processed) %>% 
+    left_join(subsampling %>% dplyr::select(sample_label, base_cations_g)) %>% 
+    rename(fm_g = base_cations_g) %>% 
+    mutate(ppm = as.numeric(ppm),
+           od_g = fm_g/((gwc_perc/100)+1),
+           soilwater_g = fm_g - od_g,
+           ug_g = ppm * ((25 + soilwater_g)/od_g),
+           ug_g = round(ug_g, 2)) %>% 
+    dplyr::select(sample_label, species, ppm, ug_g) %>% 
+    arrange(sample_label) %>% 
+    pivot_longer(-c(sample_label, species)) %>% 
+    mutate(name = paste0(species, "_", name)) %>% 
+    dplyr::select(-species) %>% 
+    filter(!grepl("blank", sample_label)) %>% 
+    pivot_wider()
+
+  icp_samples
 }
 
 

### 2022-10-25 2-code/b-functions_analysis.R

@@ -156,3 +156,56 @@ plot_din = function(din_processed, sample_key){
        gg_nh4 = gg_nh4)
   
 }
+
+plot_icp = function(icp_processed, sample_key){
+  
+  icp = 
+    icp_processed %>% 
+    dplyr::select(sample_label, ends_with("_ug_g")) %>% 
+    pivot_longer(-sample_label, names_to = "species", values_to = "ug_g") %>% 
+    left_join(sample_key) %>% 
+    filter(!is.na(site)) %>% 
+    reorder_transect() %>% 
+    reorder_horizon()
+
+#  icp_long_outliers = 
+#    icp_long %>% 
+#    group_by(region, site, transect, horizon, species) %>% 
+#    mutate(sd = sd(ppm),
+#           mean = mean(ppm),
+#           outlier = ppm > mean + (3*sd) | ppm < mean - (3*sd),
+#           outlier2 = case_when(outlier == TRUE ~ 1,
+#                                outlier == FALSE ~ 0))
+
+  icp_labels = 
+    icp %>% 
+    mutate(group = case_when(grepl("Ca|Mg|Na|K", species) ~ "1-base cations",
+                             grepl("Al|Fe|Mn", species) ~ "2-other cations",
+                             grepl("S", species) ~ "3-others")) %>% 
+    filter(!is.na(group))
+  
+  gg_base_cations = 
+    icp_labels %>% 
+    filter(grepl("base cations", group)) %>% 
+    ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(species ~ region, scales = "free")
+  
+  gg_other_cations = 
+    icp_labels %>% 
+    filter(grepl("other cations", group)) %>% 
+    ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(species ~ region, scales = "free")
+  
+  gg_others = 
+    icp_labels %>% 
+    filter(grepl("others", group)) %>% 
+    ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(species ~ region, scales = "free")
+
+  list(gg_base_cations = gg_base_cations,
+       gg_other_cations = gg_other_cations,
+       gg_others = gg_others)
+}

### 2022-10-25 _targets.R

@@ -48,6 +48,8 @@ list(
   #tar_target(din_data, readxl::read_xlsx("1-data/din/223013 Patel Final Report.xlsx", sheet = "NO3-N and NH4-N data")),
   tar_target(din_data, import_din_data("1-data/din")),
   tar_target(din_processed, process_din(din_data, analysis_key, moisture_processed, subsampling)),
+  tar_target(icp_data, import_icp_data(FILEPATH = "1-data/icp")),
+  tar_target(icp_processed, process_icp(icp_data, moisture_processed, subsampling)),
   
   # analysis - graphs
   tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),
@@ -57,6 +59,7 @@ list(
   tar_target(gg_tctnts, plot_tctnts(tctnts_data_samples, sample_key)),
   tar_target(gg_weoc, plot_weoc(weoc_processed, sample_key)),
   tar_target(gg_din, plot_din(din_processed, sample_key)),
+  tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
 
   # report  
   tar_render(report, path = "3-reports/characterization_report.Rmd")

### 2022-10-25 2-code/b-functions_analysis.R
color palette

@@ -189,21 +189,24 @@ plot_icp = function(icp_processed, sample_key){
     filter(grepl("base cations", group)) %>% 
     ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
     geom_point(position = position_dodge(width = 0.3))+
-    facet_grid(species ~ region, scales = "free")
+    facet_grid(species ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)
   
   gg_other_cations = 
     icp_labels %>% 
     filter(grepl("other cations", group)) %>% 
     ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
     geom_point(position = position_dodge(width = 0.3))+
-    facet_grid(species ~ region, scales = "free")
+    facet_grid(species ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)
   
   gg_others = 
     icp_labels %>% 
     filter(grepl("others", group)) %>% 
     ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
     geom_point(position = position_dodge(width = 0.3))+
-    facet_grid(species ~ region, scales = "free")
+    facet_grid(species ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)
 
   list(gg_base_cations = gg_base_cations,
        gg_other_cations = gg_other_cations,

### 2022-10-25 3-reports/characterization_report.Rmd
loss on ignition
ICP data
update

@@
 ---
-title: "Untitled"
+title: "Synpotic Soil Characterization"
 output: 
@@
 
+---
+
+**Gravimetric moisture**
+
 ```{r moisture}
@@
 ```
+
+**Loss on ignition (% organic matter)**
+
+```{r loi}
+tar_read(gg_loi)
+```
+
+**pH**
 
@@
 
+**specific conductance**
+
 ```{r sp_cond}
@@
 
+**total CNS**
 ```{r tctnts}
@@
 ```
+
+**Water extractable organic C**
 
@@
 
+**Extractable inorganic N**
+
+Inorganic N was extracted using 2M KCl
+
 ```{r din}
 tar_read(gg_din)
+```
+
+**ICP results: base cations**
+
+Extracted using 1M NH4Cl
+
+```{r icp-base cations, fig.height=12, fig.width=7}
+tar_read(gg_icp)$gg_base_cations
+```
+
+**ICP results: other cations**
+
+Extracted using 1M NH4Cl
+
+```{r icp-other cations, fig.height=9, fig.width=7}
+tar_read(gg_icp)$gg_other_cations
+```
+
+**ICP results: others**
+
+Extracted using 1M NH4Cl
+
+```{r icp-others, fig.height=4, fig.width=7}
+tar_read(gg_icp)$gg_others
 ```

### 2022-12-13 2-code/a-functions_processing.R
processing and analysis scripts for ferrozine

@@ -232,7 +232,7 @@ import_icp_data = function(FILEPATH){
   
 }
 
-process_icp = function(icp_data, moisture_processed, subsampling){
+process_icp = function(icp_data, analysis_key, moisture_processed, subsampling){
   
   icp_processed = 
     icp_data %>% 
@@ -270,8 +270,114 @@ process_icp = function(icp_data, moisture_processed, subsampling){
 }
 
 
+# Iron - ferrozine
+import_iron = function(FILEPATH){
+  
+  # import map
+  cb = read_sheet("1pzvGUvjK6qV8BVYSfoc2cid88dx3v7ZOD7TnYFgyWbc", sheet = "cb") %>% mutate_all(as.character) %>% mutate(region = "CB")
+  wle = read_sheet("1pzvGUvjK6qV8BVYSfoc2cid88dx3v7ZOD7TnYFgyWbc", sheet = "wle") %>% mutate_all(as.character) %>% mutate(region = "WLE")
+  ferrozine_map = bind_rows(cb, wle)
+  
+  # import data files (plate reader)
+  filePaths_ferrozine <- list.files(path = FILEPATH, pattern = "csv", full.names = TRUE, recursive = TRUE)
+  ferrozine_data <- do.call(bind_rows, lapply(filePaths_ferrozine, function(path) {
+    df <- read.csv(path, skip = 24, header = TRUE) %>% mutate_all(as.character) %>% janitor::clean_names()
+    df = df %>% mutate(source = basename(path))
+    df}))
+  
+  list(ferrozine_map = ferrozine_map,
+       ferrozine_data = ferrozine_data)
+  
+}
 
 
-FILEPATH = "1-data/icp"
-analysis_key = read.csv("1-data/analysis_key.csv")
-sample_key = read.csv("1-data/sample_key.csv")
+process_iron = function(ferrozine_map, ferrozine_data, moisture_processed, subsampling){
+  
+  # clean the map
+  map_processed = 
+    ferrozine_map %>% 
+    mutate(tray_number = parse_number(tray_number)) %>% 
+    filter(!is.na(sample_name) & !is.na(tray_number)) %>% 
+    rename(sample_label = sample_name)
+  
+  # clean the data
+  data_processed = 
+    ferrozine_data %>% 
+    mutate_all(na_if,"") %>% 
+    dplyr::select(-x) %>% 
+    fill(x_1) %>% 
+    filter(x_2 == "562") %>% 
[…]

### 2022-12-13 2-code/b-functions_analysis.R

@@ -213,3 +213,40 @@ plot_icp = function(icp_processed, sample_key){
        gg_other_cations = gg_other_cations,
        gg_others = gg_others)
 }
+
+plot_ferrozine = function(ferrozine_processed, sample_key){
+  
+  ferrozine = 
+    ferrozine_processed %>% 
+    mutate(Fe_2_3 = Fe2_ppm/Fe3_ppm) %>% 
+   # dplyr::select(sample_label, ends_with("_ug_g")) %>% 
+  #  pivot_longer(-sample_label, names_to = "species", values_to = "ug_g") %>% 
+    left_join(sample_key) %>% 
+    filter(!is.na(site)) %>% 
+    reorder_transect() %>% 
+    reorder_horizon()
+  
+  gg_fe_2_3 = 
+    ferrozine %>% 
+    filter(Fe_2_3 > 0) %>% 
+    ggplot(aes(x = site, y = Fe_2_3, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(. ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)+
+    labs(title = "Fe-II/III",
+         subtitle = "ferrozine method")
+  
+  gg_fe_total = 
+    ferrozine %>% 
+    #filter(Fe_2_3 > 0) %>% 
+    ggplot(aes(x = site, y = Fe_total_ug_g, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(. ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)+
+    labs(title = "total Fe",
+         subtitle = "ferrozine method")
+  
+  list(gg_fe_2_3 = gg_fe_2_3,
+       gg_fe_total = gg_fe_total)
+  
+}

### 2022-12-13 _targets.R

@@ -50,7 +50,10 @@ list(
   tar_target(din_data, import_din_data("1-data/din")),
   tar_target(din_processed, process_din(din_data, analysis_key, moisture_processed, subsampling)),
   tar_target(icp_data, import_icp_data(FILEPATH = "1-data/icp")),
-  tar_target(icp_processed, process_icp(icp_data, moisture_processed, subsampling)),
+  tar_target(icp_processed, process_icp(icp_data, analysis_key, moisture_processed, subsampling)),
+  tar_target(ferrozine_data, import_iron(FILEPATH = "1-data/iron-ferrozine")$ferrozine_data),
+  tar_target(ferrozine_map, import_iron(FILEPATH = "1-data/iron-ferrozine")$ferrozine_map),
+  tar_target(ferrozine_processed, process_iron(ferrozine_map, ferrozine_data, moisture_processed, subsampling)),
   
   # analysis - graphs
   tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),
@@ -61,6 +64,7 @@ list(
   tar_target(gg_weoc, plot_weoc(weoc_processed, sample_key)),
   tar_target(gg_din, plot_din(din_processed, sample_key)),
   tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
+  tar_target(gg_ferrozine, plot_ferrozine(ferrozine_processed, sample_key)),
 
   # report  
   tar_render(report, path = "3-reports/characterization_report.Rmd")

### 2022-12-20 2-code/b-functions_analysis.R
mehlich-extractable P

@@ -250,3 +250,212 @@ plot_ferrozine = function(ferrozine_processed, sample_key){
        gg_fe_total = gg_fe_total)
   
 }
+
+plot_mehlich = function(mehlich_processed, sample_key){
+  
+  mehlich = 
+    mehlich_processed %>% 
+    left_join(sample_key) %>% 
+    filter(!is.na(site)) %>% 
+    reorder_transect() %>% 
+    reorder_horizon()
+  
+  #gg_mehlich = 
+    mehlich %>% 
+    ggplot(aes(x = site, y = mehlichp_ugg, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_grid(. ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)+
+    labs(title = "Mehlich-3 extractable P")
+  
+}
+
+
+
+
+
+
+
+
+combine_data = function(){
+  
+  df_list = list(moisture_processed, pH_processed, tctnts_data_samples, weoc_processed, din_processed, icp_processed, ferrozine_processed)
+  all_combined = 
+    df_list %>% reduce(full_join) %>% 
+    dplyr::select(-notes, -ends_with(c("_ppm", "_mgL", "_flag")))
+  
+  all_combined
+ #   left_join(sample_key) %>% 
+ #   filter(!grepl("_vac|rep", sample_label)) %>% 
+ #   dplyr::select(-S_ug_g) %>% 
+ #   #rowwise() %>% 
+ #   #mutate(SUM = rowSums(across(where(is.numeric))))
+ #   #replace(.,is.na(.),0) %>% 
+ #   force()
+}
+
+#all_data_combined = combine_data()
+
+compute_analyte_matrix = function(all_data_combined, sample_key){
+  
+  analyte_completion_matrix = 
+    all_data_combined %>% 
+    distinct()
+  #  group_by(sample_label) %>% 
+  #  dplyr::summarise(n = n())
+  #  pivot_longer(-sample_label) %>% 
+    left_join(sample_key) %>% 
+    group_by(name, region, site, transect, horizon) %>% 
+    dplyr::summarise(n = n()) %>% 
+    pivot_wider(values_from = "n")
+  
+  
+}
+
+#analyte_completion_matrix = compute_analyte_matrix(all_data_combined, sample_key)
+
+
+
+
+compute_pca = function(){
+  library(ggbiplot)
+  
+
+  
+  
[…]

### 2022-12-20 _targets.R

@@ -76,6 +76,8 @@ list(
   tar_target(gg_din, plot_din(din_processed, sample_key)),
   tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
   tar_target(gg_ferrozine, plot_ferrozine(ferrozine_processed, sample_key)),
+  tar_target(gg_mehlich, plot_mehlich(mehlich_processed, sample_key)),
+  
 
   # report  
   tar_render(report, path = "3-reports/characterization_report.Rmd")

### 2022-12-20 2-code/c-fticrrr-functions.R
icr: processing, van krevelens

@@ -13,6 +13,19 @@
 ################################################## #
 ################################################## #
 
+import_fticr_data = function(FILEPATH){
+  
+  df <- 
+    list.files(path=FILEPATH, full.names = TRUE) %>% 
+    lapply(read_csv) %>% 
+    bind_rows
+  
+}
+
+
+
+
+
 # 1. PROCESSING FUNCTIONS -------------------------------------------------
 ## LEVEL I FUNCTIONS -------------------------------------------------------
 ## for metadata file
@@ -221,3 +234,93 @@ make_fticr_data = function(icr_report, analysis_key, sample_key){
        data_long_blank_corrected = data_long_blank_corrected)
   
 }
+
+#
+# 2.  ANALYSIS FUNCTIONS --------------------------------------------------
+
+## Van Krevelen -----------------------------------------------------------
+
+gg_vankrev <- function(data,mapping){
+  ggplot(data,mapping) +
+    # plot points
+    geom_point(size=0.5, alpha = 0.5) + # set size and transparency
+    # axis labels
+    ylab("H/C") +
+    xlab("O/C") +
+    # axis limits
+    xlim(0,1.25) +
+    ylim(0,2.5) +
+    # add boundary lines for Van Krevelen regions
+    geom_segment(x = 0.0, y = 1.5, xend = 1.2, yend = 1.5,color="black",linetype="longdash") +
+    geom_segment(x = 0.0, y = 0.7, xend = 1.2, yend = 0.4,color="black",linetype="longdash") +
+    geom_segment(x = 0.0, y = 1.06, xend = 1.2, yend = 0.51,color="black",linetype="longdash") +
+    guides(colour = guide_legend(override.aes = list(alpha=1, size = 1)))
+}
+
+plot_vankrevelen = function(icr_data_trt, icr_meta){
+  
+  vk_domains = 
+    icr_meta %>% 
+    dplyr::select(formula, HC, OC, Class_detailed, Class) %>% 
+    gg_vankrev(aes(x = OC, y = HC, color = Class))
+  
+  
+  data_hcoc = 
+    icr_data_trt %>% 
+    left_join(icr_meta %>% dplyr::select(formula, HC, OC))
+  
+  vk_wle = 
+    data_hcoc %>% 
[…]

### 2022-12-20 _targets.R

@@ -60,10 +60,10 @@ list(
   tar_target(mehlich_map, import_mehlich(FILEPATH = "1-data/phosphorus-mehlich")$mehlich_map),
   tar_target(mehlich_data, import_mehlich(FILEPATH = "1-data/phosphorus-mehlich")$mehlich_data),
   tar_target(mehlich_processed, process_mehlich(mehlich_map, mehlich_data, moisture_processed, subsampling)),
-  tar_target(icr_report= import_fticr_data(FILEPATH = "1-data/icr")),
-  tar_target(icr_meta= make_fticr_meta(icr_report)$meta2),
-  tar_target(icr_data_long= make_fticr_data(icr_report, analysis_key, sample_key)$data_long_blank_corrected),
-  tar_target(icr_data_trt= make_fticr_data(icr_report, analysis_key, sample_key)$data_long_trt),
+  tar_target(icr_report, import_fticr_data(FILEPATH = "1-data/icr")),
+  tar_target(icr_meta, make_fticr_meta(icr_report)$meta2),
+  tar_target(icr_data_long, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_blank_corrected),
+  tar_target(icr_data_trt, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_trt),
   
   
   # analysis - graphs
@@ -77,6 +77,8 @@ list(
   tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
   tar_target(gg_ferrozine, plot_ferrozine(ferrozine_processed, sample_key)),
   tar_target(gg_mehlich, plot_mehlich(mehlich_processed, sample_key)),
+  tar_target(gg_icr_vankrevelen, plot_vankrevelen(icr_data_trt, icr_meta)),
+  tar_target(gg_icr_unique, plot_vankrevelen_unique(icr_data_trt, icr_meta)),
   
 
   # report  

### 2022-12-21 2-code/a-functions_processing.R
ions/IC import and processing

@@ -567,3 +567,97 @@ process_mehlich = function(mehlich_map, mehlich_data, moisture_processed, subsam
   samples2
   
 }
+
+# Ions (IC)
+import_ions = function(FILEPATH){
+  
+  anions <- 
+    list.files(path=FILEPATH, pattern = c("anion", ".csv"), full.names = TRUE) %>% 
+    lapply(read_csv, skip = 5, id = "source") %>% 
+    bind_rows %>% 
+    rename(analysis_ID = "...3")
+  
+  cations <- 
+    list.files(path=FILEPATH, pattern = c("cation", ".csv"), full.names = TRUE) %>% 
+    lapply(read_csv, skip = 5, id = "source") %>% 
+    bind_rows %>% 
+    rename(analysis_ID = "...3")
+  
+  ions <- 
+    full_join(cations, anions, by = "analysis_ID")
+  
+  ions
+}
+#ions_data = import_ions(FILEPATH = "1-data/ions")
+
+process_ions = function(ions_data, analysis_key, sample_key, moisture_processed, subsampling){
+  
+  ions = 
+    ions_data %>% 
+    mutate_all(as.character) %>% 
+    pivot_longer(-c(analysis_ID, source.x, source.y)) %>% 
+    mutate(date_run = str_extract(source.x, "[0-9]{8}"),
+           date_run = lubridate::ymd(date_run)) %>% 
+    dplyr::select(-starts_with("source")) %>% 
+    filter(!grepl("...[0-9]", name)) %>% 
+    filter(!name %in% c("Nitrate", "Nitrite")) %>% 
+    mutate(name = str_remove(name, " UV"),
+           value = as.numeric(value)) %>% 
+    filter(!is.na(value)) %>% 
+    rename(ppm = value,
+           ion = name)
+  
+  samples = 
+    ions %>% 
+    filter(grepl("DOC_CMPS", analysis_ID))
+  
+  standards = 
+    ions %>% 
+    filter(grepl("x", analysis_ID))
+  
+  standards_max = 
+    standards %>% 
+    group_by(ion) %>% 
+    dplyr::summarise(max = max(ppm))
+
+  high_samples = 
+    samples %>% 
+    left_join(standards_max) %>% 
+    mutate(HIGH = ppm > max) %>% 
+    filter(HIGH)
+    
+#  high_samples %>% write.csv("high.csv")
+
+  dilutions = 
[…]

### 2022-12-21 _targets.R

@@ -24,6 +24,7 @@ source("2-code/0-packages.R")
 #source("2-code/1-initial_processing.R")
 source("2-code/a-functions_processing.R")
 source("2-code/b-functions_analysis.R")
+source("2-code/c-functions_fticrrr.R")
 # source("other_functions.R") # Source other scripts as needed. # nolint
 
 # Replace the target list below with your own:
@@ -65,6 +66,8 @@ list(
   tar_target(icr_data_long, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_blank_corrected),
   tar_target(icr_data_trt, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_trt),
   tar_target(icr_relabund_samples, compute_icr_relabund(icr_data_long, icr_meta)),
+  tar_target(ions_data, import_ions(FILEPATH = "1-data/ions")),
+  tar_target(ions_processed, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)),
   
   
   # analysis - graphs

### 2022-12-21 3-reports/characterization_report.Rmd
ferrozine graphs
mehlich-extractable P
icr graphs

@@
 
+**HCl-extractable iron - ferrozine method**
+
+Extracted using 0.5 M HCl
+
+```{r iron-ferrozine, fig.height=5, fig.width=7}
+tar_read(gg_ferrozine)
+```
+
+**Mehlich-3 extractable P**
+
+Extracted using Mehlich-3 solution, analyzed with molybdate method
+
+```{r phosphorus-mehlich, fig.height=5, fig.width=7}
+tar_read(gg_mehlich)
+```
+
+**FT-ICR-MS**
+
+```{r, icr-vk}
+tar_read(gg_icr_vankrevelen)
+```
+
+```{r, icr-unique}
+tar_read(gg_icr_unique)
+```
+
+```{r, icr-pca}
+tar_read(gg_icr_pca)
+```
+
 ---

### 2022-12-21 2-code/b-functions_analysis.R
plot ions

@@ -270,7 +270,43 @@ plot_mehlich = function(mehlich_processed, sample_key){
   
 }
 
-
+plot_ions = function(ions_processed, sample_key){
+  
+  ions = 
+    ions_processed %>% 
+    dplyr::select(sample_label, ends_with("_ug_g")) %>% 
+    pivot_longer(-sample_label, names_to = "species", values_to = "ug_g") %>% 
+    left_join(sample_key) %>% 
+    filter(!is.na(site)) %>% 
+    reorder_transect() %>% 
+    reorder_horizon()
+  
+  #  icp_long_outliers = 
+  #    icp_long %>% 
+  #    group_by(region, site, transect, horizon, species) %>% 
+  #    mutate(sd = sd(ppm),
+  #           mean = mean(ppm),
+  #           outlier = ppm > mean + (3*sd) | ppm < mean - (3*sd),
+  #           outlier2 = case_when(outlier == TRUE ~ 1,
+  #                                outlier == FALSE ~ 0))
+  
+  
+  
+  ions_subset = 
+    ions %>% 
+    mutate(species = str_remove(species, "_ug_g")) %>% 
+    filter(species %in% c("Sodium", "Potassium", "Magnesium",
+                          "Calcium", "Chloride", "Sulfate", "Nitrate", "Phosphate"))
+  
+  ions_subset %>% 
+    ggplot(aes(x = site, y = ug_g, color = transect, shape = horizon)) +
+    geom_point(position = position_dodge(width = 0.3))+
+    facet_wrap(species ~ region, scales = "free")+
+    scale_color_manual(values = pal_transect)
+  
+  
+  
+}
 
 
 

### 2022-12-21 _targets.R

@@ -84,6 +84,7 @@ list(
   tar_target(gg_icr_vankrevelen, plot_vankrevelen(icr_data_trt, icr_meta)),
   tar_target(gg_icr_unique, plot_vankrevelen_unique(icr_data_trt, icr_meta)),
   tar_target(gg_icr_pca, compute_icr_pca(icr_relabund_samples, sample_key)),
+  tar_target(gg_ions, plot_ions(ions_processed, sample_key)),
   
 
   # report  

### 2022-12-21 2-code/a-functions_processing.R
add "analysis" column to all processed outputs. combine all processed results

@@ -39,7 +39,8 @@ process_moisture = function(moisture_dat){
     # ((wet-dry)/dry) * 100
     mutate(gwc_perc = 100 * (wt_tray_moist_soil_g - wt_tray_dry_soil_g)/(wt_tray_dry_soil_g - wt_tray_g),
            gwc_perc = round(gwc_perc, 2)) %>% 
-    dplyr::select(sample_label, gwc_perc)
+    dplyr::select(sample_label, gwc_perc) %>% 
+    mutate(analysis = "GWC")
   
   
   #
@@ -53,7 +54,8 @@ process_loi = function(loi_data){
   loi_data %>%
     mutate_at(vars(contains("wt_")), as.numeric) %>% 
   mutate(percent_om = ((wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g))*100,
-         percent_om = round(percent_om, 2))  
+         percent_om = round(percent_om, 2))  %>% 
+    mutate(analysis = "LOI")
   
 }
 
@@ -64,7 +66,8 @@ process_pH = function(pH_data){
   dplyr::select(sample_label, pH, specific_conductance_ms_cm) %>% 
   mutate_at(vars(-sample_label), as.numeric)  %>% 
   filter(grepl("COMPASS", sample_label)) %>% 
-  filter(!is.na(pH) | !is.na(specific_conductance_ms_cm))
+  filter(!is.na(pH) | !is.na(specific_conductance_ms_cm))%>% 
+    mutate(analysis = "PH")
 }
 
 
@@ -124,7 +127,8 @@ process_tctnts = function(tctn_data, ts_data){
     tctn_samples %>% 
     left_join(ts_samples) %>% 
     rename(sample_label = Name) %>% 
-      arrange(sample_label)
+      arrange(sample_label) %>% 
+      mutate(analysis = "TCTNTS")
   
 }
 
@@ -170,7 +174,8 @@ process_weoc = function(weoc_data, analysis_key, moisture_processed, subsampling
   
   npoc_samples = 
     npoc_processed %>% 
-    filter(grepl("COMPASS", sample_label))
+    filter(grepl("COMPASS", sample_label))%>% 
+    mutate(analysis = "NPOC")
   
   npoc_samples
 }
@@ -221,7 +226,8 @@ process_din = function(din_data, analysis_key, moisture_processed, subsampling){
[…]

### 2022-12-21 2-code/b-functions_analysis.R
analyte completion matrix. correlation matrix. overall PCA.

@@ -312,55 +312,29 @@ plot_ions = function(ions_processed, sample_key){
 
 
 
+compute_analysis_matrix = function(data_combined, sample_key){
 
-combine_data = function(){
-  
-  df_list = list(moisture_processed, pH_processed, tctnts_data_samples, weoc_processed, din_processed, icp_processed, ferrozine_processed)
-  all_combined = 
-    df_list %>% reduce(full_join) %>% 
-    dplyr::select(-notes, -ends_with(c("_ppm", "_mgL", "_flag")))
-  
-  all_combined
- #   left_join(sample_key) %>% 
- #   filter(!grepl("_vac|rep", sample_label)) %>% 
- #   dplyr::select(-S_ug_g) %>% 
- #   #rowwise() %>% 
- #   #mutate(SUM = rowSums(across(where(is.numeric))))
- #   #replace(.,is.na(.),0) %>% 
- #   force()
-}
-
-#all_data_combined = combine_data()
-
-compute_analyte_matrix = function(all_data_combined, sample_key){
-  
-  analyte_completion_matrix = 
-    all_data_combined %>% 
-    distinct()
-  #  group_by(sample_label) %>% 
-  #  dplyr::summarise(n = n())
-  #  pivot_longer(-sample_label) %>% 
+  #data_combined2 = 
+  data_combined %>% 
+    distinct(sample_label, analysis) %>% 
+    filter(!is.na(analysis)) %>% 
     left_join(sample_key) %>% 
-    group_by(name, region, site, transect, horizon) %>% 
+    filter(!is.na(region)) %>% 
+    group_by(region, site, transect, horizon, analysis) %>% 
     dplyr::summarise(n = n()) %>% 
-    pivot_wider(values_from = "n")
-  
+    pivot_wider(values_from = "n", names_from = "analysis") %>% 
+    reorder_transect() %>% reorder_horizon() %>% 
+    arrange(desc(region), site, transect, horizon)
   
 }
 
-#analyte_completion_matrix = compute_analyte_matrix(all_data_combined, sample_key)
 
 
-
-
-compute_pca = function(){
+compute_overall_pca = function(data_combined_clean_surface, sample_key){
   library(ggbiplot)
   
-
-  
-  
   fit_pca_function = function(dat){
-
+    
     dat %>% 
       drop_na()
     
@@ -368,9 +342,8 @@ compute_pca = function(){
       dat %>%       
       dplyr::select(where(is.numeric)) %>%
       dplyr::mutate(row = row_number()) %>% 
[…]

### 2022-12-21 _targets.R

@@ -69,7 +69,6 @@ list(
   tar_target(ions_data, import_ions(FILEPATH = "1-data/ions")),
   tar_target(ions_processed, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)),
   
-  
   # analysis - graphs
   tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),
   tar_target(gg_loi, plot_loi(loi_processed, sample_key)),
@@ -87,6 +86,20 @@ list(
   tar_target(gg_ions, plot_ions(ions_processed, sample_key)),
   
 
+  # combined data
+  tar_target(data_combined, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
+                                         weoc_processed, din_processed, icp_processed, 
+                                         ferrozine_processed, mehlich_processed, ions_processed, sample_key)$data_combined),
+  tar_target(data_combined_clean_surface, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
+                                                       weoc_processed, din_processed, icp_processed, 
+                                                       ferrozine_processed, mehlich_processed, ions_processed, sample_key)$combined_surface),
+  tar_target(analysis_completion_matrix, compute_analysis_matrix(data_combined, sample_key)),
+  tar_target(gg_pca_all, compute_overall_pca(data_combined_clean_surface, sample_key)),
+  tar_target(gg_correlations, compute_correlations(data_combined_clean_surface, sample_key)),
+  
+  
+  
+  
   # report  
   tar_render(report, path = "3-reports/characterization_report.Rmd")
   

### 2022-12-21 2-code/c-functions_fticrrr.R
rename fticrrr-functions

@@ -343,33 +343,33 @@ compute_icr_relabund = function(icr_data_long, icr_meta){
     dplyr::mutate(total = sum(abund),
                   relabund  = ((abund/total)*100))
 }
-icr_relabund_samples = compute_icr_relabund(icr_data_long, icr_meta)
+## #icr_relabund_samples = compute_icr_relabund(icr_data_long, icr_meta)
 
 
-fticr_relabund_per_sample %>%
-  left_join(sample_key) %>% 
-  ggplot(aes(x = sample_label, y = relabund, fill = Class))+
-  geom_bar(stat = "identity")+
-  facet_wrap(~region, scales = "free_x")
+## fticr_relabund_per_sample %>%
+##   left_join(sample_key) %>% 
+##   ggplot(aes(x = sample_label, y = relabund, fill = Class))+
+##   geom_bar(stat = "identity")+
+##   facet_wrap(~region, scales = "free_x")
 
-fticr_relabund_per_sample %>%
-  left_join(sample_key) %>% 
-  filter(region == "CB" & horizon != "B") %>% 
-  ggplot(aes(x = sample_label, y = relabund, fill = Class))+
-  geom_bar(stat = "identity")+
-  facet_wrap(~site + transect, scales = "free_x")
+## fticr_relabund_per_sample %>%
+##   left_join(sample_key) %>% 
+##   filter(region == "CB" & horizon != "B") %>% 
+##   ggplot(aes(x = sample_label, y = relabund, fill = Class))+
+##   geom_bar(stat = "identity")+
+##   facet_wrap(~site + transect, scales = "free_x")
 
-fticr_relabund_per_sample %>%
-  left_join(sample_key) %>% 
-  filter(region == "WLE" & horizon != "B") %>% 
-  ggplot(aes(x = sample_label, y = relabund, fill = Class))+
-  geom_bar(stat = "identity")+
-  facet_wrap(~site + transect, scales = "free_x")
+## fticr_relabund_per_sample %>%
+##   left_join(sample_key) %>% 
+##   filter(region == "WLE" & horizon != "B") %>% 
+##   ggplot(aes(x = sample_label, y = relabund, fill = Class))+
+##   geom_bar(stat = "identity")+
+##   facet_wrap(~site + transect, scales = "free_x")
 
 
-y = fticr_relabund_per_sample %>% 
-  group_by(sample_label) %>% 
-  dplyr::summarise(total2 = sum(relabund))
[…]

### 2022-12-21 2-code/b-functions_analysis.R
use ggcorrplot

@@ -445,12 +445,16 @@ compute_correlations = function(data_combined_clean_surface, TITLE){
     
     
     m = cor(num_clean)
-    corrplot::corrplot(m, #method="color", 
-                       type = "lower",
-                       title = TITLE,
-                       mar = c(0,0,1,0) 
-                       #, order = "hclust"
-    )
+ #   x = corrplot::corrplot(m, #method="color", 
+ #                      type = "lower",
+ #                      title = TITLE,
+ #                      mar = c(0,0,1,0) 
+ #                      #, order = "hclust"
+ #                      )
+                       
+    ggcorrplot::ggcorrplot(m, type = "lower",
+                           title = TITLE)
+    
     
     
     

### 2022-12-21 3-reports/characterization_report.Rmd
plot ions
overall summaries in report

@@
                       warning = FALSE)
+```
+
+---
+
+## OVERALL SUMMARIES
+
+```{r analyte-completion}
+tar_read(analysis_completion_matrix) %>% knitr::kable()
+```
+
+```{r overall-pca}
+tar_read(gg_pca_all)
+```
+
+```{r correlations-all}
+tar_read(gg_correlations)
 ```
@@
 
-```{r, icr-vk}
+```{r, icr-vk, fig.height=16, fig.width=16}
 tar_read(gg_icr_vankrevelen)
@@
 
-```{r, icr-unique}
+```{r, icr-unique, fig.height=16, fig.width=16}
 tar_read(gg_icr_unique)
@@
 
-```{r, icr-pca}
-tar_read(gg_icr_pca)
+```{r, icr-pca, fig.height=5, fig.width=5}
+tar_read(gg_icr_pca)$biplot_all
+```
+
+```{r, icr-pca-regions, fig.height=5, fig.width=10}
+tar_read(gg_icr_pca)$biplot_regions
+```
+
+```{r, ions, fig.height=16, fig.width=16}
+tar_read(gg_ions)
 ```

### 2022-12-23 2-code/a-functions_processing.R
fix ferrozine fe dilution. data_combined_wide

@@ -365,7 +365,9 @@ process_iron = function(ferrozine_map, ferrozine_data, moisture_processed, subsa
       calibrate_ferrozine_data(data_processed) %>% 
       filter(grepl("COMPASS", sample_label)) %>% 
       dplyr::select(region, sample_label, analysis, ppm_calculated) %>% 
-      mutate(analysis = recode(analysis, "Fe2 (water only)" = "Fe2", "total-Fe (ascorbic)" = "Fe_total")) %>% 
+      mutate(analysis = recode(analysis, "Fe2 (water only)" = "Fe2", "total-Fe (ascorbic)" = "Fe_total"),
+             ppm_calculated = case_when(analysis == "Fe_total" ~ ppm_calculated * 2,
+                                        analysis == "Fe2" ~ ppm_calculated)) %>% 
       pivot_wider(names_from = "analysis", values_from = "ppm_calculated") %>% 
       mutate(across(where(is.numeric), round, 2)) %>% 
       mutate(Fe3 = Fe_total - Fe2)
@@ -619,7 +621,7 @@ process_ions = function(ions_data, analysis_key, sample_key, moisture_processed,
     filter(!is.na(value)) %>% 
     rename(ppm = value,
            ion = name) %>% 
-    filter(!ion %in% "Lithium")
+    filter(!ion %in% c("Lithium", "Nitrite"))
   
   samples = 
     ions %>% 
@@ -715,9 +717,26 @@ combine_data = function(moisture_processed, pH_processed, tctnts_data_samples,
     left_join(sample_key) %>% 
     mutate(transect = recode(transect, "wc" = "wetland")) %>% 
     reorder_transect() %>% 
+    reorder_horizon() %>% 
     force() 
   
-  data_combined
+  data_combined_wide = 
+    data_combined %>% 
+    separate(name, sep = "_", into = "variable", remove = F) %>% 
+    mutate(name = paste0(variable, " (", analysis, ")")) %>% 
+    dplyr::select(sample_label, name, value) %>% 
+    pivot_wider() %>% 
+    #  dplyr::select(-c("Ammonia (IC)","Bromide (IC)", "Nitrite (IC)", "Fluoride (IC)")) %>% 
+    left_join(sample_key) %>% 
+    #  mutate(`Phosphate (IC)` = case_when(is.na(`Phosphate (IC)`) & site != "GCREW" ~ 0,
+    #                                      TRUE ~ `Phosphate (IC)`)) %>% 
[…]

### 2022-12-23 2-code/b-functions_analysis.R
clean up pca, correlations

@@ -317,13 +317,12 @@ OLD_plots = function(){
 }
 
 
-compute_analysis_matrix = function(data_combined, sample_key){
+compute_analysis_matrix = function(data_combined){
 
   #data_combined2 = 
   data_combined %>% 
-    distinct(sample_label, analysis) %>% 
+    distinct(sample_label, analysis, region, site, transect, horizon) %>% 
     filter(!is.na(analysis)) %>% 
-    left_join(sample_key) %>% 
     filter(!is.na(region)) %>% 
     group_by(region, site, transect, horizon, analysis) %>% 
     dplyr::summarise(n = n()) %>% 
@@ -335,7 +334,7 @@ compute_analysis_matrix = function(data_combined, sample_key){
 
 
 
-compute_overall_pca = function(data_combined_clean_surface, sample_key){
+compute_overall_pca = function(data_combined_wide, sample_key){
   library(ggbiplot)
   
   fit_pca_function = function(dat){
@@ -366,14 +365,21 @@ compute_overall_pca = function(data_combined_clean_surface, sample_key){
          pca_int = pca_int)
   }
   
-  data_combined_clean_surface_no_ferrozine = 
-    data_combined_clean_surface %>% 
+  combined_surface = 
+    data_combined_wide %>% 
+    filter(horizon != "B") %>% 
+    dplyr::select(-"Bromide (IC)") %>% 
+    filter(sample_label != "COMPASS_Dec2021_016") %>% 
+    force() 
+  
+  combined_surface_no_ferrozine = 
+    combined_surface %>% 
     dplyr::select(-ends_with("(Ferrozine)"))
   
   ## PCA input files ----
-  pca_overall = fit_pca_function(data_combined_clean_surface_no_ferrozine)
-  pca_overall_wle = fit_pca_function(data_combined_clean_surface_no_ferrozine %>% filter(region == "WLE"))
-  pca_overall_cb = fit_pca_function(data_combined_clean_surface %>% filter(region == "CB") %>% dplyr::select(-ends_with("(IC)"), -"S (ICP)"))
+  pca_overall = fit_pca_function(combined_surface_no_ferrozine %>% dplyr::select(-ends_with("(IC)"), -"S (ICP)"))
+  pca_overall_wle = fit_pca_function(combined_surface_no_ferrozine %>% filter(region == "WLE"))
[…]

### 2022-12-23 _targets.R
fix targets

@@ -88,13 +88,15 @@ list(
   # combined data
   tar_target(data_combined, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
                                          weoc_processed, din_processed, icp_processed, 
-                                         ferrozine_processed, mehlich_processed, ions_processed, sample_key)),
-  tar_target(data_combined_clean_surface, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
-                                                       weoc_processed, din_processed, icp_processed, 
-                                                       ferrozine_processed, mehlich_processed, ions_processed, sample_key)$combined_surface),
-  tar_target(analysis_completion_matrix, compute_analysis_matrix(data_combined, sample_key)),
-  tar_target(gg_pca_all, compute_overall_pca(data_combined_clean_surface, sample_key)),
-  tar_target(gg_correlations, compute_correlations(data_combined_clean_surface, sample_key)),
+                                         ferrozine_processed, mehlich_processed, ions_processed, 
+                                         sample_key)$data_combined),
+  tar_target(data_combined_wide, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
+                                         weoc_processed, din_processed, icp_processed, 
+                                         ferrozine_processed, mehlich_processed, ions_processed, 
+                                         sample_key)$data_combined_wide),
+  tar_target(analysis_completion_matrix, compute_analysis_matrix(data_combined)),
+  tar_target(gg_pca_all, compute_overall_pca(data_combined_wide, sample_key)),
+  tar_target(gg_correlations, compute_correlations(data_combined_wide, sample_key)),
   
   
 tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),

### 2022-12-23 3-reports/characterization_report.Rmd
clean up markdown, simpler functions for all graphs

--- 2022-10-25 3-reports/characterization_report.Rmd
@@
 
-**Gravimetric moisture**
+## OVERALL SUMMARIES
 
-```{r moisture}
-tar_read(gg_moisture)
+### Analysis completion matrix
+
+<details>
+  <summary>click to open</summary>
+
+Number of reps available per analysis. Most were 8-9 reps, but some (e.g. WLE B-horizons) had fewer reps.
+
+```{r analyte-completion}
+tar_read(analysis_completion_matrix) %>% knitr::kable()
 ```
 
-**Loss on ignition (% organic matter)**
+</details>
 
-```{r loi}
-tar_read(gg_loi)
+### Overall trends - PCAs and Correlations
+
+<details>
+  <summary>click to open</summary>
+
+**PCAs**
+
+```{r overall-pca, fig.height=6, fig.width=6}
+tar_read(gg_pca_all)$gg_pca_overall
 ```
 
-**pH**
-
-```{r pH}
-tar_read(gg_pH)
+```{r overall-pca-regions, fig.height=6, fig.width=12}
+tar_read(gg_pca_all)$gg_pca_regions
 ```
 
-**specific conductance**
+**CORRELATIONS**
 
-```{r sp_cond}
-tar_read(gg_sp_cond)
+```{r correlations-all, fig.height=7, fig.width=7}
+tar_read(gg_correlations)$corr_all
 ```
 
-**total CNS**
-```{r tctnts}
-tar_read(gg_tctnts)
+```{r correlations-regions, fig.height=8, fig.width=18}
+tar_read(gg_correlations)$corr_regions
 ```
 
-**Water extractable organic C**
+</details>
 
-```{r weoc}
-tar_read(gg_weoc)
+---
+
+## EACH ANALYSIS
+
+### by transect
+
+<details>
+  <summary>click to open</summary>
+
+```{r by_transect, fig.height=4, fig.width=10}
+tar_read(gg_by_transect)
 ```
 
-**Extractable inorganic N**
+</details>
 
-Inorganic N was extracted using 2M KCl
+### by site - OA horizons
 
-```{r din}
-tar_read(gg_din)
+<details>
+  <summary>click to open</summary>
+
+```{r by_site_oa, fig.height=4, fig.width=10}
+tar_read(gg_by_site_oa)
 ```
 
-**ICP results: base cations**
+</details>
 
-Extracted using 1M NH4Cl
+### by site - OAB horizons
 
-```{r icp-base cations, fig.height=12, fig.width=7}
-tar_read(gg_icp)$gg_base_cations
+<details>
+  <summary>click to open</summary>
+
+```{r by_site_oab, fig.height=4, fig.width=10}
+tar_read(gg_by_site_oab)
 ```
 
-**ICP results: other cations**
+</details>
 
-Extracted using 1M NH4Cl
+---
 
-```{r icp-other cations, fig.height=9, fig.width=7}
-tar_read(gg_icp)$gg_other_cations
+## FT-ICR-MS
+
+<details>
+  <summary>click to open</summary>
+
+```{r, icr-vk-domains, fig.height=5, fig.width=5}
+tar_read(gg_icr_vankrevelen)$vk_domains
 ```
 
-**ICP results: others**
+```{r, icr-vk-regions, fig.height=7, fig.width=16}
+tar_read(gg_icr_vankrevelen)$vk_wle
+tar_read(gg_icr_vankrevelen)$vk_cb
 
-Extracted using 1M NH4Cl
-
-```{r icp-others, fig.height=4, fig.width=7}
-tar_read(gg_icp)$gg_others
 ```
 
+```{r, icr-unique, fig.height=16, fig.width=16}
+tar_read(gg_icr_unique)
+```
+
+```{r, icr-pca, fig.height=5, fig.width=5}
+tar_read(gg_icr_pca)$biplot_all
+```
+
+```{r, icr-pca-regions, fig.height=5, fig.width=10}
+tar_read(gg_icr_pca)$biplot_regions
+```
+
+</details>
+
 ---
+
+## Session Info
 

### 2022-12-23 _targets.R

@@ -101,10 +101,10 @@ list(
   
 tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),
 tar_target(gg_by_site_oa, make_graphs_by_site(data_combined %>% filter(horizon != "B"))),
-tar_target(gg_by_site_oab, make_graphs_by_site(data_combined))
+tar_target(gg_by_site_oab, make_graphs_by_site(data_combined)),
 
   
   # report  
-#  tar_render(report, path = "3-reports/characterization_report.Rmd")
+  tar_render(report, path = "3-reports/characterization_report.Rmd")
   
 )

### 2022-12-28 2-code/b-functions_analysis.R
add title, subtitle to plotting functions

@@ -510,7 +510,7 @@ compute_correlations = function(data_combined_wide, TITLE){
 ## the function names describe the variable on the x-axis
 
 
-plot_transect_as_x = function(data, YLAB){
+plot_transect_as_x = function(data, YLAB, TITLE = "", SUBTITLE = ""){
   
   ggplot(data,
          aes(x = transect,
@@ -531,6 +531,8 @@ plot_transect_as_x = function(data, YLAB){
     theme_kp()+
     labs(x = "",
          y = YLAB,
+         title = TITLE,
+         subtitle = SUBTITLE,
          shape = "Site",
          alpha = "Horizon",
          color = "Horizon")+
@@ -565,17 +567,19 @@ plot_region_as_x = function(data, YLAB){
     NULL
   
 }
-plot_site_as_x = function(data, YLAB){
+plot_site_as_x = function(data, YLAB, TITLE = "", SUBTITLE = ""){
   
   data %>% 
     ggplot(aes(x = site, 
                y = value, 
                color = transect, shape = horizon))+
     geom_point(position = position_dodge(width = 0.4),
-               size = 2)+
+               size = 2, stroke = 1)+
     facet_wrap(region ~ ., scales = "free_x")+
     labs(y = YLAB,
-         x = "")+
+         x = "",
+         title = TITLE,
+         subtitle = SUBTITLE)+
     scale_color_manual(values = pal_transect)+
     theme(legend.position = "right", legend.box = "vertical")+
     NULL
@@ -590,48 +594,52 @@ make_graphs_by_transect = function(data_combined){
     force() 
   
   gwc <- combined_surface %>% filter(analysis == "GWC")
-  gg_gwc <- plot_transect_as_x(data = gwc, YLAB = "Gravimetric water, %")
+  gg_gwc <- plot_transect_as_x(data = gwc, YLAB = "Gravimetric water, %", TITLE = "Gravimetric Water Content")
   
   pH <- combined_surface %>% filter(analysis == "PH")
-  gg_ph <- plot_transect_as_x(data = pH %>% filter(name == "pH"), YLAB = "pH")
[…]

### 2022-12-28 _targets.R
update

@@ -80,9 +80,9 @@ list(
   ## tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
   ## tar_target(gg_ferrozine, plot_ferrozine(ferrozine_processed, sample_key)),
   ## tar_target(gg_mehlich, plot_mehlich(mehlich_processed, sample_key)),
-  ## tar_target(gg_icr_vankrevelen, plot_vankrevelen(icr_data_trt, icr_meta)),
-  ## tar_target(gg_icr_unique, plot_vankrevelen_unique(icr_data_trt, icr_meta)),
-  ## tar_target(gg_icr_pca, compute_icr_pca(icr_relabund_samples, sample_key)),
+   tar_target(gg_icr_vankrevelen, plot_vankrevelen(icr_data_trt, icr_meta)),
+   tar_target(gg_icr_unique, plot_vankrevelen_unique(icr_data_trt, icr_meta)),
+   tar_target(gg_icr_pca, compute_icr_pca(icr_relabund_samples, sample_key)),
   ## tar_target(gg_ions, plot_ions(ions_processed, sample_key)),
   
   # combined data

### 2022-12-28 2-code/c-functions_fticrrr.R
clean up graphs

@@ -267,13 +267,20 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
   
   data_hcoc = 
     icr_data_trt %>% 
-    left_join(icr_meta %>% dplyr::select(formula, HC, OC))
+    filter(horizon != "B") %>% 
+    left_join(icr_meta %>% dplyr::select(formula, HC, OC)) %>% 
+    mutate(transect = recode(transect, "wc" = "wetland")) %>% 
+    reorder_horizon() %>% reorder_transect()
   
   vk_wle = 
     data_hcoc %>% 
-    filter(region == "WLE" & horizon != "B") %>% 
+    filter(region == "WLE") %>% 
     gg_vankrev(aes(x = OC, y = HC, color = transect))+
     stat_ellipse(level = 0.90, show.legend = F)+
+    scale_color_manual(values = pal_transect)+
+    labs(color = "",
+         title = "FTICR: all peaks (blank corrected)",
+         subtitle = "WLE sites")+
     facet_wrap(~site+horizon)
   
   vk_cb = 
@@ -281,6 +288,11 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
     filter(region == "CB" ) %>% 
     gg_vankrev(aes(x = OC, y = HC, color = transect))+
     stat_ellipse(level = 0.90, show.legend = F)+
+    scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"),
+                       values = pal_transect)+
+    labs(color = "",
+         title = "FTICR: all peaks (blank corrected)",
+         subtitle = "CB sites")+
     facet_wrap(~site+horizon)
   
   
@@ -293,10 +305,14 @@ plot_vankrevelen_unique = function(icr_data_trt, icr_meta){
   
   unique_hcoc = 
     icr_data_trt %>% 
+    filter(horizon != "B") %>% 
+    filter(!(site == "MSM" & horizon == "A")) %>% 
     group_by(formula, region, site, horizon) %>% 
     dplyr::mutate(n = n()) %>% 
     filter(n == 1) %>% 
-    left_join(icr_meta %>% dplyr::select(formula, HC, OC))
+    left_join(icr_meta %>% dplyr::select(formula, HC, OC)) %>% 
+    mutate(transect = recode(transect, "wc" = "wetland")) %>% 
+    reorder_horizon() %>% reorder_transect()
   
   
   vk_unique_wle = 
@@ -304,7 +320,10 @@ plot_vankrevelen_unique = function(icr_data_trt, icr_meta){
[…]

### 2022-12-28 2-code/b-functions_analysis.R
add SCALES to the function

@@ -510,7 +510,7 @@ compute_correlations = function(data_combined_wide, TITLE){
 ## the function names describe the variable on the x-axis
 
 
-plot_transect_as_x = function(data, YLAB, TITLE = "", SUBTITLE = ""){
+plot_transect_as_x = function(data, YLAB, TITLE = "", SUBTITLE = "", SCALES = "free_x"){
   
   ggplot(data,
          aes(x = transect,
@@ -526,7 +526,7 @@ plot_transect_as_x = function(data, YLAB, TITLE = "", SUBTITLE = ""){
     scale_alpha_manual(values = c(1, 0.3))+
     scale_color_manual(values = c("grey60", "black"))+
     scale_shape_manual(breaks = c("CC", "PR", "OWC", "MSM", "GWI", "GCREW"), values = c(1,2,3,4,5,6))+
-    facet_wrap(~region, scales = "free_x",
+    facet_wrap(~region, scales = SCALES,
                labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
     theme_kp()+
     labs(x = "",
@@ -616,12 +616,14 @@ make_graphs_by_transect = function(data_combined){
   icp <- combined_surface %>% filter(analysis == "ICP")
   gg_icp_ca <- plot_transect_as_x(data = icp %>% filter(name == "Ca_ug_g"), YLAB = "Ca, μg/g", TITLE = "ICP-OES", SUBTITLE = "extracted with 1M NH4Cl")
   gg_icp_mg <- plot_transect_as_x(data = icp %>% filter(name == "Mg_ug_g"), YLAB = "Mg, μg/g", TITLE = "ICP-OES", SUBTITLE = "extracted with 1M NH4Cl")
-  gg_icp_na <- plot_transect_as_x(data = icp %>% filter(name == "Na_ug_g"), YLAB = "Na, μg/g", TITLE = "ICP-OES", SUBTITLE = "extracted with 1M NH4Cl")
+  gg_icp_na <- plot_transect_as_x(data = icp %>% filter(name == "Na_ug_g"), YLAB = "Na, μg/g", TITLE = "ICP-OES", SUBTITLE = "extracted with 1M NH4Cl",
+                                  SCALES = "free")
[…]

### 2022-12-28 3-reports/characterization_report.Rmd
update rmd. delete old image files

@@
 
-```{r, icr-vk-regions, fig.height=7, fig.width=16}
+```{r, icr-vk-wle, fig.height=5, fig.width=10}
 tar_read(gg_icr_vankrevelen)$vk_wle
+
+```
+
+```{r, icr-vk-cb, fig.height=10, fig.width=10}
 tar_read(gg_icr_vankrevelen)$vk_cb
@@
 
-```{r, icr-unique, fig.height=16, fig.width=16}
-tar_read(gg_icr_unique)
+```{r, icr-unique-wle, fig.height=10, fig.width=4}
+tar_read(gg_icr_unique)$vk_unique_wle
+```
+
+```{r, icr-unique-cb, fig.height=10, fig.width=7}
+tar_read(gg_icr_unique)$vk_unique_cb
 ```

### 2023-01-06 2-code/b-functions_analysis.R
add p-value cutoff for ggcorrplot. add cn ratio plot. mehlich

@@ -465,6 +465,8 @@ compute_correlations = function(data_combined_wide, TITLE){
     
     
     m = cor(num_clean)
+    p.mat <- ggcorrplot::cor_pmat(num_clean)
+    
  #   x = corrplot::corrplot(m, #method="color", 
  #                      type = "lower",
  #                      title = TITLE,
@@ -473,13 +475,13 @@ compute_correlations = function(data_combined_wide, TITLE){
  #                      )
                        
     ggcorrplot::ggcorrplot(m, type = "lower",
+                           p.mat = p.mat,
+                           outline.color = "black",
+                           #   lab = TRUE, 
+                           insig = "blank",
+                           colors = c("#E46726", "white", "#6D9EC1"),
                            title = TITLE)
-    
-    
-    
-    
-    
-    
+
   }
 
   combined_surface = 
@@ -605,6 +607,12 @@ make_graphs_by_transect = function(data_combined){
   gg_tc <- plot_transect_as_x(data = tctnts %>% filter(name == "TC_perc"), YLAB = "Total C, %", TITLE = "Elemental Analysis")
   gg_tn <- plot_transect_as_x(data = tctnts %>% filter(name == "TN_perc"), YLAB = "Total N, %", TITLE = "Elemental Analysis")
   gg_ts <- plot_transect_as_x(data = tctnts %>% filter(name == "TS_perc"), YLAB = "Total S, %", TITLE = "Elemental Analysis")
+
+  tctn <- tctnts %>% 
+    filter(name == c("TC_perc", "TN_perc")) %>% 
+    pivot_wider(names_from = "name", values_from = "value") %>% 
+    mutate(value = TC_perc/TN_perc)
+  gg_cn <- plot_transect_as_x(data = tctn, YLAB = "C:N", TITLE = "Elemental Analysis")
   
   weoc <- combined_surface %>% filter(analysis == "NPOC")
   gg_weoc <- plot_transect_as_x(data = weoc %>% filter(name == "npoc_ug_g"), YLAB = "WEOC, μg/g", TITLE = "Water Extractable Organic Carbon")
@@ -653,6 +661,7 @@ make_graphs_by_transect = function(data_combined){
        gg_sp_conduc = gg_sp_conduc,
        gg_tc =  gg_tc,
[…]

### 2023-03-03 2-code/c-functions_fticrrr.R
add +NULL to ggplots

@@ -281,7 +281,8 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
     labs(color = "",
          title = "FTICR: all peaks (blank corrected)",
          subtitle = "WLE sites")+
-    facet_wrap(~site+horizon)
+    facet_wrap(~site+horizon)+
+    NULL
   
   vk_cb = 
     data_hcoc %>% 
@@ -293,7 +294,8 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
     labs(color = "",
          title = "FTICR: all peaks (blank corrected)",
          subtitle = "CB sites")+
-    facet_wrap(~site+horizon)
+    facet_wrap(~site+horizon)+
+    NULL
   
   
   list(vk_domains = vk_domains,

### 2023-03-03 2-code/a-functions_processing.R
process and plot xrd

@@ -733,7 +733,7 @@ import_xrd = function(FILEPATH){
     df %>% filter(!grepl("-", File))
     }))
 }
-xrd_data = import_xrd(FILEPATH = "1-data/xrd")
+#xrd_data = import_xrd(FILEPATH = "1-data/xrd")
 
 
 process_xrd = function(xrd_data, sample_key){

### 2023-03-03 2-code/b-functions_analysis.R

@@ -786,3 +786,24 @@ make_graphs_by_site = function(data_combined){
        gg_ions_po4 = gg_ions_po4
   )
 }
+
+# xrd ----
+
+plot_xrd = function(xrd_processed){
+  
+  xrd_summary = 
+    xrd_processed %>% 
+    dplyr::select(-c(tree_number, File)) %>% 
+    pivot_longer(-c(sample_label, region, site, transect, horizon),
+                 names_to = "mineral",
+                 values_to = "percentage") %>% 
+    reorder_horizon() %>% 
+    reorder_transect()
+  
+  xrd_summary %>% 
+    ggplot(aes(x = sample_label, y = percentage, fill = mineral))+
+    geom_bar(stat = "identity")+
+    facet_grid(horizon ~ region + site + transect, scales = "free_x")
+    
+  
+}

### 2023-03-03 _targets.R

@@ -69,6 +69,8 @@ list(
   tar_target(icr_relabund_samples, compute_icr_relabund(icr_data_long, icr_meta)),
   tar_target(ions_data, import_ions(FILEPATH = "1-data/ions")),
   tar_target(ions_processed, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)),
+  tar_target(xrd_data, import_xrd(FILEPATH = "1-data/xrd")),
+  tar_target(xrd_processed= process_xrd(xrd_data, sample_key)),
   
   # analysis - graphs
   ## tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),
@@ -103,6 +105,7 @@ list(
   tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),
   tar_target(gg_by_site_oa, make_graphs_by_site(data_combined %>% filter(horizon != "B"))),
   tar_target(gg_by_site_oab, make_graphs_by_site(data_combined)),
+  tar_target(gg_xrd, plot_xrd(xrd_processed)),
   
   # export
   tar_target(export, {

### 2023-03-08 2-code/0-packages.R
update element_rect per new ggplot version

@@ -13,7 +13,7 @@ theme_kp <- function() {  # this for all the elements common across plots
           legend.text = element_text(size = 12),
           legend.key.size = unit(1.5, 'lines'),
           legend.background = element_rect(colour = NA),
-          panel.border = element_rect(color="black",size=1, fill = NA),
+          panel.border = element_rect(color="black",linewidth=1, fill = NA),
           
           plot.title = element_text(hjust = 0, size = 14),
           axis.text = element_text(size = 12, color = "black"),

### 2023-03-08 2-code/a-functions_processing.R
WRC and texture functions

@@ -445,7 +445,7 @@ import_wrc_data = function(FILEPATH){
 
 process_wrc = function(wrc_data){
   
-  wrc_processed <- 
+  #wrc_processed <- 
     wrc_data %>% 
     mutate(source = str_remove(source, ".xlsx")) %>% 
     separate(source, sep = "_", into = c("EC", "kit", "site", "transect")) %>% 
@@ -455,32 +455,75 @@ process_wrc = function(wrc_data){
     mutate(transect = tolower(transect),
            kpa_eval = round((10^pf_eval)/10,2),
            kpa_fit = round((10^pf_fit)/10,2)) %>% 
-    mutate(transect = recode(transect, "wetland" = "wc")) %>% 
+    mutate(transect = recode(transect, "wetland" = "wc"),
+           region = case_when(site %in% c("CC", "OWC", "PR") ~ "WLE",
+                              site %in% c("MSM", "GWI", "GCREW") ~ "CB")) %>% 
     reorder_transect()
   
+}
+
+
+# soil texture
+compute_texture = function(hydrometer_df){
   
- wrc_processed %>% 
-    filter(pf_fit >= 0 | pf_eval >= 0) %>% 
-    ggplot(aes(y = water_content_vol_percent))+
-    geom_line(aes(x = pf_fit))+
-    geom_point(aes(x = pf_eval), shape = 1)+
-    facet_wrap(~site+transect)
- 
- 
- wrc_processed %>% 
-   filter(pf_fit >= 0 | pf_eval >= 0) %>% 
-   ggplot(aes(y = water_content_vol_percent, color = transect))+
-   geom_line(aes(x = kpa_fit/1000))+
-   geom_point(aes(x = kpa_eval/1000), shape = 1)+
-   scale_x_log10(labels = scales::comma)+
-   scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
-   facet_wrap(~site)
+  hydrometer_data_processed = 
+    hydrometer_df %>% 
+    mutate_at(vars(-c(site, sample_id, date_started, notes)), as.numeric) %>% 
+    mutate(wt_dry_soil_g = (wt_jar_soil_g - wt_half_gallon_jar_g) + (wt_sieve_soil_53um_g - wt_sieve_g))
   
+  #
+  # II. COMPUTING PERCENT SAND-SILT-CLAY -----------------------------------
   
+  ## This function will use the equations provided in Gee & Bauder
+  ## to compute % sand, clay, silt
   
[…]

### 2023-03-08 2-code/b-functions_analysis.R

@@ -807,3 +807,57 @@ plot_xrd = function(xrd_processed){
     
   
 }
+
+
+
+# soil physics ----
+## water retention curves
+
+plot_wrc = function(wrc_processed){
+  
+  wrc_processed %>% 
+    filter(kpa_fit >= 0 | kpa_eval >= 0) %>% 
+    mutate(region = factor(region, levels = c("WLE", "CB"))) %>% 
+    ggplot(aes(y = water_content_vol_percent, color = transect))+
+    geom_line(aes(x = kpa_fit))+
+    geom_point(aes(x = kpa_eval), shape = 1)+
+    scale_x_log10(labels = scales::comma)+
+    scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
+    labs(color = "")+
+    facet_wrap(~region + site)
+  
+}
+
+## texture
+
+plot_texture = function(texture_processed){
+
+  # plot the soil texture triangle
+  library(ggtern) 
+  data(USDA)
+  
+  gg_texture = 
+    ggtern(data = texture_processed,
+           aes(x = percent_sand, y = percent_clay, z = percent_silt)) +
+    geom_polygon(data = USDA, 
+                 aes(x = Sand, y = Clay, z = Silt, group = Label),
+                 fill = NA, size = 0.3, alpha = 0.5, color = "grey30")+
+    geom_point(aes(color = site, shape = transect),
+               size = 3)+
+    theme_bw()+
+    theme_showarrows()+
+    theme_hidetitles()+
+    theme_clockwise() 
+  
+  USDA_text <- 
+    USDA  %>% 
+    group_by(Label) %>%
+    summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+    ungroup()
+  
+  #textures_names<-
+  gg_texture +
+    geom_text(data = USDA_text,
+              aes(x = Sand, y = Clay, z = Silt, label = Label),
+              size = 3, color = "grey30")
+}

### 2023-03-08 _targets.R

@@ -70,7 +70,14 @@ list(
   tar_target(ions_data, import_ions(FILEPATH = "1-data/ions")),
   tar_target(ions_processed, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)),
   tar_target(xrd_data, import_xrd(FILEPATH = "1-data/xrd")),
-  tar_target(xrd_processed= process_xrd(xrd_data, sample_key)),
+  tar_target(xrd_processed, process_xrd(xrd_data, sample_key)),
+  tar_target(wrc_data, import_wrc_data(FILEPATH = "1-data/wrc")),
+  tar_target(wrc_processed, process_wrc(wrc_data)),
+  tar_target(hydrometer_data, "1-data/particle_size.csv", format = "file"),
+  tar_target(hydrometer_df, read.csv(hydrometer_data)),
+  tar_target(texture_processed, compute_texture(hydrometer_df)),
+  tar_target(gg_wrc, plot_wrc(wrc_processed)),
+  tar_target(gg_texture, plot_texture(texture_processed)),
   
   # analysis - graphs
   ## tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),

### 2023-03-11 2-code/b-functions_analysis.R
fix oab graphs. fix pca and patchwork issues

@@ -370,6 +370,7 @@ compute_overall_pca = function(data_combined_wide, sample_key){
     filter(horizon != "B") %>% 
     dplyr::select(-"Bromide (IC)") %>% 
     filter(sample_label != "COMPASS_Dec2021_016") %>% 
+    dplyr::select(-ends_with("(IC)"), -"S (ICP)", -"Fe (ICP)", -"P (ICP)") %>% 
     force() 
   
   combined_surface_no_ferrozine = 
@@ -377,9 +378,9 @@ compute_overall_pca = function(data_combined_wide, sample_key){
     dplyr::select(-ends_with("(Ferrozine)"))
   
   ## PCA input files ----
-  pca_overall = fit_pca_function(combined_surface_no_ferrozine %>% dplyr::select(-ends_with("(IC)"), -"S (ICP)"))
-  pca_overall_wle = fit_pca_function(combined_surface_no_ferrozine %>% filter(region == "WLE"))
-  pca_overall_cb = fit_pca_function(combined_surface %>% filter(region == "CB") %>% dplyr::select(-ends_with("(IC)"), -"S (ICP)"))
+  pca_overall = fit_pca_function(combined_surface_no_ferrozine)
+  pca_overall_wle = fit_pca_function(combined_surface_no_ferrozine %>% filter(region == "WLE") %>% dplyr::select(-ends_with("(IC)")))
+  pca_overall_cb = fit_pca_function(combined_surface %>% filter(region == "CB") %>% dplyr::select(-ends_with("(IC)")))
   
   
   ## PCA plots overall ----
@@ -441,8 +442,11 @@ compute_overall_pca = function(data_combined_wide, sample_key){
     theme(legend.position = "top", legend.box = "vertical")+
     NULL
   
+  library(patchwork)
+  gg_pca_regions = gg_pca_wle + gg_pca_cb
+  
   list(gg_pca_overall = gg_pca_overall,
-       gg_pca_regions = gg_pca_wle + gg_pca_cb)
+       gg_pca_regions = gg_pca_regions)
   
 }
 
@@ -578,6 +582,7 @@ plot_site_as_x = function(data, YLAB, TITLE = "", SUBTITLE = ""){
     geom_point(position = position_dodge(width = 0.4),
                size = 2, stroke = 1)+
     facet_wrap(region ~ ., scales = "free_x")+
+    theme_kp()+
     labs(y = YLAB,
          x = "",
          title = TITLE,
[…]

### 2023-03-11 2-code/c-functions_fticrrr.R
fix patchwork issues

@@ -493,6 +493,7 @@ compute_icr_pca = function(icr_relabund_samples, sample_key){
     theme(legend.position = "top", legend.box = "vertical")+
     NULL
   
+  library(patchwork)
   biplot_regions = 
     biplot_wle + biplot_cb 
   

### 2023-03-11 2-code/b-functions_analysis.R
piper plot PRELIMINARY code

@@ -813,6 +813,221 @@ plot_xrd = function(xrd_processed){
   
 }
 
+plot_ions_piper = function(ions_processed){
+  
+  tic = read_tsv("1-data/tic_owc.txt", skip = 10) %>% janitor::clean_names()
+
+  tic_processed = 
+    tic %>% 
+    dplyr::select(sample_name, result_ic) %>% 
+    rename(analysis_ID = sample_name,
+           tic_mgL = result_ic) %>% 
+    # keep only sampple rows 
+    filter(grepl("DOC_", analysis_ID)) %>% 
+    # join the analysis key to get the sample_label
+    left_join(analysis_key %>% dplyr::select(analysis_ID, sample_label)) %>% 
+    mutate(Bicarbonate_ppm = tic_mgL * 5) %>% 
+    dplyr::select(sample_label, Bicarbonate_ppm) %>% 
+    left_join(sample_key)
+  
+  
+  
+  
+  
+  
+  ions_long = 
+    ions_processed %>% 
+    dplyr::select(sample_label, ends_with("ppm")) %>% 
+    left_join(tic_processed) %>% 
+    pivot_longer(-sample_label,
+                 names_to = "ion", 
+                 values_to = "ppm") %>% 
+    mutate(group = case_when(grepl("Calcium", ion) ~ "Calcium",
+                             grepl("Magnesium", ion) ~ "Magnesium",
+                             grepl("Sodium|Potassium|Ammoni", ion) ~ "Sodium",
+                             grepl("Chloride|Nitrate|Phosphate", ion) ~ "Chloride",
+                             grepl("Carbonate|Bicarbonate", ion) ~ "Bicarbonate",
+                             grepl("Sulfate", ion) ~ "Sulfate")) %>% 
+    mutate(cation_anion = case_when(grepl("Calcium|Magnesium|Sodium", group) ~ "cation",
+                                    grepl("Chloride|Bicarbonate|Sulfate", group) ~ "anion")) %>% 
+    filter(!is.na(group)) %>% 
+    group_by(sample_label, cation_anion, group) %>% 
+    dplyr::summarise(ppm = sum(ppm)) %>% 
+    group_by(sample_label, cation_anion) %>% 
+    dplyr::mutate(total = sum(ppm),
[…]

### 2023-03-11 2-code/a-functions_processing.R
loss on ignition data and code

@@ -47,15 +47,32 @@ process_moisture = function(moisture_dat){
 }
 
 
-# loss on ignition
+## Loss on Ignition
+import_loi = function(FILEPATH){
+  
+  df <- 
+    list.files(path=FILEPATH, pattern = ".csv", full.names = TRUE) %>% 
+    lapply(read_csv, id = "source") %>% 
+    bind_rows
+  
+  df
+}
 process_loi = function(loi_data){
   
-  #loi = 
   loi_data %>%
-    mutate_at(vars(contains("wt_")), as.numeric) %>% 
-  mutate(percent_om = ((wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g))*100,
-         percent_om = round(percent_om, 2))  %>% 
-    mutate(analysis = "LOI")
+    janitor::clean_names() %>% 
+    mutate(region = case_when(grepl("WLE", source) ~ "WLE",
+                              grepl("CB", source) ~ "CB")) %>% 
+    mutate(label = case_when((customer_id >= 129 & customer_id <= 214) ~ paste0("COMPASS_May22_", customer_id),
+                             customer_id >= 215 ~ paste0("COMPASS_Aug22_", customer_id)),
+           sample_label = case_when(region == "WLE" ~ sample_label,
+                                    region == "CB" ~ label)) %>% 
+    mutate(percent_om = case_when(region == "CB" ~ percent_om,
+                                  region == "WLE" ~ 100*(wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g)),
+           percent_om = round(percent_om, 2)) %>%
+    dplyr::select(sample_label, percent_om) %>% 
+    mutate(analysis = "LOI") %>% 
+    force()
   
 }
 
@@ -808,12 +825,12 @@ process_xrd = function(xrd_data, sample_key){
 
 
 # Combined data
-combine_data = function(moisture_processed, pH_processed, tctnts_data_samples, 
+combine_data = function(moisture_processed, pH_processed, tctnts_data_samples, loi_processed, 
                         weoc_processed, din_processed, icp_processed, cec_processed,
                         ferrozine_processed, mehlich_processed, ions_processed,
                         sample_key){
   
-  df_list = list(moisture_processed, pH_processed, tctnts_data_samples, 
[…]

### 2023-03-11 _targets.R

@@ -41,7 +41,7 @@ list(
   # data files
   tar_target(moisture_data, import_gsheet("1nYzExPmpv01IYalqo2L3pmDsBDvhjdvHYMJo31N5_MA")),
   tar_target(moisture_processed, process_moisture(moisture_data)),
-  tar_target(loi_data, import_gsheet("1vvNKOtewhjR0PVYCLfvpRxY9QiOWzUvRV0WX5ou2qT4")),
+  tar_target(loi_data, import_loi(FILEPATH = "1-data/loi")),
   tar_target(loi_processed, process_loi(loi_data)),
   tar_target(pH_data, import_gsheet("177ZR7O9JpsgVmFfLh72LfJ72oM-oTc6k_r1qE0XHL64")),
   tar_target(pH_processed, process_pH(pH_data)),
@@ -96,11 +96,11 @@ list(
   ## tar_target(gg_ions, plot_ions(ions_processed, sample_key)),
   
   # combined data
-  tar_target(data_combined, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
+  tar_target(data_combined, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
                                          weoc_processed, din_processed, icp_processed, cec_processed,
                                          ferrozine_processed, mehlich_processed, ions_processed, 
                                          sample_key)$data_combined),
-  tar_target(data_combined_wide, combine_data(moisture_processed, pH_processed, tctnts_data_samples, 
+  tar_target(data_combined_wide, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
                                          weoc_processed, din_processed, icp_processed, cec_processed,
                                          ferrozine_processed, mehlich_processed, ions_processed, 
                                          sample_key)$data_combined_wide),

### 2023-03-11 2-code/b-functions_analysis.R
add LOI to graphs

@@ -619,6 +619,9 @@ make_graphs_by_transect = function(data_combined){
     mutate(value = TC_perc/TN_perc)
   gg_cn <- plot_transect_as_x(data = tctn, YLAB = "C:N", TITLE = "Elemental Analysis")
   
+  loi <- combined_surface %>% filter(analysis == "LOI")
+  gg_loi <- plot_transect_as_x(data = loi, YLAB = "% OM", TITLE = "Loss on ignition")
+  
   weoc <- combined_surface %>% filter(analysis == "NPOC")
   gg_weoc <- plot_transect_as_x(data = weoc %>% filter(name == "npoc_ug_g"), YLAB = "WEOC, μg/g", TITLE = "Water Extractable Organic Carbon")
   
@@ -677,6 +680,7 @@ make_graphs_by_transect = function(data_combined){
        gg_tn = gg_tn,
        gg_cn = gg_cn,
        gg_ts = gg_ts,
+       gg_loi = gg_loi,
        gg_weoc = gg_weoc,
        gg_din_nh4n = gg_din_nh4n,
        gg_din_no3n = gg_din_no3n,
@@ -725,6 +729,9 @@ make_graphs_by_site = function(data_combined){
   gg_tn <- plot_site_as_x(data = tctnts %>% filter(name == "TN_perc"), YLAB = "Total N, %", TITLE = "Elemental Analysis")
   gg_ts <- plot_site_as_x(data = tctnts %>% filter(name == "TS_perc"), YLAB = "Total S, %", TITLE = "Elemental Analysis")
   
+  loi <- combined_surface %>% filter(analysis == "LOI")
+  gg_loi <- plot_site_as_x(data = loi, YLAB = "% OM", TITLE = "Loss on ignition")
+  
   weoc <- combined_surface %>% filter(analysis == "NPOC")
   gg_weoc <- plot_site_as_x(data = weoc %>% filter(name == "npoc_ug_g"), YLAB = "WEOC, μg/g", TITLE = "Water Extractable Organic Carbon")
   
@@ -768,6 +775,7 @@ make_graphs_by_site = function(data_combined){
        gg_tc =  gg_tc,
        gg_tn = gg_tn,
        gg_ts = gg_ts,
+       gg_loi = gg_loi,
        gg_weoc = gg_weoc,
        gg_din_nh4n = gg_din_nh4n,
        gg_din_no3n = gg_din_no3n,

### 2023-03-11 2-code/c-functions_fticrrr.R
rename pca function for icr

@@ -406,7 +406,7 @@ library(vegan)
 library(patchwork)
 
 
-fit_pca_function = function(icr_relabund_samples, sample_key){
+fit_pca_function_icr = function(icr_relabund_samples, sample_key){
   relabund_pca =
     icr_relabund_samples %>% 
     left_join(sample_key) %>% 
@@ -439,9 +439,9 @@ compute_icr_pca = function(icr_relabund_samples, sample_key){
     sample_key %>% 
     mutate(transect = recode(transect, "wc" = "wetland"))
   
-  pca_overall = fit_pca_function(icr_relabund_samples, sample_key)
-  pca_wle = fit_pca_function(icr_relabund_samples, sample_key %>% filter(region == "WLE"))
-  pca_cb = fit_pca_function(icr_relabund_samples, sample_key %>% filter(region == "CB"))
+  pca_overall = fit_pca_function_icr(icr_relabund_samples, sample_key)
+  pca_wle = fit_pca_function_icr(icr_relabund_samples, sample_key %>% filter(region == "WLE"))
+  pca_cb = fit_pca_function_icr(icr_relabund_samples, sample_key %>% filter(region == "CB"))
   
   
   # PCA biplots

### 2023-03-11 2-code/a-functions_processing.R
reorganizing the functions

@@ -18,15 +18,15 @@ reorder_horizon = function(dat){
   dat %>% 
     mutate(horizon = factor(horizon, levels = c("O", "A", "B")))
 }
-
 reorder_transect = function(dat){
   dat %>% 
     mutate(transect = factor(transect, levels = c("upland", "transition", "wte", "wc", "wetland")))
 }
 
 
-# process data ------------------------------------------------------------
-# Moisture
+# Chemistry data ----------------------------------------------------------
+
+## Moisture
 process_moisture = function(moisture_dat){
 #  moisture_processed = 
     moisture_dat %>% 
@@ -75,6 +75,7 @@ process_loi = function(loi_data){
     force()
   
 }
+
 
 # pH
 process_pH = function(pH_data){
@@ -342,8 +343,6 @@ import_iron = function(FILEPATH){
        ferrozine_data = ferrozine_data)
   
 }
-
-
 process_iron = function(ferrozine_map, ferrozine_data, moisture_processed, subsampling){
   
   # clean the map
@@ -438,109 +437,6 @@ process_iron = function(ferrozine_map, ferrozine_data, moisture_processed, subsa
     samples2
   }
 
-# Water retention curves
-import_wrc_data = function(FILEPATH){
-  
-  filePaths_wrc <- list.files(path = FILEPATH, pattern = "xlsx", full.names = TRUE, recursive = FALSE)
-  wrc_data <- do.call(bind_rows, lapply(filePaths_wrc, function(path) {
-    
-    # importing both, the evaluated values and the fitted values
-    df_eval <- readxl::read_excel(path, sheet = "Evaluation-Retention Θ(pF)") %>% mutate_all(as.character) %>% janitor::clean_names()
-    df_eval = df_eval %>% mutate(source = basename(path)) %>% dplyr::select(p_f, water_content_vol_percent, source) %>% rename(pf_eval = p_f)
-    
-    df_fit <- readxl::read_excel(path, sheet = "Fitting-Retention Θ(pF)") %>% mutate_all(as.character) %>% janitor::clean_names()
-    df_fit = df_fit %>% mutate(source = basename(path)) %>% dplyr::select(p_f, water_content_vol_percent, source) %>% rename(pf_fit = p_f)
-    
-    df <- full_join(df_eval, df_fit)
-    df
-  }
-  
-  ))
-
-}
[…]

### 2023-03-11 2-code/b-functions_analysis.R
update theme

@@ -816,7 +816,8 @@ plot_xrd = function(xrd_processed){
   xrd_summary %>% 
     ggplot(aes(x = sample_label, y = percentage, fill = mineral))+
     geom_bar(stat = "identity")+
-    facet_grid(horizon ~ region + site + transect, scales = "free_x")
+    facet_grid(horizon ~ region + site + transect, scales = "free_x")+
+    theme_kp()
     
   
 }
@@ -1053,7 +1054,8 @@ plot_wrc = function(wrc_processed){
     scale_x_log10(labels = scales::comma)+
     scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
     labs(color = "")+
-    facet_wrap(~region + site)
+    facet_wrap(~region + site)+
+    theme_kp()
   
 }
 

### 2023-03-11 3-reports/characterization_report.Rmd
update graphs
process and plot xrd
WRC and texture functions
update markdown report

@@
 
-```{r correlations-all, fig.height=7, fig.width=7}
+```{r correlations-all, fig.height=8, fig.width=8}
 tar_read(gg_correlations)$corr_all
@@
 
+## XRD
+```{r xrd, fig.height=4, fig.width=15}
+tar_read(gg_xrd)
+```
+
+---
+
+## Soil Physics
+
+### WRC
+
+```{r wrc, fig.height=7, fig.width=8}
+tar_read(gg_wrc)
+
+```
+
+### Texture
+```{r texture}
+tar_read(gg_texture)
+```
+
+---
+
 ## Session Info

### 2023-03-11 2-code/wrc.R
remove old wrc file

@@ -1,57 +1,0 @@
-# SOIL CHARACTERIZATION FOR WLE AND CB SITES
-
-## wrc.R
-## This script will import and clean data for water retention curves
-
-## KFP, July 2022
-
-######################## ####
-######################## ####
-
-# load packages -----------------------------------------------------------
-source("2-code/0-packages.R")
-
-#
-# import data -------------------------------------------------------------
-import_wrc_data = function(FILEPATH){
-  
-  filePaths_wrc <- list.files(path = FILEPATH, pattern = "*.xlsx", full.names = TRUE)
-  wrc_dat <- do.call(bind_rows, lapply(filePaths_wrc, function(path) {
-    df <- readxl::read_xlsx(path, sheet = "Fitting-Retention Θ(pF)")
-    df %<>%
-      rename(pF = `pF [-]`,
-             vol_water_perc = `Water Content [Vol%]`) %>% 
-      mutate(source = rep(path, nrow(.))) %>% 
-      force()
-    df}))
-  
-}
-
-wrc_data = import_wrc_data("1-data/wrc")
-
-#
-# process data ------------------------------------------------------------
-wrc_processed = 
-  wrc_data %>% 
-  separate(source, sep = "_EC1_", into = c("delete", "name")) %>% 
-  separate(name, sep = "_", into = c("EC_kit", "site", "transect")) %>% 
-  mutate(transect = str_remove(transect, ".xlsx"),
-         kPa = round((10^pF)/10,2)) %>% 
-  dplyr::select(-delete)
-
-#
-
-# graphs ------------------------------------------------------------------
-wrc_processed %>% 
-  filter(EC_kit != "K008") %>% 
-  filter(EC_kit == c("K004", "K011")) %>% 
-  ggplot(aes(x = kPa/1000, y = vol_water_perc, colour = transect))+
-  geom_line()+
-  scale_x_log10()+
-  facet_wrap(~site)
-
-#
-
-# export ------------------------------------------------------------------
-wrc_processed %>% write.csv("1-data/processed/wrc_2022-07-18.csv", row.names = F, na = "")
-

### 2023-03-11 2-code/scratchpad.R
scratchpad (delete later)

@@ -1,0 +1,143 @@
+
+
+loi %>% 
+  filter(percent_om < 25) %>% 
+  group_by(site, transect) %>% 
+  dplyr::summarise(mean = mean(percent_om))
+
+moisture %>% 
+  filter(site == "OWC") %>% 
+  group_by(site, transect, horizon) %>% 
+  dplyr::summarise(mean = mean(gwc_perc))
+
+pH %>% 
+  filter(site == "OWC") %>% 
+  group_by(site, transect, horizon) %>% 
+  dplyr::summarise(mean = mean(pH))
+
+
+tctnts %>% 
+  filter(site == "OWC") %>% 
+  group_by(site, transect, horizon) %>% 
+  dplyr::summarise(mean = mean(TC_perc))
+
+tctnts %>% 
+  filter(site == "OWC") %>% 
+  filter(TN_perc < 5) %>% 
+  group_by(site, transect, horizon) %>% 
+  dplyr::summarise(mean = mean(TN_perc))
+
+tctnts %>% 
+  filter(site == "OWC") %>% 
+  group_by(site, transect, horizon) %>% 
+  dplyr::summarise(mean = mean(TS_perc))
+
+
+
+
+icp_data= import_icp_data(FILEPATH = "1-data/icp")
+
+
+
+
+
+
+
+FILEPATH = "1-data/icp"
+analysis_key = read.csv("1-data/analysis_key.csv")
+sample_key = read.csv("1-data/sample_key.csv")
+
+
+
+
+
+ferrozine =   
+  samples %>% 
+  left_join(sample_key)
+
+
+ferrozine %>% 
+  mutate(Fe_2_3 = Fe2/Fe3) %>% 
+  filter(Fe_2_3 > 0) %>% 
+  ggplot(aes(x = site, y = Fe_2_3, color = transect, shape = horizon))+
+  geom_point(position = position_dodge(width = 0.4),
+             size = 2)+
+  facet_wrap(~region, scales = "free_x")
+
+
+ferrozine %>% 
+  
+  ggplot(aes(x = site, y = Fe_total, color = transect, shape = horizon))+
+  geom_point(position = position_dodge(width = 0.4),
+             size = 2)+
+  facet_wrap(~region, scales = "free_x")
+
+
+
+# -------------------------------------------------------------------------
+## ICR
+
+source("2-code/c-fticrrr-functions.R")
+icr_report= import_fticr_data(FILEPATH = "1-data/icr")
[…]

### 2023-03-11 2-code/c-functions_fticrrr.R
add theme_kp

@@ -262,7 +262,8 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
   vk_domains = 
     icr_meta %>% 
     dplyr::select(formula, HC, OC, Class_detailed, Class) %>% 
-    gg_vankrev(aes(x = OC, y = HC, color = Class))
+    gg_vankrev(aes(x = OC, y = HC, color = Class))+
+    theme_kp()
   
   
   data_hcoc = 
@@ -282,6 +283,7 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
          title = "FTICR: all peaks (blank corrected)",
          subtitle = "WLE sites")+
     facet_wrap(~site+horizon)+
+    theme_kp()+
     NULL
   
   vk_cb = 
@@ -295,6 +297,7 @@ plot_vankrevelen = function(icr_data_trt, icr_meta){
          title = "FTICR: all peaks (blank corrected)",
          subtitle = "CB sites")+
     facet_wrap(~site+horizon)+
+    theme_kp()+
     NULL
   
   
@@ -327,7 +330,8 @@ plot_vankrevelen_unique = function(icr_data_trt, icr_meta){
     labs(color = "")+
     facet_grid(site ~ horizon)+
     labs(title = "FTICR Unique Peaks",
-         subtitle = "WLE sites")
+         subtitle = "WLE sites")+
+    theme_kp()
   
   
   vk_unique_cb = 
@@ -340,7 +344,8 @@ plot_vankrevelen_unique = function(icr_data_trt, icr_meta){
     labs(color = "")+
     facet_grid(site ~ horizon)+
     labs(title = "FTICR Unique Peaks",
-         subtitle = "CB sites")
+         subtitle = "CB sites")+
+    theme_kp()
   
   
   list(vk_unique_wle = vk_unique_wle,
@@ -457,6 +462,7 @@ compute_icr_pca = function(icr_relabund_samples, sample_key){
     xlim(-4,4)+
     ylim(-3.5,3.5)+
     labs(color = "", shape = "")+
+    theme_kp()+
     theme(legend.position = "top", legend.box = "vertical")+
     NULL
   
@@ -473,6 +479,7 @@ compute_icr_pca = function(icr_relabund_samples, sample_key){
          color = "", shape = "")+
     xlim(-4,4)+
     ylim(-3.5,3.5)+
+    theme_kp()+
     theme(legend.position = "top", legend.box = "vertical")+
     NULL
   
[…]

### 2023-03-29 2-code/b-functions_analysis.R
fix patchwork vs. cowplot. add `sample_key` to function inputs

@@ -325,7 +325,7 @@ make_data_subset = function(data_combined){
     filter(!transect %in% "wte") %>% 
     filter(!grepl("Bromide|Fluoride|Nitrate|Calcium|Magnesium|Potassium|Sodium|Phosphate|Ammonia", name, ignore.case = TRUE))
 }
-make_data_wide = function(data_combined_subset){
+make_data_wide = function(data_combined_subset, sample_key){
   
   #data_combined_wide = 
     data_combined_subset %>% 
@@ -361,10 +361,11 @@ compute_overall_pca = function(data_combined, sample_key){
   library(ggbiplot)
   
   data_combined_subset = make_data_subset(data_combined)
-  data_combined_wide = make_data_wide(data_combined_subset)
+  data_combined_wide = make_data_wide(data_combined_subset, sample_key)
   data_combined_wide_NO_IC = make_data_wide(data_combined_subset %>% 
                                               filter(!analysis %in% "IC") %>% 
-                                              filter(!grepl("dic", name, ignore.case = TRUE)))
+                                              filter(!grepl("dic", name, ignore.case = TRUE)),
+                                            sample_key)
   
   fit_pca_function = function(dat){
     
@@ -468,14 +469,15 @@ compute_overall_pca = function(data_combined, sample_key){
   
 }
 
-compute_correlations = function(data_combined_wide, TITLE){
+compute_correlations = function(data_combined, sample_key){
   #library(corrplot)
   
   data_combined_subset = make_data_subset(data_combined)
-  data_combined_wide = make_data_wide(data_combined_subset)
+  data_combined_wide = make_data_wide(data_combined_subset, sample_key)
   data_combined_wide_NO_IC = make_data_wide(data_combined_subset %>% 
                                               filter(!analysis %in% "IC") %>% 
-                                              filter(!grepl("dic", name, ignore.case = TRUE)))
+                                              filter(!grepl("dic", name, ignore.case = TRUE)),
+                                            sample_key)
   
   fit_correlations_function = function(dat, TITLE){
     num = 
@@ -516,8 +518,9 @@ compute_correlations = function(data_combined_wide, TITLE){
   corr_wle = fit_correlations_function(dat = data_combined_wide_NO_IC %>% filter(region == "WLE"), TITLE = "WLE" )
[…]

### 2023-03-29 2-code/b-functions_analysis.R
fix function inputs

@@ -1329,13 +1329,13 @@ plot_site_as_color = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES
   
   # make summary table for HSD letters ----
   hsd_transect_overall = 
-    data_combined_subset %>% 
+    data %>% 
     group_by(analysis, name, region) %>% 
     do(fit_lme_hsd_transect(.))
   
   ## then calculate the y-axis position for the labels
   lme_y =
-    data_combined %>% 
+    data %>% 
     group_by(region, analysis, name) %>% 
     dplyr::summarise(max = max(value),
                      y = max + max/5) %>% 
@@ -1361,7 +1361,7 @@ plot_site_as_color = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES
                size = 2.5, stroke = 1,
                position = position_dodge(width = 0.4))+
     # plot HSD letters
-    geom_text(data = lme_for_graphs, aes(y = y, label = label), size = 5)+
+    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
     scale_alpha_manual(values = c(1, 0.3))+
     scale_color_manual(breaks = c("CC", "PR", "OWC", "MSM", "GWI", "GCREW"), values = c('#ED6e85', '#ffc115', '#7f4420', '#90BE6D', '#03045E', '#00B4D8'))+
     facet_wrap(~region, scales = SCALES,
@@ -1378,7 +1378,9 @@ plot_site_as_color = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES
     NULL
   
 }
-make_graphs_by_transect_SITE_AS_COLOR = function(data_combined_subset){
+make_graphs_by_transect_SITE_AS_COLOR = function(data_combined){
+  
+  data_combined_subset = make_data_subset(data_combined)
   
   mehlich_p = data_combined_subset %>% filter(name == "mehlichp_ugg")
   gg_wle_mehlich = plot_site_as_color(data = mehlich_p %>% filter(region == "WLE"), YLAB = "P, μg/g", TITLE = "Mehlich-3 extractable P", 

### 2023-03-29 2-code/b-functions_analysis.R
fix spConductance name

@@ -645,7 +645,7 @@ make_graphs_by_transect = function(data_combined){
   
   pH <- combined_surface %>% filter(analysis == "PH")
   gg_ph <- plot_transect_as_x(data = pH %>% filter(name == "pH"), YLAB = "pH", TITLE = "pH")
-  gg_sp_conduc <- plot_transect_as_x(data = pH %>% filter(grepl("specific_conductance", name)), 
+  gg_sp_conduc <- plot_transect_as_x(data = pH %>% filter(grepl("spConductance_mscm", name)), 
                                      YLAB = "Specific Conductance μS/cm", TITLE = "Specific Conductance")
   
   tctnts <- combined_surface %>% filter(analysis == "TCTNTS")
@@ -761,7 +761,7 @@ make_graphs_by_site = function(data_combined){
   
   pH <- combined_surface %>% filter(analysis == "PH")
   gg_ph <- plot_site_as_x(data = pH %>% filter(name == "pH"), YLAB = "pH", TITLE = "pH")
-  gg_sp_conduc <- plot_site_as_x(data = pH %>% filter(grepl("specific_conductance", name)), 
+  gg_sp_conduc <- plot_site_as_x(data = pH %>% filter(grepl("spConductance_mscm", name)), 
                                      YLAB = "Specific Conductance μS/cm", TITLE = "Specific Conductance")
   
   tctnts <- combined_surface %>% filter(analysis == "TCTNTS")

### 2023-03-29 _targets.R
update targets for data_combined. export surface vs. all horizons. summary tables. colorsites.

@@ -99,35 +99,35 @@ list(
   ## tar_target(gg_ions, plot_ions(ions_processed, sample_key)),
   
   # combined data
-  tar_target(data_combined, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
+  tar_target(data_combined_all_horizons, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
                                          weoc_processed, dic_processed, din_processed, icp_processed,
                                          ferrozine_processed, mehlich_processed, ions_processed_meq, 
-                                         sample_key)$data_combined),
-  tar_target(data_combined_wide, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
-                                         weoc_processed, dic_processed, din_processed, icp_processed,
-                                         ferrozine_processed, mehlich_processed, ions_processed_meq, 
-                                         sample_key)$data_combined_wide),
+                                         sample_key)),
+  tar_target(data_combined, subset_surface_horizons(data_combined_all_horizons)),
   tar_target(analysis_completion_matrix, compute_analysis_matrix(data_combined)),
-  tar_target(gg_pca_all, compute_overall_pca(data_combined_wide, sample_key)),
-  tar_target(gg_correlations, compute_correlations(data_combined_wide, sample_key)),
+  tar_target(gg_pca_all, compute_overall_pca(data_combined, sample_key)),
+  tar_target(gg_correlations, compute_correlations(data_combined, sample_key)),
   
-  
+  tar_target(gg_by_transect_colorsites, make_graphs_by_transect_SITE_AS_COLOR(data_combined)),
   tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),
   tar_target(gg_by_site_oa, make_graphs_by_site(data_combined %>% filter(horizon != "B"))),
   tar_target(gg_by_site_oab, make_graphs_by_site(data_combined)),
   tar_target(gg_xrd, plot_xrd(xrd_processed)),
   
+  tar_target(summary_tables, make_summary_tables(data_combined)),
+  
   # export
   tar_target(export, {
-    write.csv(data_combined, "1-data/processed/chemistry_combined.csv", row.names = FALSE)
+    write.csv(data_combined, "1-data/processed/chemistry_combined_surface_horizon.csv", row.names = FALSE)
[…]

### 2023-03-29 3-reports/characterization_report.Rmd
report: figures for manuscript

@@
 ## EACH ANALYSIS
+
+### figures for manuscript
+
+<details>
+  <summary>click to open</summary>
+
+```{r ms_mehlich, fig.height=5, fig.width=6}
+tar_read(gg_by_transect_colorsites)$gg_wle_mehlich
+```
+
+```{r ms_wle_icp, fig.height=9, fig.width=9}
+tar_read(gg_by_transect_colorsites)$gg_wle_icp
+```
+
+```{r ms_cb_icp, fig.height=9, fig.width=9}
+tar_read(gg_by_transect_colorsites)$gg_cb_icp
+```
+
+```{r ms_wle_anions, fig.height=5, fig.width=9}
+tar_read(gg_by_transect_colorsites)$gg_wle_anions
+```
+
+```{r ms_cb_anions, fig.height=5, fig.width=9}
+tar_read(gg_by_transect_colorsites)$gg_cb_anions
+```
+
+```{r ms_cb_fe, fig.height=5, fig.width=6}
+tar_read(gg_by_transect_colorsites)$gg_cb_fe
+```
+
+```{r ms_cb_cond, fig.height=5, fig.width=6}
+tar_read(gg_by_transect_colorsites)$gg_cb_spcond
+```
+
+</details>
 

### 2023-10-11 2-code/b-functions_analysis.R
xrd processing and analysis

@@ -844,6 +844,28 @@ make_graphs_by_site = function(data_combined){
 
 plot_xrd = function(xrd_processed){
   
+##  In general the pattern quality was much better than the Exchange samples. 
+##  The agreement index was generally less than the 10% acceptable value. I checked the few that were over this and they had a little preferred orientation.
+##  
+##  I am refining my technique for these analyses, which I also used for the Exchange samples. 
+##  I start by refining the expected basal peak positions for the clay minerals since if I don’t detect those I am not happy saying there is some of that clay mineral there. 
+##  If the basal peak is below a threshold value I exclude that clay from the refinement. 
+##  If I don’t do this, I run the risk of the program using intensity in the higher angle peaks 
+##  (e.g. in areas where preferred orientation needs more calculated intensity) to suggest the presence of clay minerals when there isn’t a basal spacing. 
+##  For the other minerals, I use an percentage cut-off (typically around 2%) to eliminate weak quantities of minerals that cannot be justified by the measured pattern.
+##  
+##  What is new in the COMPASS analysis is a calculation of a “crystallinity” value to give you an indication of amorphous content. 
+##  I noticed that some samples had a very large amorphous background, but because the results are scaled to 100 % it suggests they are 100% quartz, for example. 
+##  The crystallinity index is simply the refined scale factor for quartz divided by the refined weight fraction of quartz and the actual value is arbitrary but comparisons are useful. 
+##  Since these were all measured using the same experimental conditions, the scale factor 
+##  (a number used to scale the theoretical intensity to the measured counts) and weight fraction should strongly correlate when the sample is 100% crystalline. 
+##  
+##  If you look at the graph in the excel sheet for this parameter, you will see the samples fall into 3 main groups:
+##  Where the crystallinity number is around 0.04, corresponding to samples with around 100 % crystallinity (I assume, because most of the samples are here).
+##  Crystallinity numbers larger than 0.04, corresponding to samples with quartz preferred orientation.
[…]

### 2023-10-11 2-code/a-functions_processing.R
remove cations and unnecessary ions from IC. keep only sulfate and chloride

@@ -770,7 +770,8 @@ process_ions = function(ions_data, analysis_key, sample_key, moisture_processed,
     pivot_wider(names_from = "ion") %>% 
     mutate(analysis = "IC") %>% 
     mutate_all(as.character) %>% 
-    mutate_at(vars(ends_with(c("ppm", "ug_g"))), as.numeric)
+    mutate_at(vars(ends_with(c("ppm", "ug_g"))), as.numeric) %>% 
+    dplyr::select(sample_label, analysis, starts_with(c("Chloride", "Sulfate")))
   
   # convert to meq
   charges = 

### 2023-10-11 2-code/b-functions_analysis.R

@@ -699,18 +699,18 @@ make_graphs_by_transect = function(data_combined){
  #                                       SUBTITLE = "extracted with 0.5M HCl, measured colorimetrically using ferrozine")+ylim (0, 5)
   
   ions <- combined_surface %>% filter(analysis == "IC")
-  gg_ions_ca <- plot_transect_as_x(data = ions %>% filter(name == "Calcium_meq100g"), YLAB = "Ca, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water")
-  gg_ions_na <- plot_transect_as_x(data = ions %>% filter(name == "Sodium_meq100g"), YLAB = "Na, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water",
-                                   SCALES = "free")
-  gg_ions_mg <- plot_transect_as_x(data = ions %>% filter(name == "Magnesium_meq100g"), YLAB = "Mg, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water",
-                                   SCALES = "free")
-  gg_ions_k <- plot_transect_as_x(data = ions %>% filter(name == "Potassium_meq100g"), YLAB = "K, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water")
+##  gg_ions_ca <- plot_transect_as_x(data = ions %>% filter(name == "Calcium_meq100g"), YLAB = "Ca, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water")
+##  gg_ions_na <- plot_transect_as_x(data = ions %>% filter(name == "Sodium_meq100g"), YLAB = "Na, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water",
+##                                   SCALES = "free")
+##  gg_ions_mg <- plot_transect_as_x(data = ions %>% filter(name == "Magnesium_meq100g"), YLAB = "Mg, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water",
+##                                   SCALES = "free")
+##  gg_ions_k <- plot_transect_as_x(data = ions %>% filter(name == "Potassium_meq100g"), YLAB = "K, μg/g", TITLE = "IC", SUBTITLE = "extracted with MQ water")
[…]

### 2023-10-11 3-reports/characterization_report.Rmd
remove icr from targets

@@
 
-## FT-ICR-MS
-
-<details>
-  <summary>click to open</summary>
-
-```{r, icr-vk-domains, fig.height=5, fig.width=5}
-tar_read(gg_icr_vankrevelen)$vk_domains
-```
-
-```{r, icr-vk-wle, fig.height=5, fig.width=10}
-tar_read(gg_icr_vankrevelen)$vk_wle
-
-```
-
-```{r, icr-vk-cb, fig.height=10, fig.width=10}
-tar_read(gg_icr_vankrevelen)$vk_cb
-
-```
-
-```{r, icr-unique-wle, fig.height=10, fig.width=4}
-tar_read(gg_icr_unique)$vk_unique_wle
-```
-
-```{r, icr-unique-cb, fig.height=10, fig.width=7}
-tar_read(gg_icr_unique)$vk_unique_cb
-```
-
-```{r, icr-pca, fig.height=5, fig.width=5}
-tar_read(gg_icr_pca)$biplot_all
-```
-
-```{r, icr-pca-regions, fig.height=5, fig.width=10}
-tar_read(gg_icr_pca)$biplot_regions
-```
-
-</details>
-
 ---

### 2023-10-11 _targets.R

@@ -64,11 +64,11 @@ list(
   tar_target(mehlich_map, import_mehlich(FILEPATH = "1-data/phosphorus-mehlich")$mehlich_map),
   tar_target(mehlich_data, import_mehlich(FILEPATH = "1-data/phosphorus-mehlich")$mehlich_data),
   tar_target(mehlich_processed, process_mehlich(mehlich_map, mehlich_data, moisture_processed, subsampling)),
-  tar_target(icr_report, import_fticr_data(FILEPATH = "1-data/icr")),
-  tar_target(icr_meta, make_fticr_meta(icr_report)$meta2),
-  tar_target(icr_data_long, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_blank_corrected),
-  tar_target(icr_data_trt, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_trt),
-  tar_target(icr_relabund_samples, compute_icr_relabund(icr_data_long, icr_meta)),
+##  tar_target(icr_report, import_fticr_data(FILEPATH = "1-data/icr")),
+##  tar_target(icr_meta, make_fticr_meta(icr_report)$meta2),
+##  tar_target(icr_data_long, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_blank_corrected),
+##  tar_target(icr_data_trt, make_fticr_data(icr_report, analysis_key, sample_key)$data_long_trt),
+##  tar_target(icr_relabund_samples, compute_icr_relabund(icr_data_long, icr_meta)),
   tar_target(ions_data, import_ions(FILEPATH = "1-data/ions")),
   tar_target(ions_processed, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)$samples3),
   tar_target(ions_processed_meq, process_ions(ions_data, analysis_key, sample_key, moisture_processed, subsampling)$samples_meq),
@@ -93,9 +93,9 @@ list(
   ## tar_target(gg_icp, plot_icp(icp_processed, sample_key)),
   ## tar_target(gg_ferrozine, plot_ferrozine(ferrozine_processed, sample_key)),
   ## tar_target(gg_mehlich, plot_mehlich(mehlich_processed, sample_key)),
-   tar_target(gg_icr_vankrevelen, plot_vankrevelen(icr_data_trt, icr_meta)),
-   tar_target(gg_icr_unique, plot_vankrevelen_unique(icr_data_trt, icr_meta)),
[…]

### 2023-10-11 _targets.R
fix OAB

@@ -111,7 +111,7 @@ list(
   tar_target(gg_by_transect_colorsites, make_graphs_by_transect_SITE_AS_COLOR(data_combined)),
   tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),
   tar_target(gg_by_site_oa, make_graphs_by_site(data_combined %>% filter(horizon != "B"))),
-  tar_target(gg_by_site_oab, make_graphs_by_site(data_combined)),
+  tar_target(gg_by_site_oab, make_graphs_by_site(data_combined_all_horizons)),
   tar_target(gg_xrd, plot_xrd(xrd_processed)),
   
   tar_target(summary_tables, make_summary_tables(data_combined)),

### 2023-10-11 3-reports/playing_with_pcas.Rmd
playing with pcas

---
title: "Playing with PCAs"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
source("2-code/0-packages.R")
library(ggbiplot)

pal_transect = rev(soilpalettes::soil_palette("redox2", 4))
theme_set(theme_kp())
```

```{r}
targets::tar_load_everything()
```

```{r}
make_data_subset = function(data_combined){

 # data_combined_subset = 
    data_combined %>% 
    filter(!transect %in% "wte") %>% 
    #filter(!grepl("Bromide|Fluoride|Nitrate|Calcium|Magnesium|Potassium|Sodium|Phosphate|Ammonia", name, ignore.case = TRUE)) %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

  #data_combined_wide = 
    data_combined_subset %>% 
    dplyr::select(sample_label, analysis, name, value) %>% 
    separate(name, sep = "_", into = "variable", remove = F) %>% 
    #mutate(name = paste0(variable, " (", analysis, ")")) %>% 
    dplyr::select(sample_label, variable, value) %>% 
    pivot_wider(names_from = "variable") %>% 
    left_join(sample_key) %>% 
    filter(!grepl("016", sample_label)) %>% # this one sample is very weird
    force()

}

```

```{r}

data_combined_subset = make_data_subset(data_combined) %>% filter(!analysis %in% "LOI")
data_combined_wide = make_data_wide(data_combined_subset, sample_key)
data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
                                             filter(!grepl("dic", name, ignore.case = TRUE)),
                                           sample_key)

fit_pca_function = function(dat){

  dat %>% 
    drop_na()

  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    dplyr::mutate(row = row_number()) %>% 
    drop_na()

  num_row_numbers = num %>% dplyr::select(row)

  grp = 
    dat %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

  num = num %>% dplyr::select(-row)
  pca_int = prcomp(num, scale. = T)

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}
```

---

## OVERALL

```{r}
## PCA input files ----
pca_overall = fit_pca_function(data_combined_wide_NO_DIC) 
pca_overall_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "CB"))
```

```{r}
## PCA plots overall ----
#gg_pca_overall = 
  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(#shape = pca_overall$grp$transect,
               color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL
```

### regions separated - colored by transect

```{r, fig.width=10, fig.height=6}
gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

#  library(patchwork)
#  gg_pca_regions = gg_pca_wle + gg_pca_cb
gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

### regions separated - colored by site

```{r, fig.width=10, fig.height=6}
gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$site), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$transect,
                 color = groups))+ 
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$site), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$transect,
                 color = groups))+ 
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

#  library(patchwork)
#  gg_pca_regions = gg_pca_wle + gg_pca_cb
gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

---

## BY TRANSECT

```{r}
# BY TRANSECT ----

pca_overall_upland = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland"))
pca_overall_transition = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition"))
pca_overall_wetland = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc"))

gg_overall_upland = 
  ggbiplot(pca_overall_upland$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_upland$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_upland$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA upland, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_transition = 
  ggbiplot(pca_overall_transition$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_transition$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_transition$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA transition, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_wetland = 
  ggbiplot(pca_overall_wetland$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wetland$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wetland$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA wetland, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_upland
gg_overall_transition
gg_overall_wetland

#

```

---

## BY TRANSECT + REGION

```{r, fig.width=10, fig.height=6}
# BY TRANSECT + REGION ----

pca_upland_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland" & region == "WLE"))
pca_upland_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland" & region == "CB"))

gg_upland_wle = 
  ggbiplot(pca_upland_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_upland_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA upland, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_upland_cb = 
  ggbiplot(pca_upland_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_upland_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA upland, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_upland_wle , gg_upland_cb)

```

```{r, fig.width=10, fig.height=6}

pca_transition_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition" & region == "WLE"))
pca_transition_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition" & region == "CB"))

gg_transition_wle = 
  ggbiplot(pca_transition_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_transition_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA transition, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_transition_cb = 
  ggbiplot(pca_transition_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_transition_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA transition, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_transition_wle , gg_transition_cb)

```

```{r, fig.width=10, fig.height=6}

pca_wetland_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc" & region == "WLE"))
pca_wetland_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc" & region == "CB"))

gg_wetland_wle = 
  ggbiplot(pca_wetland_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_wetland_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA wetland, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_wetland_cb = 
  ggbiplot(pca_wetland_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_wetland_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA wetland, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_wetland_wle , gg_wetland_cb)

```

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>

### 2023-12-06 3-reports/playing_with_pcas.Rmd
clean up PCA, remove GWC and DIC

@@
 
-```{r}
+```{r, eval=FALSE}
 
@@
                                            sample_key)
+```
 
+```{r}
+## Set df for PCA
+
+data_wide_PCA = data_combined_wide %>% dplyr::select(-c("dic", "gwc"))
+
+```
+
+```{r}
 fit_pca_function = function(dat){
@@
 }
+
 ```
@@
 ## PCA input files ----
-pca_overall = fit_pca_function(data_combined_wide_NO_DIC) 
-pca_overall_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "WLE"))
-pca_overall_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "CB"))
+pca_overall = fit_pca_function(data_wide_PCA) 
+pca_overall_wle = fit_pca_function(data_wide_PCA %>% filter(region == "WLE"))
+pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
 ```
@@
   NULL
+
+  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall$grp$region,
+               color = pca_overall$grp$transect))+ 
+ # scale_color_manual(values = c("#16879C", "#BB281E" ))+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  # xlim(-4,20)+
+  # ylim(-8,8)+
+  labs(shape="",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
 ```

### 2023-12-06 3-reports/playing_with_pcas.Rmd
fix PCA nomenclature. plot PC1

@@
 data_combined_subset = make_data_subset(data_combined) %>% filter(!analysis %in% "LOI")
-data_combined_wide = make_data_wide(data_combined_subset, sample_key)
 data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
@@
 
-data_wide_PCA = data_combined_wide %>% dplyr::select(-c("dic", "gwc"))
+data_combined_wide = make_data_wide(data_combined, sample_key)
+data_wide_PCA = 
+  data_combined_wide %>% 
+  dplyr::select(-c("dic", "gwc", "percentOM")) %>% 
+  rename(CEC = cec,
+         NH4N = nh4n,
+         NO3N = no3n,
+         P = mehlichp,
+         WEOC = npoc)
 
@@
 
-  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
-           groups = as.character(pca_overall$grp$transect), 
-           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
-  geom_point(size=3,stroke=1, alpha = 1,
-             aes(shape = pca_overall$grp$region,
-               color = pca_overall$grp$transect))+ 
- # scale_color_manual(values = c("#16879C", "#BB281E" ))+
-  #scale_shape_manual(values = c(21, 19))+
-  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
-  # xlim(-4,20)+
-  # ylim(-8,8)+
-  labs(shape="",
-       title = "Overall PCA, both regions",
-       subtitle = "Surface horizons only")+
-  theme_kp()+
-  NULL
+#  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+#           groups = as.character(pca_overall$grp$transect), 
+#           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+#  geom_point(size=3,stroke=1, alpha = 1,
+#             aes(shape = pca_overall$grp$region,
+#               color = pca_overall$grp$transect))+ 
+# # scale_color_manual(values = c("#16879C", "#BB281E" ))+
+#  #scale_shape_manual(values = c(21, 19))+
+#  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+#  # xlim(-4,20)+
+#  # ylim(-8,8)+
+#  labs(shape="",
+#       title = "Overall PCA, both regions",
+#       subtitle = "Surface horizons only")+
+#  theme_kp()+
+#  NULL
 
@@
 
-pca_overall_upland = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland"))
-pca_overall_transition = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition"))
-pca_overall_wetland = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc"))
+pca_overall_upland = fit_pca_function(data_wide_PCA %>% filter(transect == "upland"))
+pca_overall_transition = fit_pca_function(data_wide_PCA %>% filter(transect == "transition"))
+pca_overall_wetland = fit_pca_function(data_wide_PCA %>% filter(transect == "wc"))
 
@@
 
-pca_upland_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland" & region == "WLE"))
-pca_upland_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "upland" & region == "CB"))
+pca_upland_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "upland" & region == "WLE"))
+pca_upland_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "upland" & region == "CB"))
 
@@
 
-pca_transition_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition" & region == "WLE"))
-pca_transition_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "transition" & region == "CB"))
+pca_transition_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "transition" & region == "WLE"))
+pca_transition_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "transition" & region == "CB"))
 
@@
 
-pca_wetland_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc" & region == "WLE"))
-pca_wetland_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(transect == "wc" & region == "CB"))
+pca_wetland_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "wc" & region == "WLE"))
+pca_wetland_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "wc" & region == "CB"))
 
@@
 
+## Plotting PC1 (drivers of variability)
+
+```{r}
+get_pc_values = function(dat){
+
+  pca = fit_pca_function(dat)
+
+  x = pca$pca_int$rotation
+  x2 = 
+    x %>% 
+    as.data.frame() %>% 
+    dplyr::select(PC1) %>% 
+    arrange(desc(PC1)) %>% 
+    rownames_to_column("variable") %>% 
+    mutate(variable = fct_reorder(variable, (PC1)))
+
+  x2
+}
+```
+
+```{r, fig.height=8, fig.width=6}
+
+pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
+pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
+pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")
+
+pc_all = 
+  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
+  mutate(variable = fct_reorder(variable, (PC1)))
+
+pc_all %>% 
+  mutate(variable = fct_reorder(variable, (PC1))) %>% 
+  filter(region != "combined") %>% 
+  ggplot(aes(x = PC1, y = variable, color = region))+
+  geom_point(size = 5)+
+  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
+pc_all %>% 
+  mutate(variable = fct_reorder(variable, (PC1))) %>% 
+  filter(region != "combined") %>% 
+  ggplot(aes(x = abs(PC1), y = variable, color = region))+
+  geom_point(size = 5)+
+  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
+ # xlim(0, 0.4)+
+  #geom_vline(xintercept = 0)+
+  labs(y = "",
+       x = "|PC1|")+
+  theme_kp()
+
+```
+
+---
+
 ## Session Info

### 2023-12-06 2-code/b2-functions_clustering.R
clustering code

@@ -1,0 +1,194 @@
+
+
+# hierarchical clusters ----
+
+library(cluster)
+library(factoextra)
+
+data(iris)
+scale(iris)
+
+df <- USArrests 
+
+df <- na.omit(df)
+df <- scale(df)
+
+
+
+
+df = iris %>%  dplyr::select(where(is.numeric))
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 5, border = 2:5)
+
+fviz_cluster(list(data = df, cluster = sub_grp))
+
+
+df = data_combined_wide %>% filter(region == "CB") %>%  dplyr::select(where(is.numeric))
+grp = data_combined_wide %>%  filter(region == "CB") %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+# plot
+set.seed(12345)
+gg = 
+  df_cl %>% 
+  ggplot(aes(y = transect, x = as.character(cluster), color = site))+
+  geom_point(aes(group = site, text = sample_label), 
+             size = 4, 
+             position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0.1, 
+                                             dodge.width = 0.5))
+
+plotly::ggplotly(gg)
+
+
+# pca
+library(ggbiplot)
+
+pca_wle_clusters = fit_pca_function(df_cl %>% dplyr::select(-dic)) 
+
+ggbiplot(pca_wle_clusters$pca_int, obs.scale = 1, var.scale = 1,
[…]

### 2024-01-11 3-reports/playing_with_pcas.Rmd
remove wte from pca. plot PCA by transect

@@
 
-data_combined_subset = make_data_subset(data_combined) %>% filter(!analysis %in% "LOI")
+data_combined_subset = make_data_subset(data_combined) # %>% filter(!analysis %in% "LOI")
 data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
@@
 
-data_combined_wide = make_data_wide(data_combined, sample_key)
+data_combined_wide = make_data_wide(data_combined_subset, sample_key)
 data_wide_PCA = 
@@
 
-#  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
-#           groups = as.character(pca_overall$grp$transect), 
-#           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
-#  geom_point(size=3,stroke=1, alpha = 1,
-#             aes(shape = pca_overall$grp$region,
-#               color = pca_overall$grp$transect))+ 
-# # scale_color_manual(values = c("#16879C", "#BB281E" ))+
-#  #scale_shape_manual(values = c(21, 19))+
-#  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
-#  # xlim(-4,20)+
-#  # ylim(-8,8)+
-#  labs(shape="",
-#       title = "Overall PCA, both regions",
-#       subtitle = "Surface horizons only")+
-#  theme_kp()+
-#  NULL
+  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall$grp$region,
+               color = pca_overall$grp$transect))+ 
+ # scale_color_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  # xlim(-4,20)+
+  # ylim(-8,8)+
+  labs(shape="",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
 

### 2024-01-11 3-reports/playing_with_pcas.Rmd
PC1 loadings by transect and region

@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 
@@
 
+PC1 loadings by transect
+```{r}
+
+pc_transect = 
+  data_wide_PCA %>% 
+  group_by(transect) %>% 
+  do(get_pc_values(.)) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wc")))
+
+pc_transect %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5)+
+  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
+ # xlim(0, 0.4)+
+  #geom_vline(xintercept = 0)+
+  labs(y = "",
+       x = "|PC1|")+
+  facet_grid(. ~ transect)+
+  theme_kp()
+
+```
+
+PC1 loadings by region+transect
+```{r}
+
+pc_region_transect = 
+  data_wide_PCA %>% 
+  group_by(region, transect) %>% 
+  do(get_pc_values(.)) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wc")))
+
+pc_region_transect %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable, color = region))+
+  geom_point(size = 5)+
+  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
+ # xlim(0, 0.4)+
+  #geom_vline(xintercept = 0)+
+  labs(y = "",
+       x = "|PC1|")+
+  facet_grid(. ~ transect)+
+  theme_kp()
+
+```
+
 ---

### 2024-03-15 3-reports/playing_with_pcas.Rmd
cleaning

@@
 pca_overall = fit_pca_function(data_wide_PCA) 
-pca_overall_wle = fit_pca_function(data_wide_PCA %>% filter(region == "WLE"))
+pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
 pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
@@
            groups = as.character(pca_overall_wle$grp$transect), 
-           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
   geom_point(size=3,stroke=1, alpha = 1,
@@
 
+---
+
+---
+
 ## Session Info

### 2024-03-15 3-reports/playing_with_pcas.Rmd
pc_region_transect

@@
 
+pc_region_transect %>% 
+  filter(transect != "transition") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable, color = transect))+
+  geom_point(size = 5)+
+  scale_colour_manual(values = c("red", "blue"))+
+ # xlim(0, 0.4)+
+  #geom_vline(xintercept = 0)+
+  labs(y = "",
+       x = "|PC1|")+
+  facet_grid(. ~ region)+
+  theme_kp()+
+  geom_vline(xintercept = 0.3, linetype = "dashed")+
+  geom_vline(xintercept = 0.1, linetype = "dashed")
+
 ```

### 2024-03-15 3-reports/playing_with_pcas.Rmd
CV

@@
 
+## CV
+
+```{r}
+cv = 
+  data_combined %>% 
+  group_by(region, site, transect, horizon, analysis, name) %>% 
+  dplyr::summarise(cv = 100 * sd(value)/mean(value),
+                   n = n(),
+                   mean = mean(value)) %>% 
+  filter(!transect %in% "wte")
+```
+
+```{r}
+(cv %>% 
+   filter(!(name == "Al_meq100g" & region == "WLE")) %>% 
+  ggplot(aes(x = transect, y = cv))+
+  geom_violin(scale = "area", fill = "grey90")+
+  geom_jitter(width = 0.1, size = 3, aes(color = site, text = name))+
+   geom_boxplot(width = 0.05, outlier.color = NA)+
+  facet_wrap(~region)+
+   theme_kp())  %>% plotly::ggplotly()
+
+cv %>% 
+  ggplot(aes(x = region, y = cv))+
+  geom_point(width = 0.1, aes(color = transect), 
+             position = position_dodge(width = 0.2))+
+  expand_limits(y = 0)+
+  facet_wrap(~ name, scales = "free_y")
+
+View(data_combined_wide %>% dplyr::select(region, site, transect, horizon, contains(c("dic", "pH"))))
+```
+
 ---

### 2024-03-15 3-reports/playing_with_pcas.Rmd
clusters

@@
 
+## CLUSTERS
+
+```{r}
+
+library(cluster)
+library(factoextra)
+
+df = data_wide_PCA %>% filter(region == "CB") %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA %>% filter(region == "CB")%>% drop_na() %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+```
+
+```{r}
+library(ggConvexHull)
+
+pca_cluster = fit_pca_function(df_cl)
+
+ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
+                   groups = as.character(pca_cluster$grp$cluster), 
+                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_convexhull(aes(group = pca_cluster$grp$cluster,
+                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_cluster$grp$site,
+                 color = groups))+ 
+  # scale_color_manual(values = c("#16879C", "#BB281E" ))+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  # xlim(-4,20)+
+  # ylim(-8,8)+
+  labs(fill="cluster",
+       color = "cluster",
+       shape = "region",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+```
+
+```{r}
+df_cl %>%
+ # distinct(region, site, transect, cluster) %>% 
+  ggplot(aes(y = cluster, x = site))+
+  geom_point(aes(color = transect, shape = transect),
+             size = 4, stroke = 1,
+             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
+                                             dodge.width = 0.4))
+
+devtools::install_github("liamgilbey/ggwaffle")
+library(ggwaffle)
+
+df_cl %>% 
+  ggplot()+
+  geom_waffle(aes(x = transect, y = site, fill = cluster))
+
+```
+
 ---

### 2024-03-15 3-reports/playing_with_pcas.Rmd
do_pca function (for grouping)
pca by site

@@
 
+## BY SITE
+```{r}
+
+pca_owc = fit_pca_function(data_wide_PCA %>% filter(site == "OWC")) 
+pca_ptr = fit_pca_function(data_wide_PCA %>% filter(site == "PTR")) 
+pca_crc = fit_pca_function(data_wide_PCA %>% filter(site == "CRC")) 
+pca_gcw = fit_pca_function(data_wide_PCA %>% filter(site == "GCW")) 
+pca_msm = fit_pca_function(data_wide_PCA %>% filter(site == "MSM")) 
+pca_gwi = fit_pca_function(data_wide_PCA %>% filter(site == "GWI"))
+
+gg_pca_owc = ggbiplot(pca_owc$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_owc$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA OWC", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+gg_pca_ptr = ggbiplot(pca_ptr$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_ptr$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA PTR", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+gg_pca_crc = ggbiplot(pca_crc$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_crc$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA CRC", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+gg_pca_gcw = ggbiplot(pca_gcw$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_gcw$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA GCW", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+gg_pca_msm = ggbiplot(pca_msm$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_msm$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA MSM", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+gg_pca_gwi = ggbiplot(pca_gwi$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_gwi$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(color = groups))+ 
+  labs(shape="", title = "PCA GWI", subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+library(patchwork)
+
+gg_pca_gcw + gg_pca_msm + gg_pca_gwi +
+  plot_layout(guides = "collect")&
+  theme(legend.position = "bottom")
+
+gg_pca_owc + gg_pca_ptr + gg_pca_crc +
+  plot_layout(guides = "collect")&
+  theme(legend.position = "bottom")
+
+```
+
 ---
@@
 }
+```
+
+```{r}
+do_pca = function(dat){
+
+  pca = fit_pca_function(dat)
+
+  x = pca$pca_int$rotation %>% as.data.frame() %>% rownames_to_column("variable")
+
+}
+
+x = data_wide_PCA %>% group_by(region, transect) %>% do(do_pca(.))
+
+x %>% 
+  ggplot(aes(x = PC1, y = PC2, label = variable))+
+  geom_point()+geom_text()+
+  facet_grid(transect ~ region)
+
 ```

### 2024-03-15 2-code/b-functions_analysis.R
ternary plot for texture

@@ -1231,7 +1231,8 @@ plot_wrc = function(wrc_processed){
 plot_texture = function(texture_processed, sample_key){
 
   texture = 
-    texture_processed %>% 
+#    texture_processed %>% 
+    soil_texture %>% 
     left_join(sample_key)
   
   # plot the soil texture triangle
@@ -1239,7 +1240,7 @@ plot_texture = function(texture_processed, sample_key){
   data(USDA)
   
   gg_texture = 
-    ggtern(data = texture,
+    ggtern(data = texture %>% filter(region == "CB"),
            aes(x = percent_sand, y = percent_clay, z = percent_silt)) +
     geom_polygon(data = USDA, 
                  aes(x = Sand, y = Clay, z = Silt, group = Label),
@@ -1250,8 +1251,26 @@ plot_texture = function(texture_processed, sample_key){
     theme_showarrows()+
     theme_hidetitles()+
     theme_clockwise() +
-    #facet_wrap(~ site)+
+    facet_wrap(~ site + horizon)+
     NULL
+  
+  ggtern(data = texture %>% filter(region == "CB" & site == "MSM"),
+         aes(x = percent_sand, y = percent_clay, z = percent_silt)) +
+    geom_polygon(data = USDA, 
+                 aes(x = Sand, y = Clay, z = Silt, group = Label),
+                 fill = NA, size = 0.3, alpha = 0.5, color = "grey30")+
+    geom_point(aes(color = transect, shape = horizon),
+               size = 5, stroke = 1)+
+ #   scale_shape_manual(values = c(21, 22, 23))+
+    theme_bw()+
+    theme_showarrows()+
+    theme_hidetitles()+
+    theme_clockwise() +
+    facet_wrap(~ horizon)+
+    NULL+
+    geom_text(data = USDA_text,
+              aes(x = Sand, y = Clay, z = Silt, label = Label),
+              size = 3, color = "grey30")
   
   USDA_text <- 
     USDA  %>% 

### 2024-03-15 2-code/scratchpad.R
scratchpad

@@ -762,3 +762,480 @@ icp_ic_long =
   drop_na() %>% 
   left_join(sample_key)
 icp_ic_long %>% write.csv("1-data/processed/synoptic_ICP_IC_all_horizons.csv", row.names = F)
+
+
+
+# -------------------------------------------------------------------------
+
+theme_set(theme_bw(base_size = 14))
+
+
+x = pca_overall$pca_int$rotation
+
+x2 = 
+  x %>% 
+  as.data.frame() %>% 
+  dplyr::select(PC1) %>% 
+  arrange(desc(PC1)) %>% 
+  rownames_to_column("variable") %>% 
+  mutate(variable = fct_reorder(variable, (PC1)))
+
+x2 %>% 
+  ggplot(aes(x = PC1, y = variable))+
+  geom_point(size = 3)
+
+
+
+pca_overall = fit_pca_function(data_combined_wide_NO_DIC) 
+
+data_wide = 
+  data_combined_wide_NO_DIC %>% 
+  dplyr::select(-c(gwc, Mg, Na))
+
+
+
+
+
+
+
+get_pc_values = function(dat){
+  
+  pca = fit_pca_function(dat)
+  
+  x = pca$pca_int$rotation
+  x2 = 
+    x %>% 
+    as.data.frame() %>% 
+    dplyr::select(PC1) %>% 
+    arrange(desc(PC1)) %>% 
+    rownames_to_column("variable") %>% 
+    mutate(variable = fct_reorder(variable, (PC1)))
+  
+  x2
+}
+
+# round 1
+pc_overall = get_pc_values(pca_overall) %>% mutate(region = "combined")
+pc_wle = get_pc_values(pca_overall_wle) %>% mutate(region = "WLE")
+pc_cb = get_pc_values(pca_overall_cb) %>% mutate(region = "CB")
+
+pc_all = 
+  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
+  mutate(variable = fct_reorder(variable, (PC1)))
+
+pc_all %>% 
+  filter(region == "combined") %>% 
+  mutate(variable = fct_reorder(variable, (PC1))) %>% 
+  ggplot(aes(x = PC1, y = variable, color = region))+
+  geom_point(size = 3)+
+  xlim(-0.4, 0.4)
+
+
+# round 2
+pc_overall = get_pc_values(data_wide) %>% mutate(region = "combined")
+pc_wle = get_pc_values(data_wide %>% filter(region == "WLE")) %>% mutate(region = "WLE")
[…]

### 2024-04-05 3-reports/playing_with_pcas.Rmd
cv and distance

@@
 
+```{r}
+wine.pca <- prcomp(wine[treatments], scale = TRUE)
+
+df.wine.x$groups <- wine.class
+wine$groups <- wine.class
+feature.centroids <- aggregate(wine[treatments], list(Type = wine$groups), mean)
+
+data(wine)
+treatments <- 1:2 #treatments to be considerd for this calculation
+wine.pca <- prcomp(wine[treatments], scale. = TRUE)
+#calculate the centroids for the feature/treatment space and the pca space
+df.wine.x <- as.data.frame(wine.pca$x)
+df.wine.x$groups <- wine.class
+wine$groups <- wine.class
+feature.centroids <- aggregate(wine[treatments], list(Type = wine$groups), mean)
+
+dist(rbind(feature.centroids[feature.centroids$Type == "barolo",][-1],
+           feature.centroids[feature.centroids$Type == "grignolino",][-1]), method = "euclidean")
+
+data.pca = prcomp(data_wide_PCA %>% drop_na() %>% dplyr::select(where(is.numeric)))
+data.feature.centroids <- aggregate(data_wide_PCA %>% drop_na() %>% dplyr::select(where(is.numeric)), 
+                                    list(Type = (data_wide_PCA %>% drop_na())$region), 
+                                    mean, na.action = na.pass)
+
+dist(rbind(data.feature.centroids[data.feature.centroids$Type == "CB",][-1],
+           data.feature.centroids[data.feature.centroids$Type == "WLE",][-1]), method = "euclidean")
+
+centroid.pca = fit_pca_function(data_wide_PCA)$pca_int
+centroid.feature = aggregate(data_wide_PCA)
+
+data.feature.centroids = 
+  data_combined_subset %>%
+  filter(site == "GWI") %>% 
+  group_by(transect, name) %>% 
+  dplyr::summarise(mean = mean(value, na.rm = T)) %>% 
+  pivot_wider(names_from = "name", values_from = "mean")
+
+dist(rbind(data.feature.centroids[data.feature.centroids$transect == "transition",][-1],
+           data.feature.centroids[data.feature.centroids$transect == "wetland",][-1]), method = "euclidean")
+
+dist(rbind(data.feature.centroids[data.feature.centroids$transect == "upland",][-1],
+           data.feature.centroids[data.feature.centroids$transect == "wetland",][-1]), method = "euclidean")
+
+dist(rbind(data.feature.centroids[data.feature.centroids$transect == "transition",][-1],
+           data.feature.centroids[data.feature.centroids$transect == "upland",][-1]), method = "euclidean")
+
+df = data(iris)
+dist.iris<-vegdist(iris[,1:4],method = "bray") #calculate Bray-Curtis dissim
+dist.iris
+
+#avg distance to group centroids
+permdisp<-betadisper(dist.iris,iris$Species,type="centroid")
+permdisp
+library(usedist)
+dist_between_centroids(dist.iris,1:50, 51:100) #1:50 = group setosa, 51:100 = group versicolor
+
+library(vegan)
+TukeyHSD(permdisp)
+
+dat.owc = data_wide_PCA %>% filter(site == "OWC") %>% drop_na() 
+dist.owc <- vegdist(dat.owc %>% dplyr::select(where(is.numeric)), 
+                    method = "bray") #calculate Bray-Curtis dissim
+permdisp.owc <- betadisper(dist.owc, dat.owc$transect, type = "centroid")
+TukeyHSD(permdisp.owc)
+
+dat.gcw = data_wide_PCA %>% filter(site == "GCW") %>% drop_na() 
+dist.gcw <- vegdist(dat.gcw %>% dplyr::select(where(is.numeric)), 
+                    method = "bray") #calculate Bray-Curtis dissim
+permdisp.gcw <- betadisper(dist.gcw, dat.gcw$transect, type = "centroid")
+TukeyHSD(permdisp.gcw)
+
+dat.gwi = data_wide_PCA %>% filter(site == "GWI") %>% drop_na() 
+dist.gwi <- vegdist(dat.gwi %>% dplyr::select(where(is.numeric)), 
+                    method = "bray") #calculate Bray-Curtis dissim
+permdisp.gwi <- betadisper(dist.gwi, dat.gcw$transect, type = "centroid")
+TukeyHSD(permdisp.gwi)
+
+```
+
+```{r}
+library(mosaic)
+x1 <- c(1,0,0); x2 <- c(1,2,3); y1 <- c(3,4,5); y2 <- rnorm(3)
+# projection onto the 1 vector gives the mean vector
+mean(y2)            
+mosaic::project(y2, 1)
+# return the length of the vector, rather than the vector itself
+project(y2, 1, type='length')
+project(y1 ~ x1 + x2) -> pr; pr
+
+project(x, y, type = "length")
+
+```
+
 ---
@@
   dplyr::summarise(cv = 100 * sd(value)/mean(value),
+                   sd = sd(value),
                    n = n(),
                    mean = mean(value)) %>% 
-  filter(!transect %in% "wte")
+  filter(!transect %in% "wte") %>% 
+  filter(!(name == "Al_meq100g" & region == "WLE")) %>% 
+  filter(name != "gwc_perc")
+
 ```
@@
   geom_jitter(width = 0.1, size = 3, aes(color = site, text = name))+
-   geom_boxplot(width = 0.05, outlier.color = NA)+
+   geom_boxplot(width = 0.05, outlier.color = NA, fill = "white")+
   facet_wrap(~region)+
@@
   expand_limits(y = 0)+
-  facet_wrap(~ name, scales = "free_y")
+  scale_y_log10()+
+  facet_wrap(~ name, 
+             #scales = "free_y"
+             )
+
+cv %>% 
+  filter(!(name == "Al_meq100g" & region == "WLE")) %>% 
+  ggplot(aes(x = name, y = cv))+
+  geom_violin()+
+  geom_point(width = 0.1, aes(color = region), 
+            # position = position_dodge(width = 0.9)
+             )+
+  expand_limits(y = 0)+
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
+  facet_grid(transect ~ .)+
+#  scale_y_log10()+
+  NULL
+
+cv %>% 
+  filter(!(name == "Al_meq100g" & region == "WLE")) %>% 
+  ggplot(aes(y = name, x = cv))+
+  geom_violin(aes(color = region))+
+  geom_point(width = 0.1, aes(color = region), 
+             position = position_dodge(width = 0.9)
+             )+
+  expand_limits(x = 0)+
+#  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
+  facet_wrap(~transect)+
+#  scale_y_log10()+
+  NULL
+
+cv %>% 
+  ggplot(aes(x = mean, y = cv, color = region))+
+  geom_point()+
+  facet_wrap(~name, scales = "free_x")
 
 View(data_combined_wide %>% dplyr::select(region, site, transect, horizon, contains(c("dic", "pH"))))
+```
+
+```{r}
+library(ggwaffle)
+cv %>% 
+  ggplot(aes(x = transect, y = name, fill = cv))+
+  geom_waffle()+
+  geom_text(aes(label = round(cv,2)), color = "black")+
+  scale_fill_gradientn(trans = "log",
+                       colors = PNWColors::pnw_palette("Bay", 12), breaks = c(0, 10, 20, 40, 100, 200))+
+  facet_wrap(~region + site)
+
+cv[cv == "NaN"] <- NA
+cv = v %>% drop_na()
+
+aov(cv ~ transect + region + name, data = cv) %>% summary()
+
+library(nlme)
+nlme::lme(cv ~ transect, random = ~1|name, data = cv %>% drop_na() %>% filter(region == "WLE")) %>% anova()
+nlme::lme(cv ~ region, random = ~1|transect, data = cv %>% drop_na()) %>% anova()
+
+cv %>% group_by(region, transect) %>% dplyr::summarise(mean_cv = mean(cv, na.rm = T))
+
+a = aov((cv) ~ transect, data = cv %>% filter(region == "CB"))
+agricolae::HSD.test(a, "transect") %>% print()
+
+a = aov((cv) ~ transect, data = cv %>% filter(region == "WLE"))
+agricolae::HSD.test(a, "transect") %>% print()
 ```

### 2024-04-05 2-code/scratchpad.R
scaled distance by analyte

@@ -1253,17 +1253,53 @@ data_scaled =
          upland_scaled = (upland - upland) / (wetland - upland),
          wetland_scaled = (wetland - upland) / (wetland - upland))
 
+
+data_scaled %>%
+  dplyr::select(region, site, horizon, name, ends_with("scaled")) %>% 
+  pivot_longer(cols = c(upland_scaled, transition_scaled, wetland_scaled), names_to = "transect") %>% 
+  ggplot(aes(x = value, y = name, color = transect))+
+  geom_segment(aes(x = 0, xend = value, yend = name), color = "black", linetype = "dashed", linewidth = 0.2)+
+  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
+  geom_point(size = 3)+
+  # xlim(-1, 2.5)+
+  #scale_x_log10()+
+  facet_wrap(~region + site)+
+  theme_bw()+
+  theme(panel.grid = element_blank(),
+        #axis.text.x = element_blank()
+  )
+
+
+
+
+data_long = 
+  data_combined_wide %>% 
+  dplyr::select(-"gwc") %>% 
+  rename(CEC = cec,
+         NH4N = nh4n,
+         NO3N = no3n,
+         P = mehlichp,
+         WEOC = npoc,
+         DIC = dic,
+         SOM = percentOM,
+         Cl = Chloride,
+         SO4 = Sulfate) %>% 
+  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
+  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
+  filter(!is.na(value))
+
+
 data_scaling_max_min = 
-  data_combined_subset %>% 
+  data_long %>% 
   filter(transect != "transition") %>% 
   group_by(region, site, transect, horizon, name) %>% 
   dplyr::summarise(value = mean(value)) %>% 
   group_by(region, site, horizon, name) %>% 
-  dplyr::summarise(max = max(value),
-                   min = min(value))
+  dplyr::summarise(max = max(value, na.rm = T),
+                   min = min(value, na.rm = T))
 
 data_scaled2 = 
-  data_combined_subset %>% 
+  data_long %>% 
   group_by(region, site, transect, horizon, name) %>% 
   dplyr::summarise(value = mean(value)) %>% 
   left_join(data_scaling_max_min) %>% 
@@ -1280,54 +1316,89 @@ data_scaled2 =
[…]

### 2024-07-29 2-code/scratchpad.R
scratchpads

@@ -1093,7 +1093,7 @@ compute_euclidean_distance <- function(data_normalized){
 euc <- 
   data_normalized %>% 
   filter(name != "percentOM") %>% 
-  group_by(region, site, name) %>% 
+  group_by(region, site) %>% 
   do(compute_euclidean_distance(.))
 
 
@@ -1186,7 +1186,7 @@ fticr_eucl =
   geom_hline(yintercept = 1.713, linetype = "dashed")+
   
   theme_kp()
-```
+
 
 
 
@@ -1327,19 +1327,19 @@ data_scaled2 =
                                         )))
   
 
-    # data_scaled2 %>%
-    #   ggplot(aes(x = scaled2, y = name, color = transect))+
-    #   geom_segment(aes(x = 0, xend = scaled2, yend = name), color = "black", linetype = "dashed", linewidth = 0.2)+
-    #   geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
-    #   geom_point(size = 3)+
-    #   #geom_text(aes(x = scaled2, y = name, label = label), color = "black")+
-    #    xlim(-1, 5)+
-    #  # scale_x_log10()+
-    #   facet_wrap(~region + site)+
-    #   theme_bw()+
-    #   theme(panel.grid = element_blank(),
-    #         #axis.text.x = element_blank()
-    #   )
+     data_scaled2 %>%
+       ggplot(aes(x = scaled2, y = name, color = transect))+
+       geom_segment(aes(x = 0, xend = scaled2, yend = name), color = "black", linetype = "dashed", linewidth = 0.2)+
+       geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
+       geom_point(size = 3)+
+       #geom_text(aes(x = scaled2, y = name, label = label), color = "black")+
+        xlim(-1, 5)+
+      # scale_x_log10()+
+       facet_wrap(~region + site)+
+       theme_bw()+
+       theme(panel.grid = element_blank(),
+             #axis.text.x = element_blank()
+       )
 
 gg_cb = 
   data_scaled2 %>%
@@ -1402,3 +1402,191 @@ gg_wle =
     )
 
 cowplot::plot_grid(gg_wle, gg_cb, rel_widths = c(1, 0.85))
+
+
+## stoichiometry ----
+
+cnps = 
+  data_combined_wide %>% 
+  filter(!transect %in% "wte") %>% 
+  mutate(cn = TC_perc/TN_perc,
+         ns = TN_perc/TS_perc,
[…]

### 2024-07-29 2-code/scratchpad2.R

@@ -1,0 +1,345 @@
+## PCA on C-normalized data
+
+```{r}
+data_wide_PCA2 = 
+  data_wide_PCA %>% 
+  pivot_longer(cols = c(where(is.numeric), -TC)) %>% 
+  mutate(value = value/TC) %>% 
+  dplyr::select(-TC) %>% 
+  pivot_wider()
+
+pca_all_c = fit_pca_function(data_wide_PCA2)
+pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
+pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
+
+
+ggbiplot(pca_all_c$pca_int, obs.scale = 1, var.scale = 1,
+         groups = as.character(pca_all_c$grp$region), 
+         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(#shape = pca_all_c$grp$site,
+               color = groups))+ 
+  scale_color_manual(#breaks = c("upland", "transition", "wte", "wc"), 
+    values = pal_transect)+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  xlim(-6,6)+
+  ylim(-4.5,4.5)+
+  labs(shape="",
+       title = "PCA: all, normalized to C",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+
+ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
+         groups = as.character(pca_overall_wle$grp$transect), 
+         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall_wle$grp$site,
+               color = groups))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+    values = pal_transect)+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  xlim(-6,6)+
+  ylim(-4.5,4.5)+
+  labs(shape="",
+       title = "PCA: all, normalized to C",
[…]

### 2024-07-29 3-reports/playing_with_pcas.Rmd
Update playing_with_pcas.Rmd

@@
 
-pca_owc = fit_pca_function(data_wide_PCA %>% filter(site == "OWC")) 
-pca_ptr = fit_pca_function(data_wide_PCA %>% filter(site == "PTR")) 
-pca_crc = fit_pca_function(data_wide_PCA %>% filter(site == "CRC")) 
-pca_gcw = fit_pca_function(data_wide_PCA %>% filter(site == "GCW")) 
-pca_msm = fit_pca_function(data_wide_PCA %>% filter(site == "MSM")) 
-pca_gwi = fit_pca_function(data_wide_PCA %>% filter(site == "GWI"))
+pca_owc = fit_pca_function(data_wide_PCA2 %>% filter(site == "OWC")) 
+pca_ptr = fit_pca_function(data_wide_PCA2 %>% filter(site == "PTR")) 
+pca_crc = fit_pca_function(data_wide_PCA2 %>% filter(site == "CRC")) 
+pca_gcw = fit_pca_function(data_wide_PCA2 %>% filter(site == "GCW")) 
+pca_msm = fit_pca_function(data_wide_PCA2 %>% filter(site == "MSM")) 
+pca_gwi = fit_pca_function(data_wide_PCA2 %>% filter(site == "GWI"))
 
@@
 
+pc_region_transect %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable, color = transect))+
+  geom_point(size = 5)+
+#  scale_colour_manual(values = c("red", "blue"))+
+ # xlim(0, 0.4)+
+  #geom_vline(xintercept = 0)+
+  labs(y = "",
+       x = "|PC1|")+
+  facet_grid(. ~ region)+
+  theme_kp()+
+  geom_vline(xintercept = 0.3, linetype = "dashed")+
+  geom_vline(xintercept = 0.1, linetype = "dashed")
+
 ```
@@
 
-df = data_wide_PCA %>% filter(region == "CB") %>%   drop_na() %>% dplyr::select(where(is.numeric))
-grp = data_wide_PCA %>% filter(region == "CB")%>% drop_na() %>% dplyr::select(!where(is.numeric))
+df = data_wide_PCA2 %>% filter(region == "WLE") %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA2 %>% filter(region == "WLE")%>% drop_na() %>% dplyr::select(!where(is.numeric))
 
@@
   geom_point(size=3,stroke=1, alpha = 1,
-             aes(shape = pca_cluster$grp$site,
+             aes(shape = pca_cluster$grp$transect,
                  color = groups))+ 
@@
 
+## PCA on C-normalized data
+
+```{r}
+data_wide_PCA2 = 
+  data_wide_PCA %>% 
+  pivot_longer(cols = c(where(is.numeric), -TC)) %>% 
+  mutate(value = value/TC) %>% 
+  dplyr::select(-TC) %>% 
+  pivot_wider()
+
+pca_all_c = fit_pca_function(data_wide_PCA2)
+
+ggbiplot(pca_all_c$pca_int, obs.scale = 1, var.scale = 1,
+         groups = as.character(pca_all_c$grp$region), 
+         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(#shape = pca_all_c$grp$site,
+                 color = groups))+ 
+  scale_color_manual(#breaks = c("upland", "transition", "wte", "wc"), 
+                     values = pal_transect)+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  xlim(-6,6)+
+  ylim(-4.5,4.5)+
+  labs(shape="",
+       title = "PCA: all, normalized to C",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+```
+
+```{r}
+df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+library(ggConvexHull)
+
+pca_cluster = fit_pca_function(df_cl)
+
+ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
+                   groups = as.character(pca_cluster$grp$cluster), 
+                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_convexhull(aes(group = pca_cluster$grp$cluster,
+                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_cluster$grp$site,
+                 color = groups))+ 
+  # scale_color_manual(values = c("#16879C", "#BB281E" ))+
+  #scale_shape_manual(values = c(21, 19))+
+  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
+  # xlim(-4,20)+
+  # ylim(-8,8)+
+  labs(fill="cluster",
+       color = "cluster",
+       shape = "region",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+df_cl %>% 
+  ggplot(aes(x = transect, y = cluster, color = site))+
+  geom_jitter(width = 0.1, height = 0.1, size = 3)+
+  facet_wrap(~region)+
+  theme_kp()
+```
+
+---
+
 ## Session Info

### 2024-07-29 3-reports/manuscript_figures.Rmd
manuscript_figures report

---
title: "Synpotic Soil Characterization"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

---

```{r}
source("2-code/0-packages.R")
library(ggbiplot)

pal_transect = rev(soilpalettes::soil_palette("redox2", 4))
theme_set(theme_kp())
```

```{r targets}
targets::tar_load(c(data_combined, sample_key))
```

```{r data_processing_functions}
make_data_subset = function(data_combined){

 # data_combined_subset = 
    data_combined %>% 
    filter(!transect %in% "wte") %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

  #data_combined_wide = 
    data_combined_subset %>% 
    dplyr::select(sample_label, analysis, name, value) %>% 
    separate(name, sep = "_", into = "variable", remove = F) %>% 
    dplyr::select(sample_label, variable, value) %>% 
    pivot_wider(names_from = "variable") %>% 
    left_join(sample_key) %>% 
    filter(!grepl("016", sample_label)) %>% # this one sample is very weird
    force()

}
```

```{r data_processing, eval=TRUE}

data_combined_subset = make_data_subset(data_combined)

data_combined_wide = make_data_wide(data_combined_subset, sample_key) %>% 
  dplyr::select(-gwc) %>% 
    rename(CEC = cec,
         NH4N = nh4n,
         NO3N = no3n,
         P = mehlichp,
         WEOC = npoc,
         DIC = dic,
         SOM = percentOM)

data_wide_PCA = 
  data_combined_wide %>% 
  dplyr::select(-c("DIC", "SOM"))

print("df created: `data_combined_wide`, `data_wide_PCA`")
```

---

## 1. PCAs

```{r pca_function}
fit_pca_function = function(dat){

  dat %>% 
    drop_na()

  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    dplyr::mutate(row = row_number()) %>% 
    drop_na()

  num_row_numbers = num %>% dplyr::select(row)

  grp = 
    dat %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

  num = num %>% dplyr::select(-row)
  pca_int = prcomp(num, scale. = T)

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}
```

---

### Overall PCAs

```{r pca_input}
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
```

```{r pca_gg, fig.width=10, fig.height=6}
## PCA plots overall
#gg_pca_overall = 
ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_overall$grp$region), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
                 fill = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E"))+
  scale_fill_manual(values = c("#16879C", "#BB281E"))+
  scale_shape_manual(values = c(23, 21))+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  guides(fill = "none")+
  NULL

```

```{r pca_gg_regions, fig.width=10, fig.height=6}

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

### Drivers and loadings

```{r pc_loadings}
get_pc_values = function(dat){

  pca = fit_pca_function(dat)

  x = pca$pca_int$rotation
  x2 = 
    x %>% 
    as.data.frame() %>% 
    dplyr::select(PC1) %>% 
    arrange(desc(PC1)) %>% 
    rownames_to_column("variable") %>% 
    mutate(variable = fct_reorder(variable, (PC1)))

  x2
}

pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")

```

```{r pc_gg, fig.height=8, fig.width=6}

pc_all = 
  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
  mutate(variable = fct_reorder(variable, (PC1)))

non_absolute = 
  pc_all %>% 
  mutate(variable = fct_reorder(variable, (PC1))) %>% 
  filter(region != "combined") %>% 
  ggplot(aes(x = PC1, y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

pc_all %>% 
  filter(region != "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+
  labs(y = "",
       x = "|PC1|")+
  theme_kp()

```

### Clustering

```{r cluster_calc, include=FALSE}

library(cluster)
library(factoextra)

df = data_wide_PCA %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA %>% drop_na() %>% dplyr::select(!where(is.numeric))

fviz_nbclust(df, FUN = hcut, method = "wss")
fviz_nbclust(df, FUN = hcut, method = "silhouette")
# three clusters

d <- dist(df, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
plot(hc1, cex = 0.6, hang = -1)
hc5 <- hclust(d, method = "ward.D2" )
sub_grp <- cutree(hc5, k = 3)
table(sub_grp)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 3, border = 2:5)

df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
count(df_cl,cluster)

```

```{r cluster_gg}
df_cl %>%
  ggplot(aes(y = cluster, x = site))+
  geom_point(aes(color = transect, shape = transect),
             size = 4, stroke = 1,
             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
                                             dodge.width = 0.4))+
  scale_color_manual(values = pal_transect)

```

### PCA with clusters

```{r cluster_pca, fig.width=10, fig.height=6}
library(ggConvexHull)

pca_cluster = fit_pca_function(df_cl)

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$region,
                 fill = groups), color = "black")+ 
  scale_shape_manual(values = c(23, 21))+
  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

```

---

## 2. Each analyte

### Normalized values - v1

```{r scaled_gg1, fig.height = 8, fig.width = 7}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaling_max_min = 
  data_long %>% 
  filter(transect != "transition") %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::summarise(max = max(value, na.rm = T),
                   min = min(value, na.rm = T))

data_scaled2 = 
  data_long %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  left_join(data_scaling_max_min) %>% 
  mutate(scaled = (value - min) / (max - min)) %>% 
  dplyr::select(region, site, transect, horizon, name, scaled) %>%
  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
  mutate(correction = case_when(upland < wetland ~ 1,
                                upland > wetland ~ -1)) %>% 
  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
  mutate(scaled = scaled * correction,
         scaled = case_when(transect == "upland" ~ 0,
                            transect == "wetland" ~ 1,
                            TRUE ~ scaled),
         scaled2 = case_when(scaled > 4 ~ 4,
                             scaled < -1 ~ -1,
                             TRUE ~ scaled),
         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
  # order analytes
  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
                                        "pH", "spConductance",
                                        "WEOC", "DIC",
                                        "Fe", "P",
                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
                                        "NH4N", "NO3N", "Cl", "SO4")),
         site = factor(site, levels = c("CRC", "PTR", "OWC",
                                        "GCW", "MSM", "GWI")))

data_scaled2 %>%
  ggplot(aes(x = scaled2, y = name))+
  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
               color = "black", linetype = "dashed", linewidth = 0.2)+
  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
             color = "black",
             size = 1)+
  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
  annotate("text", label = "U", x = 0, y = 21)+
  annotate("text", label = "W", x = 1, y = 21)+
  annotate("text", label = "", x = 1, y = 21.5)+
  xlim(-1, 5)+
  scale_y_discrete(limits = rev)+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                     values = c(21, 22, 24, 21, 22, 24))+
  labs(x = "",
       y = "",
       shape = "transition soils at")+
  facet_wrap(~region)+
  theme_kp()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
        legend.background = element_blank()
  )

```

### Normalized values - v2

```{r scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaled_NEW = 
  data_long_all %>% 
  group_by(region, site, horizon, transect, name) %>% 
  dplyr::summarise(value_scaled = median(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::mutate(value_scaled = scales::rescale(value_scaled)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland")))

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

```{r scaled_gg2_split, fig.height = 16, fig.width = 16}
data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name + region)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

### GWI only

```{r scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name + region)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

### Specific analytes

```{r analytes, fig.height = 4, fig.width = 4}

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 800, label = label), size = 5)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 450, label = label), size = 5)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

```

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>

### 2024-07-30 3-reports/manuscript_figures.Rmd
overall drivers. permanova. analyte by region.

@@
 
+overall = 
+  pc_all %>% 
+  filter(region == "combined") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
 non_absolute = 
@@
              aes(shape = pca_cluster$grp$region,
-                 fill = groups), color = "black")+ 
+                 fill = groups))+ 
   scale_shape_manual(values = c(23, 21))+
+  scale_color_manual(values = pnw_palette("Sailboat", 3))+
   scale_fill_manual(values = pnw_palette("Sailboat", 3))+
@@
 ---
+
+## 1b. PERMANOVA
+
+```{r}
+library(vegan)
+
+adonis2((data_wide_PCA %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
+         (region + transect + horizon + site)^2, 
+       data = data_wide_PCA %>% drop_na)
+
+```
 
@@
 data_combined_wide %>% 
+  filter(region == "CB") %>% 
   mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
@@
        y = "Extractable Fe, μg/g")+
+  facet_wrap(~region)+
   scale_fill_manual(values = pal_transect)+
@@
 data_combined_wide %>% 
+    filter(region == "WLE") %>% 
   mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
@@
        y = "Extractable P, μg/g")+
+  facet_wrap(~region)+
   scale_fill_manual(values = pal_transect)+

### 2024-07-30 3-reports/manuscript_figures.Rmd
renaming bc f-ing sync issues
analytes figures

@@
 
-```{r analytes, fig.height = 4, fig.width = 4}
+Version 1 - one region only
+
+```{r analytesx, fig.height = 4, fig.width = 4}
 
@@
 
+Version 2 - both regions
+
+```{r analytesx-2, fig.height = 4, fig.width = 8}
+data_combined_wide %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+  ggplot(aes(x = transect, y = Fe))+
+  geom_violin(aes(fill = transect), alpha = 0.4)+
+  geom_jitter(width = 0.1)+
+  labs(x = "",
+       y = "Extractable Fe, μg/g")+
+  facet_wrap(~region)+
+  scale_fill_manual(values = pal_transect)+
+  theme(legend.position = "none")
+
+data_combined_wide %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+  ggplot(aes(x = transect, y = P))+
+  geom_violin(aes(fill = transect), alpha = 0.4)+
+  geom_jitter(width = 0.1)+
+  labs(x = "",
+       y = "Extractable P, μg/g")+
+  facet_wrap(~region)+
+  scale_fill_manual(values = pal_transect)+
+  theme(legend.position = "none")
+```
+
 ---

### 2024-08-06 pr_ml/240804_initial_RF.Rmd
Upload initial Random Forest models

---
title: "First looks at RF and MI"
author: "PR"
date: "2024-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)
```

#### Goals

##### Use Random Forest and Mutual Information analyses to understand BGC relationsips across synoptic sites

Both Random Forests (RF) and Mutual Information (MI) are designed to understand how multiple predictors relate to some dependent variable. Since we're interested in how parameters relate to each other across various delineations (regions, sites, transect locations), we'll be using these as our dependents. I'll start with Random Forests because I have more experience with them and they're easier to interpret IMO.

We'll construct RF models for

1.  CB v WLE
2.  UP v TR v W
3.  Each site

Because the goal is to understand relationships in biogeochemistry and not to predict a specific variable, we will not be splitting data into training and testing datasets. This means we **should not** interpret goodness-of-fit metrics as how well we can "predict" if a sample comes from a given location, but rather how well the variables used as predictors explain the variability in the dataset. This is something of a subtle difference, but I think an important one.

```{r read in data}

#install_github("kaizadp/soilpalettes")

source("../2-code/0-packages.R")

require(pacman)
p_load(janitor, #I compulsively use clean_names()... sorry to add dependencie
       ranger, #actual RF algorithm
       PNWcolors, #change to your preferred color scheme, just using something familiar
       cowplot,
       tidymodels) #make modeling tidy again

## Read in data and add white noise as a diagnostic
df_raw <- read_csv("../1-data/processed/chemistry_combined_PCA.csv") %>% 
  clean_names() %>% 
  mutate(white_noise = rnorm(1:n(), mean = 0, sd = 1)) %>% 
  dplyr::select(-c(sample_label, tree_number))

## Losing 28 rows, primarily to ts, fe, p, but also some other stuff
# df_raw %>% #non-predictors
#   filter(if_any(everything(), is.na))

## We'll go with more vars over more rows for now
df <- df_raw %>% 
  drop_na()
```

### Random forest classification model results {.tabset}

```{r set up model function}

make_rf_model <- function(var){

  # 1) Dynamically create formula
  formula <- as.formula(paste(var, "~ ."))

  # 2) Tune mtry and ntree hyperparameters
  rf_spec <- rand_forest() %>%
    set_engine("ranger", importance = "impurity") %>% #using the ranger pkg instead of rf; importance needed for FI
    set_mode("classification")

  # 3) Define a recipe
  recipe <- recipe(formula, data = df) %>%
    step_integer(all_nominal_predictors()) %>% # integerize all dbl
    step_corr(all_predictors()) %>% # check correlations between predictors and remove any strong correlations
    step_normalize(all_predictors(), -all_outcomes()) # normalize non-normally distributed predictors

  # 4) Define a workflow
  wf <- workflow() %>%
    add_model(rf_spec) %>%
    add_recipe(recipe)

  # 5) Run the  model
  rf_model <- wf %>%
    update_model(rf_spec) %>%
    fit(data = df)

  ## Extract predictions
  predictions <- predict(rf_model, df) %>% rename("predicted" = .pred_class)

  ### Feature Importance
  ## Set vectors for renaming stuff
  var_names <- rf_model$fit$fit$fit$variable.importance

  ## Convert feature importance to a tibble with variables as a column
  fi0 <- as.data.frame(var_names) %>%
    tibble::rownames_to_column() %>%
    as_tibble() %>% 
    rename("predictor" = rowname, 
           "raw_fi" = var_names) %>% 
    mutate(fi_n = raw_fi / sum(raw_fi))

  ## Bind predictions to data
  prediction_df <- bind_cols(df, predictions)

  ## Return two things: how good the model did and which variables are most important
  return(list(predictions = prediction_df,
                feature_importance = fi0))
}
```

```{r set up wrapper plotting function}

# Wrapper function for plotting
plot_rf_results <- function(var, plot_title) {

  # Call the make_rf_model function
  model_results <- make_rf_model(var)

  ## Plot for classification equivalent to goodness-of-fit
  p1 <- ggplot(model_results$predictions, aes_string(x = var, y = "predicted")) + 
    geom_jitter(width = 0.2, 
                height = 0.2,
                alpha = 0.5) + 
    ggtitle(paste("Predictions:", var))

  ## Set up a color palette
  var_colors <- PNWColors::pnw_palette("Bay", n = length(unique(model_results$feature_importance$predictor)))

  ## Plot of feature importance
  p2 <- ggplot(model_results$feature_importance, aes(x = fi_n * 100, 
                                                     y = reorder(predictor, fi_n), 
                                                     fill = predictor)) + 
    geom_col(alpha = 0.8, show.legend = F, width = 0.7) + 
    scale_fill_manual(values = var_colors) + 
    labs(x = "Feature Importance (%)", 
         y = "", fill = "") + 
    ggtitle(paste("Feature Imp.:", var))

  ## Combine into a plot
  plot_grid(p1, p2, nrow = 1)
}

```

#### Between regions

```{r}
plot_rf_results("region")
```

**Initial thoughts - Region**
 - looks like region was correctly predicted for all data
 - NO3 as the most important seems sensible, a little surprised it wasn't salt-related though
 - white noise near bottom (transect does not contribute more than random noise to the model so we should ignore it here)

#### Between transect locations

```{r}
plot_rf_results("transect")
```

**Initial thoughts - Transect**
 - all but one correctly classified (one Transition looked like Upland)
 - sulfate is interesting, maybe the relative transition in redox is more consistent than the relative transition in other things across CB and WLE sites
 - neither horizon nor region is important to this model

#### Between sites

```{r}
plot_rf_results("site")
```

**Initial thoughts - Site**
 - All correctly identified

###

#### A couple quick cautions

  1. I haven't done classification RF before, so 1) there's probably a better way to quantify GOF, and 2) it would be good if we go this direction for me to take a bit of a deep dive into properly constructing and vetting an RF classification model. 
  2. I'm a little suspicious that the models performed "too" good... I was expecting them to get maybe half of the values right, 100% makes me suspicious.

### 2024-08-06 pr_ml/temp_build_model.R

@@ -1,0 +1,122 @@
+
+
+source("2-code/0-packages.R")
+
+require(pacman)
+p_load(janitor, #I compulsively use clean_names()... sorry to add dependencie
+       ranger, #actual RF algorithm
+       PNWcolors, #change to your preferred color scheme, just using something familiar
+       cowplot,
+       tidymodels) #make modeling tidy again
+
+df_raw <- read_csv("1-data/processed/chemistry_combined_PCA.csv") %>% 
+  clean_names() %>% 
+  mutate(white_noise = rnorm(1:n(), mean = 0, sd = 1)) %>% 
+  dplyr::select(-c(sample_label, tree_number)) 
+
+df <- df_raw %>% 
+  drop_na()
+
+make_rf_model <- function(var){
+  
+  # 1) Dynamically create formula
+  formula <- as.formula(paste(var, "~ ."))
+  
+  # 2) Tune mtry and ntree hyperparameters
+  rf_spec <- rand_forest() %>%
+    set_engine("ranger", importance = "impurity") %>% #using the ranger pkg instead of rf; importance sets up for graph
+    set_mode("classification")
+  
+  # 3) Define a recipe
+  recipe <- recipe(formula, data = df) %>%
+    #step_dummy(all_nominal(), -all_outcomes()) %>% # One-hot encoding was recommended. I'm not gonna lie, I don't know what this is or what it does...
+    step_integer(all_nominal_predictors()) %>%
+    step_corr(all_predictors()) %>%
+    step_normalize(all_predictors(), -all_outcomes())
+  
+  # 4) Define a workflow
+  wf <- workflow() %>%
+    add_model(rf_spec) %>%
+    add_recipe(recipe)
+  
+  # 5) Run the  model
+  rf_model <- wf %>%
+    update_model(rf_spec) %>%
+    fit(data = df)
+  
+  ## Predictions
+  predictions <- predict(rf_model, df) %>% rename("predicted" = .pred_class)
+  
+  ### Feature Importance
+  ## Set vectors for renaming stuff
+  var_names <- rf_model$fit$fit$fit$variable.importance
+  
+  ## Convert feature importance to a tibble with variables as a column
+  fi0 <- as.data.frame(var_names) %>%
+    tibble::rownames_to_column() %>%
+    as_tibble() %>% 
+    rename("predictor" = rowname, 
+           "raw_fi" = var_names) %>% 
+    mutate(fi_n = raw_fi / sum(raw_fi))
+  
[…]

### 2024-08-06 pr_ml/240804_initial_RF.Rmd
update ugly formatting

@@
 This is something of a subtle difference, but I think an important one.
+
+Note that I'm adding a white noise variable to all models. That's there as a guide for interpreting which variables are contributing more to the model than random chance.
 
@@
 **Initial thoughts - Region**
+
   - looks like region was correctly predicted for all data 
@@
 **Initial thoughts - Transect**
+
   - all but one correctly classified (one Transition looked like Upland) 
@@
 **Initial thoughts - Site**
+
   - All correctly identified
+  - NO3 and TN as top variables followed by region and horizon seems like a nice mix! Much less "one variable to rule them all" vibes
+  - Transect not important to this model
 

### 2024-08-06 pr_ml/temp_build_model.R

@@ -95,7 +95,7 @@ plot_rf_results <- function(var, plot_title) {
   plot_grid(p1, p2, nrow = 1)
 }
 
-plot_rf_results("region")
+plot_rf_results("transect")
 
 
 

### 2024-08-20 3-reports/manuscript_figures.Rmd
update rmd

@@
 
-```{r scaled_gg2_split, fig.height = 16, fig.width = 16}
+```{r scaled_gg2_split, fig.height = 16, fig.width = 18}
+library(ggh4x) # for nested facets
+# https://teunbrand.github.io/ggh4x/reference/facet_nested.html
+
 data_scaled_NEW %>% 
@@
   geom_line(linewidth = 1) + geom_point(size = 2)+
-  facet_wrap( ~ name + region)+
+#  facet_wrap( ~ name)+
+  facet_nested_wrap(~ name + region, ncol = 8,
+                    nest_line = element_line(colour = "grey"))+ 
   scale_color_manual(breaks = c("CRC", "PTR", "OWC",
@@
 ### Specific analytes
-
-Version 1 - one region only
-
-```{r analytesx, fig.height = 4, fig.width = 4}
-
-data_combined_wide %>% 
-  filter(region == "CB") %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
-  ggplot(aes(x = transect, y = Fe))+
-  geom_violin(aes(fill = transect), alpha = 0.4)+
-  geom_jitter(width = 0.1)+
-  geom_text(data = tibble(x = c("upland", "transition", "wc"),
-                          label = c("B", "A", "B")),
-            aes(x = x, y = 800, label = label), size = 5)+
-  labs(x = "",
-       y = "Extractable Fe, μg/g")+
-  facet_wrap(~region)+
-  scale_fill_manual(values = pal_transect)+
-  theme(legend.position = "none")
-
-data_combined_wide %>% 
-    filter(region == "WLE") %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
-  ggplot(aes(x = transect, y = P))+
-  geom_violin(aes(fill = transect), alpha = 0.4)+
-  geom_jitter(width = 0.1)+
-  geom_text(data = tibble(x = c("upland", "transition", "wc"),
-                          label = c("B", "A", "B")),
-            aes(x = x, y = 450, label = label), size = 5)+
-  labs(x = "",
-       y = "Extractable P, μg/g")+
-  facet_wrap(~region)+
-  scale_fill_manual(values = pal_transect)+
-  theme(legend.position = "none")
-
-```
-
-Version 2 - both regions
 

### 2024-08-20 _targets.R
move data files to `data/raw`

@@ -41,40 +41,40 @@ list(
   # data files
   tar_target(moisture_data, import_gsheet("1nYzExPmpv01IYalqo2L3pmDsBDvhjdvHYMJo31N5_MA")),
   tar_target(moisture_processed, process_moisture(moisture_data)),
-  tar_target(loi_data, import_loi(FILEPATH = "1-data/loi")),
+  tar_target(loi_data, import_loi(FILEPATH = "1-data/raw/loi")),
   tar_target(loi_processed, process_loi(loi_data)),
   tar_target(pH_data, import_gsheet("177ZR7O9JpsgVmFfLh72LfJ72oM-oTc6k_r1qE0XHL64")),
   tar_target(pH_processed, process_pH(pH_data)),
-  tar_target(tctn_data, import_tctn_data(FILEPATH = "1-data/tctnts")),
-  tar_target(ts_data, import_ts_data(FILEPATH = "1-data/tctnts")),
+  tar_target(tctn_data, import_tctn_data(FILEPATH = "1-data/raw/tctnts")),
+  tar_target(ts_data, import_ts_data(FILEPATH = "1-data/raw/tctnts")),
   tar_target(tctnts_data_samples, process_tctnts(tctn_data, ts_data)),
-  tar_target(weoc_data, import_weoc_data(FILEPATH = "1-data/npoc", PATTERN = "Summary_Raw")),
+  tar_target(weoc_data, import_weoc_data(FILEPATH = "1-data/raw/npoc", PATTERN = "Summary_Raw")),
   tar_target(weoc_processed, process_weoc(weoc_data, analysis_key, moisture_processed, subsampling)),
-  tar_target(dic_data, import_dic_data(FILEPATH = "1-data/dic")),
+  tar_target(dic_data, import_dic_data(FILEPATH = "1-data/raw/dic")),
   tar_target(dic_processed, process_dic(dic_data, analysis_key, moisture_processed, subsampling)),
-  #tar_target(din_data, readxl::read_xlsx("1-data/din/223013 Patel Final Report.xlsx", sheet = "NO3-N and NH4-N data")),
-  tar_target(din_data, import_din_data("1-data/din")),
+  #tar_target(din_data, readxl::read_xlsx("1-data/raw/din/223013 Patel Final Report.xlsx", sheet = "NO3-N and NH4-N data")),
+  tar_target(din_data, import_din_data("1-data/raw/din")),
[…]

### 2024-08-20 2-code/b-functions_analysis.R
ignore DIC for now. #4

@@ -1306,6 +1306,7 @@ make_summary_tables <- function(data_combined){
   
   hsd_transect_by_site = 
     data_combined_subset %>% 
+    filter(!grepl("DIC", analysis)) %>% 
     group_by(analysis, name, region, site) %>% 
     do(fit_hsd_transect(.))
   

### 2024-08-20 _targets.R
skipping psa and xrd plots for now. #5

@@ -80,7 +80,7 @@ list(
   #tar_target(hydrometer_df, read.csv(hydrometer_data)),
   tar_target(texture_processed, compute_texture()),
   tar_target(gg_wrc, plot_wrc(wrc_processed)),
-  tar_target(gg_texture, plot_texture(texture_processed, sample_key)),
+#  tar_target(gg_texture, plot_texture(texture_processed, sample_key)),
   
   # analysis - graphs
   ## tar_target(gg_moisture, plot_moisture(moisture_processed, sample_key)),

### 2024-10-16 3-reports/manuscript_figures.Rmd
clean
add correlation matrix. new version of all analyses with only subset of variables (dropped the highly correlated variables)

@@
 
-```{r data_processing_functions}
+```{r data_processing}
 make_data_subset = function(data_combined){
@@
 }
-```
-
-```{r data_processing, eval=TRUE}
 
@@
 ---
+
+## 0. Correlations
+
+```{r gg_corr, fig.height=8, fig.width=8}
+
+fit_correlations_function = function(dat, TITLE){
+  num = 
+    dat %>%       
+    dplyr::select(where(is.numeric)) %>%
+    drop_na()
+
+  num_clean = 
+    num %>% 
+    rownames_to_column("row") %>% 
+    pivot_longer(-row) %>% 
+    separate(name, sep = "_", into = c("name")) %>% 
+    pivot_wider() %>% 
+    dplyr::select(-row)
+
+  m = cor(num_clean)
+  m = round(m, 1)
+
+  p.mat <- ggcorrplot::cor_pmat(num_clean)
+  ggcorrplot::ggcorrplot(m, type = "lower",
+                         hc.order = TRUE,
+                         p.mat = p.mat,
+                         outline.color = "black",
+                         lab = TRUE, 
+                         insig = "blank",
+                         #colors = c("#E46726", "white", "#6D9EC1"),
+  ) +
+    scale_fill_gradient2(low = "orangered2", mid = "white", high = "turquoise4", limit = c(-1, 1), midpoint = 0)+
+    theme(legend.position = "none")
+
+}
+
+corr_all = fit_correlations_function(dat = data_wide_PCA)
+#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "WLE"), TITLE = "WLE" )
+#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "CB"), TITLE = "CB")
+
+corr_all
+```
+
+Highly correlated variables include:
+
+- (+) TC & WEOC, TN
+- (+) TS & Chloride, Sulfate, Na, CEC
+- (+) CEC & TS, Na, K, Mg, SpConductance
+- (+) Na & SpConductance, TS, K, Mg
+- (+) Chloride & Sulfate, TS, CEC, SpConductance, Na
+- (-) pH & WEOC, Al
+- (-) Ca and Al
+
+---
+
+# VERSION 1: ALL ANALYTES
+
+<details>
+  <summary> Click to Open </summary>
 
@@
 
-library(cluster)
+#library(cluster)
 library(factoextra)
@@
 
-### Specific analytes
+</details>
+
+# VERSION 2: DROPPING CORRELATED ANALYTES
+
+### Overall PCAs
+
+```{r drop}
+# drop correlated analytes
+
+data_wide_PCA2 = data_wide_PCA %>% 
+  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance))
+```
+
+```{r v2-pca_input}
+pca_overall = fit_pca_function(data_wide_PCA2) 
+pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
+pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
+```
+
+```{r v2-pca_gg, fig.width=10, fig.height=6}
+## PCA plots overall
+#gg_pca_overall = 
+ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+         groups = as.character(pca_overall$grp$region), 
+         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=4,stroke=1, alpha = 1,
+             aes(shape = pca_overall$grp$region,
+                 fill = groups))+ 
+  scale_color_manual(values = c("#16879C", "#BB281E"))+
+  scale_fill_manual(values = c("#16879C", "#BB281E"))+
+  scale_shape_manual(values = c(23, 21))+
+  labs(shape="",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  guides(fill = "none")+
+  NULL
+
+```
+
+```{r v2-pca_gg_regions, fig.width=10, fig.height=6}
+
+gg_pca_wle = 
+  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall_wle$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall_wle$grp$site,
+                 color = groups))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+                     values = pal_transect)+
+#  xlim(-6,6)+
+#  ylim(-4.5,4.5)+
+  labs(shape="",
+       title = "PCA: WLE",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+gg_pca_cb = 
+  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall_cb$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall_cb$grp$site,
+                 color = groups))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+                     values = pal_transect)+
+#  xlim(-5.5,7.5)+
+#  ylim(-5,5)+
+  labs(shape="",
+       title = "PCA: CB",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
+gg_pca_regions
+
+```
+
+### Drivers and loadings
+
+```{r v2-pc_loadings}
+get_pc_values = function(dat){
+
+  pca = fit_pca_function(dat)
+
+  x = pca$pca_int$rotation
+  x2 = 
+    x %>% 
+    as.data.frame() %>% 
+    dplyr::select(PC1) %>% 
+    arrange(desc(PC1)) %>% 
+    rownames_to_column("variable") %>% 
+    mutate(variable = fct_reorder(variable, (PC1)))
+
+  x2
+}
+
+pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
+pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "WLE")) %>% mutate(region = "WLE")
+pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "CB")) %>% mutate(region = "CB")
+
+```
+
+```{r v2-pc_gg, fig.height=8, fig.width=6}
+
+pc_all = 
+  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
+  mutate(variable = fct_reorder(variable, (PC1)))
+
+overall = 
+  pc_all %>% 
+  filter(region == "combined") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
+non_absolute = 
+  pc_all %>% 
+  mutate(variable = fct_reorder(variable, (PC1))) %>% 
+  filter(region != "combined") %>% 
+  ggplot(aes(x = PC1, y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
+pc_all %>% 
+  filter(region != "combined") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+  labs(y = "",
+       x = "|PC1|")+
+  theme_kp()
+
+```
+
+### Clustering
+
+```{r v2-cluster_calc, include=FALSE}
+
+#library(cluster)
+library(factoextra)
+
+df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+```
+
+```{r v2-cluster_gg}
+df_cl %>%
+  ggplot(aes(y = cluster, x = site))+
+  geom_point(aes(color = transect, shape = transect),
+             size = 4, stroke = 1,
+             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
+                                             dodge.width = 0.4))+
+  scale_color_manual(values = pal_transect)
+
+```
+
+### PCA with clusters
+
+```{r v2-cluster_pca, fig.width=10, fig.height=6}
+library(ggConvexHull)
+
+pca_cluster = fit_pca_function(df_cl)
+
+ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
+                   groups = as.character(pca_cluster$grp$cluster), 
+                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_convexhull(aes(group = pca_cluster$grp$cluster,
+                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
+  geom_point(size=4,stroke=1, alpha = 1,
+             aes(shape = pca_cluster$grp$region,
+                 fill = groups))+ 
+  scale_shape_manual(values = c(23, 21))+
+  scale_color_manual(values = pnw_palette("Sailboat", 3))+
+  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
+  labs(fill="cluster",
+       color = "cluster",
+       shape = "region",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  NULL
+
+```
+
+---
+
+## 1b. PERMANOVA
+
+```{r}
+library(vegan)
+
+adonis2((data_wide_PCA2 %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
+         (region + transect + horizon + site)^2, 
+       data = data_wide_PCA2 %>% drop_na)
+
+```
+
+## 2. Each analyte
+
+### Normalized values - v1
+
+```{r v2-scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
+data_long = 
+  data_combined_wide %>% 
+  rename(Cl = Chloride,
+         SO4 = Sulfate) %>% 
+  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
+  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
+  filter(!is.na(value))
+
+data_scaling_max_min = 
+  data_long %>% 
+  filter(transect != "transition") %>% 
+  group_by(region, site, transect, horizon, name) %>% 
+  dplyr::summarise(value = mean(value)) %>% 
+  group_by(region, site, horizon, name) %>% 
+  dplyr::summarise(max = max(value, na.rm = T),
+                   min = min(value, na.rm = T))
+
+data_scaled2 = 
+  data_long %>% 
+  group_by(region, site, transect, horizon, name) %>% 
+  dplyr::summarise(value = mean(value)) %>% 
+  left_join(data_scaling_max_min) %>% 
+  mutate(scaled = (value - min) / (max - min)) %>% 
+  dplyr::select(region, site, transect, horizon, name, scaled) %>%
+  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
+  mutate(correction = case_when(upland < wetland ~ 1,
+                                upland > wetland ~ -1)) %>% 
+  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
+  mutate(scaled = scaled * correction,
+         scaled = case_when(transect == "upland" ~ 0,
+                            transect == "wetland" ~ 1,
+                            TRUE ~ scaled),
+         scaled2 = case_when(scaled > 4 ~ 4,
+                             scaled < -1 ~ -1,
+                             TRUE ~ scaled),
+         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
+  # order analytes
+  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
+                                        "pH", "spConductance",
+                                        "WEOC", "DIC",
+                                        "Fe", "P",
+                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
+                                        "NH4N", "NO3N", "Cl", "SO4")),
+         site = factor(site, levels = c("CRC", "PTR", "OWC",
+                                        "GCW", "MSM", "GWI")))
+
+data_scaled2 %>%
+  ggplot(aes(x = scaled2, y = name))+
+  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
+               color = "black", linetype = "dashed", linewidth = 0.2)+
+  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
+  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
+             color = "black",
+             size = 1)+
+  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
+             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
+  annotate("text", label = "U", x = 0, y = 21)+
+  annotate("text", label = "W", x = 1, y = 21)+
+  annotate("text", label = "", x = 1, y = 21.5)+
+  xlim(-1, 5)+
+  scale_y_discrete(limits = rev)+
+  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
+                     values = c(21, 22, 24, 21, 22, 24))+
+  labs(x = "",
+       y = "",
+       shape = "transition soils at")+
+  facet_wrap(~region)+
+  theme_kp()+
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_blank(),
+        legend.position = "right",
+        legend.background = element_blank()
+  )
+
+```
+
+### Normalized values - v2
+
+```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
+data_long_all = 
+  data_combined_wide %>% 
+  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
+  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
+  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
+  filter(!is.na(value))
+
+data_scaled_NEW = 
+  data_long_all %>% 
+  group_by(region, site, horizon, transect, name) %>% 
+  dplyr::summarise(value_scaled = median(value)) %>% 
+  group_by(region, site, horizon, name) %>% 
+  dplyr::mutate(value_scaled = scales::rescale(value_scaled)) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
+  mutate(name = factor(name, levels = c("TC", "TS", "WEOC",
+                                        "Ca", "Na", "Chloride",
+                                        "NH4N", "Fe", "P")))
+
+data_scaled_NEW %>% 
+  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
+  geom_line(linewidth = 1) + geom_point(size = 2)+
+  facet_wrap( ~ name)+
+  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c("#d73027", "#fc8d59", "#fdbb84", 
+                                "#74a9cf", "#2b8cbe", "#045a8d"))+
+  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  labs(x = "Transect", y = "Normalized concentration")+
+  theme_kp()+
+  theme(axis.text.y = element_blank(),
+        panel.grid = element_blank())
+
+```
+
+```{r v2-scaled_gg2_split, fig.height = 10, fig.width = 14}
+library(ggh4x) # for nested facets
+# https://teunbrand.github.io/ggh4x/reference/facet_nested.html
+
+data_scaled_NEW %>% 
+  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
+  geom_line(linewidth = 1) + geom_point(size = 2)+
+#  facet_wrap( ~ name)+
+  facet_nested_wrap(~ name + region, ncol = 6,
+                    nest_line = element_line(colour = "grey"))+ 
+  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c("#d73027", "#fc8d59", "#fdbb84", 
+                                "#74a9cf", "#2b8cbe", "#045a8d"))+
+  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  labs(x = "Transect", y = "Normalized concentration")+
+  theme_kp()+
+  theme(axis.text.y = element_blank(),
+        panel.grid = element_blank())
+```
+
+### GWI only
+
+```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8}
+data_scaled_NEW %>% 
+  filter(site == "GWI") %>% 
+  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
+  geom_line(linewidth = 1) + geom_point(size = 2)+
+  facet_wrap( ~ name )+
+  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c("#d73027", "#fc8d59", "#fdbb84", 
+                                "#74a9cf", "#2b8cbe", "#045a8d"))+
+  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  labs(x = "Transect", y = "Normalized concentration")+
+  theme(axis.text.y = element_blank(),
+        panel.grid = element_blank())
+
+```
+
+# Specific analytes
 

### 2024-10-23 3-reports/manuscript_figures.Rmd
clean up report

@@
 #gg_pca_overall = 
-ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
          groups = as.character(pca_overall$grp$region), 
@@
 
-```{r v2-pca_gg_regions, fig.width=10, fig.height=6}
+```{r v2-pca_gg_regions, fig.width=12, fig.height=7}
 
 gg_pca_wle = 
-  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
+  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
            groups = as.character(pca_overall_wle$grp$transect), 
@@
                      values = pal_transect)+
-#  xlim(-6,6)+
-#  ylim(-4.5,4.5)+
+  xlim(-3, 5)+
+  ylim(-2.5, 3.5)+
   labs(shape="",
-       title = "PCA: WLE",
-       subtitle = "surface horizons")+
+       title = "PCA: WLE")+
   theme_kp()+
@@
 gg_pca_cb = 
-  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
+  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
            groups = as.character(pca_overall_cb$grp$transect), 
@@
                      values = pal_transect)+
-#  xlim(-5.5,7.5)+
-#  ylim(-5,5)+
+  xlim(-5.5, 3)+
+  ylim(-3, 3.5)+
   labs(shape="",
-       title = "PCA: CB",
-       subtitle = "surface horizons")+
+       title = "PCA: CB")+
   theme_kp()+
@@
 
-gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
+gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1)
 gg_pca_regions
@@
 
-ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
+ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
                    groups = as.character(pca_cluster$grp$cluster), 
@@
 
-```{r analytesx-2, fig.height = 4, fig.width = 8}
+```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
 data_combined_wide %>% 

### 2024-10-23 3-reports/manuscript_figures.Rmd
Fe/P plotting

@@
 
+```{r analytesx-3, fig.height = 10, fig.width = 8}
+
+plot_transect_as_x = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES = "free_x"){
+
+  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+    filter(!is.na(value))
+
+  fit_lme_hsd_transect = function(dat){
+    # this function will assign post-hoc HSD letters to the transect
+    # use lme4::lmer -> multcomp::glht -> multcomp::cld
+    library(multcomp)
+
+    lmer = lme4::lmer(value ~ transect + (1|site), data = dat)
+    h = summary(glht(lmer, linfct = mcp(transect = "Tukey")), test = adjusted("holm"))
+    x = cld(h, decreasing = TRUE) # decreasing order of HSD letters
+    x$mcletters$monospacedLetters %>% # sigh
+      as.data.frame() %>% 
+      rownames_to_column("transect") %>% 
+      rename(groups = ".") %>% 
+      mutate(groups = str_remove_all(groups, " "),
+             groups = toupper(groups))
+
+  }
+
+  # make summary table for HSD letters ----
+  hsd_transect_overall = 
+    data %>% 
+    group_by(region) %>% 
+    do(fit_lme_hsd_transect(.))
+
+  ## then calculate the y-axis position for the labels
+  lme_y =
+    data %>% 
+    #group_by(region) %>% 
+    dplyr::summarise(max = max(value, na.rm =T),
+                     y = max + max/5) %>% 
+    ungroup()
+
+  ## combine them
+  lme_for_graphs = 
+    hsd_transect_overall %>% 
+    cross_join(lme_y %>% dplyr::select(-max))
+
+  #
+  # graphs ----
+
+  ##  ggplot(data,
+  ##         aes(x = transect,
+  ##             y = value
+  ##             #shape = horizon,
+  ##             #alpha = horizon
+  ##         )) +
+  ##    # plot points
+  ##    geom_point(aes(group = site,
+  ##                   color = site),
+  ##               size = 2.5, stroke = 1,
+  ##               position = position_dodge(width = 0.4))+
+  ##    # plot HSD letters
+  ##    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
+  ##    scale_alpha_manual(values = c(1, 0.3))+
+  ##    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
+  ##    facet_wrap(~region, 
+  ##               labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
+  ##    theme_kp()+
+  ##    labs(x = "",
+  ##        # y = YLAB,
+  ##        # title = TITLE,
+  ##        # subtitle = SUBTITLE,
+  ##         shape = "Site",
+  ##         alpha = "Horizon",
+  ##         color = "Site")+
+  ##    theme(legend.position = "right", legend.box = "vertical")+
+  ##    NULL
+
+  ggplot(data,
+         aes(x = transect,
+             y = value
+             #shape = horizon,
+             #alpha = horizon
+         )) +
+    # plot points
+
+    geom_violin(aes(fill = transect), alpha = 0.4)+
+    geom_jitter(width = 0.1)+
+    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
+    labs(x = "",
+         y = "Extractable P, μg/g")+
+    facet_wrap(~region)+
+    scale_fill_manual(values = pal_transect)+
+    labs(x = "",
+         y = YLAB,
+         title = TITLE,
+         subtitle = SUBTITLE,
+         shape = "Site",
+         alpha = "Horizon",
+         color = "Site")+
+    theme(legend.position = "none")+
+    NULL
+
+}
+
+gg_fe = plot_transect_as_x(data_combined_wide %>% rename(value = Fe),
+                   YLAB = "Extractable Fe, μg/g")
+
+gg_p = plot_transect_as_x(data_combined_wide %>% rename(value = P),
+                   YLAB = "Extractable P, μg/g")
+
+gg_weoc = plot_transect_as_x(data_combined_wide %>% rename(value = WEOC),
+                   YLAB = "Water-Extractable Organic C, μg/g")
+
+cowplot::plot_grid(gg_fe, gg_p, ncol = 1)
+
+```
+
 ---

### 2024-10-23 3-reports/manuscript_figures.Rmd
Chesapeake-only plots

@@
 
+## Chesapeake-only plots for GWI
+```{r analytes-gwi, fig.height = 10, fig.width = 5}
+
+# plot Specific Conductance
+gg_spc =
+  data_combined_subset %>% filter(grepl("spCond", name)) %>% filter(region == "CB") %>% 
+  ggplot(aes(x = transect, y = value)) +
+    # plot points
+    geom_point(aes(group = site, color = site),
+               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
+    scale_alpha_manual(values = c(1, 0.3))+
+    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
+                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
+  #  facet_wrap(~region, labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
+    theme_kp()+
+    labs(x = "",
+         y = "Specific Conductance, mS/cm",
+         shape = "Site",
+         color = "Site")+
+    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
+    NULL
+
+# plot Chloride
+gg_cl = 
+  data_combined_subset %>% filter(grepl("Chloride", name)) %>% filter(region == "CB") %>% 
+  ggplot(aes(x = transect, y = value)) +
+    # plot points
+    geom_point(aes(group = site, color = site),
+               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
+    scale_alpha_manual(values = c(1, 0.3))+
+    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
+                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
+  #  facet_wrap(~region, labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
+    theme_kp()+
+    labs(x = "",
+         y = "Chloride, meq/100g",
+         shape = "Site",
+         color = "Site")+
+    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
+    NULL
+
+cowplot::plot_grid(gg_spc, gg_cl, ncol = 1)
+
+```
+
 ---

### 2024-10-23 3-reports/manuscript_figures.Rmd
add tc graphs

@@
 
-gg_weoc = plot_transect_as_x(data_combined_wide %>% rename(value = WEOC),
-                   YLAB = "Water-Extractable Organic C, μg/g")
-
-cowplot::plot_grid(gg_fe, gg_p, ncol = 1)
-
+gg_tc = plot_transect_as_x(data_combined_wide %>% rename(value = TC),
+                   YLAB = "Total Carbon, %")
+
+cowplot::plot_grid(gg_tc, gg_fe, ncol = 1)
+
+```
+
+```{r analytesx-4, fig.height = 15, fig.width = 8}
+cowplot::plot_grid(gg_tc, gg_fe, gg_p, ncol = 1)
 ```

### 2024-11-11 3-reports/manuscript_figures.Rmd
manuscript figures

@@
   ggplot(aes(x = transect, y = Fe))+
-  geom_violin(aes(fill = transect), alpha = 0.4)+
+  geom_violin(aes(fill = region), alpha = 0.6)+
   geom_jitter(width = 0.1)+
@@
   facet_wrap(~region)+
-  scale_fill_manual(values = pal_transect)+
+#  scale_fill_manual(values = c("darkblue", "red"))+
+  scale_fill_manual(values = pnw_palette("Bay", 2))+
   theme(legend.position = "none")
@@
 
-  ##  ggplot(data,
-  ##         aes(x = transect,
-  ##             y = value
-  ##             #shape = horizon,
-  ##             #alpha = horizon
-  ##         )) +
-  ##    # plot points
-  ##    geom_point(aes(group = site,
-  ##                   color = site),
-  ##               size = 2.5, stroke = 1,
-  ##               position = position_dodge(width = 0.4))+
-  ##    # plot HSD letters
-  ##    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
-  ##    scale_alpha_manual(values = c(1, 0.3))+
-  ##    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
-  ##    facet_wrap(~region, 
-  ##               labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
-  ##    theme_kp()+
-  ##    labs(x = "",
-  ##        # y = YLAB,
-  ##        # title = TITLE,
-  ##        # subtitle = SUBTITLE,
-  ##         shape = "Site",
-  ##         alpha = "Horizon",
-  ##         color = "Site")+
-  ##    theme(legend.position = "right", legend.box = "vertical")+
-  ##    NULL
-
   ggplot(data,
@@
 
-    geom_violin(aes(fill = transect), alpha = 0.4)+
+    geom_violin(aes(fill = region), alpha = 0.4)+
     geom_jitter(width = 0.1)+
@@
     facet_wrap(~region)+
-    scale_fill_manual(values = pal_transect)+
+    scale_fill_manual(values = pnw_palette("Bay", 2))+
     labs(x = "",
@@
 
-cowplot::plot_grid(gg_tc, gg_fe, ncol = 1)
+cowplot::plot_grid(gg_fe, gg_p, ncol = 1)
 

### 2024-11-11 3-reports/manuscript_figures.Rmd
map

@@
 print("df created: `data_combined_wide`, `data_wide_PCA`")
+```
+
+---
+
+## Site Map
+
+```{r map, fig.height = 9.5, fig.width = 13}
+
+library(sf) # for map
+library(ggspatial) # for north arrow
+
+## Set CRS
+common_crs <- 4326
+
+## Set map size and point size
+point_size <- 2
+map_width = 9
+map_height = 6
+
+## Set regional and WLE/CB (inset) bounding boxes
+us_bbox <- c(xmin = -125, xmax = -60, ymin = 20, ymax = 50)
+region_bbox <- c(xmin = -98, xmax = -60, ymin = 32, ymax = 50)
+cb_bbox <- c(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5)
+wle_bbox <- c(xmin = -84, xmax = -82, ymin = 41, ymax = 42)
+
+## Make US states map cropped to GL/CB region
+us <- 
+  read_sf("cb_2018_us_state_5m/cb_2018_us_state_5m.shp") %>% 
+  st_transform(., crs = common_crs) %>% 
+  st_crop(., y = us_bbox)
+
+region <- st_crop(us, y = region_bbox)
+
+## Further crop states to WLE/CB region
+cb_states <- st_crop(region, y = cb_bbox)
+wle_states <- st_crop(region, y = wle_bbox)
+
+## Get state labels
+st_labels = st_centroid(us) %>% 
+  filter(!STUSPS %in% c("TN", "NC", "DC")) # remove state labels that mess with the graph
+
+## create a df with site coordinates and labels
+sites = data.frame(
+  y = c(41.61524, 41.50148, 41.37618,
+        38.43085, 37.21927, 38.874445),
+  x = c(-83.22889, -83.04611, -82.50685,
+        -76.22622, -76.40851, -76.551667),
+  region = c("Erie", "Erie", "Erie",
+             "Chesapeake", "Chesapeake", "Chesapeake"),
+  site = c("CRC", "PTR", "OWC",
+           "MSM", "GWI", "GCW"),
+  label = c("1", "2", "3", 
+            "5", "6", "4"))
+
+## Make the base map
+base_plot <- 
+  ggplot() + 
+  geom_sf(data = region) + 
+  #   geom_sf_text(data = st_labels, aes(label = STUSPS))+
+  geom_point(
+    data = sites, aes(x, y),
+    size = 2, color = "black")+
+  # CB inset rectangle
+  geom_rect(aes(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5), 
+            fill = NA, color = "black", lwd = 0.75) +
+  geom_segment(aes(x = -74.5, xend = -70, y = 38, yend = 38), 
+               color = "black", lwd = 0.75) + 
+  # WLE inset rectangle
+  geom_rect(aes(xmin = -84, xmax = -82, ymin = 41, ymax = 42), 
+            fill = NA, color = "black", lwd = 0.75) +
+  geom_segment(aes(x = -83, xend = -83, y = 42, yend = 44), 
+               color = "black", lwd = 0.75) + 
+  annotation_north_arrow(location = "tr", which_north = "true", 
+                         pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
+                         style = north_arrow_fancy_orienteering)+
+  xlim(-92, -67)+
+  ylim(35, 49)+
+  theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
+        legend.key = element_rect(fill = "transparent"), 
+        legend.position = c(0.85, 0.1)) + 
+  labs(x = "", y = "")+
+  NULL
+
+## Make the inset map with just WLE sites
+inset_plot_wle <- 
+  ggplot() + 
+  geom_sf(data = wle_states) + 
+  geom_point(
+    data = sites %>% filter(region == "Erie"), aes(x, y),
+    size = 5, color = "black")+
+  geom_text(data = sites %>% filter(region == "Erie"), aes(x, y, label = label), nudge_x = 0, nudge_y = -0.10, size = 7)+
+  scale_x_continuous(expand = c(0,0))+
+  scale_y_continuous(expand = c(0,0))+
+  cowplot::theme_map() + 
+  theme(legend.position = "none",
+        panel.background = element_rect(fill = "#56cfe1"),
+        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
+        axis.text = element_blank())
+
+## Make the inset map with just CB sites
+inset_plot_cb <- 
+  ggplot() + 
+  geom_sf(data = cb_states) + 
+  geom_point(
+    data = sites %>% filter(region == "Chesapeake"), aes(x, y),
+    size = 5, color = "black")+
+  geom_text(data = sites %>% filter(region == "Chesapeake"), aes(x, y, label = label), nudge_x = -0.25, nudge_y = -0.00, size = 7)+
+  scale_x_continuous(expand = c(0,0))+
+  scale_y_continuous(expand = c(0,0))+
+  cowplot::theme_map() + 
+  theme(legend.position = "none",
+        panel.background = element_rect(fill = "#48bfe3"),
+        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
+        axis.text = element_blank())
+
+## Combine into single figure
+base_plot + 
+  annotation_custom(
+    ggplotGrob(inset_plot_wle), 
+    xmin = -90, xmax = -80, ymin = 42, ymax = 50)+
+  annotation_custom(
+    ggplotGrob(inset_plot_cb), 
+    xmin = -75, xmax = -65, ymin = 35, ymax = 43)+
+  theme_kp()
+
+# save figure 1300 x 950
 ```

### 2024-11-11 3-reports/manuscript_figures.Rmd
fix legends and symbols

@@
            ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
-  geom_point(size=3,stroke=1, alpha = 1,
+  geom_point(size=5,stroke=1, alpha = 1,
              aes(shape = pca_overall_wle$grp$site,
-                 color = groups))+ 
+                 fill = groups), 
+             stroke = 1)+ 
   scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                      values = pal_transect)+
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "wc"), 
+                    values = pal_transect)+
+  scale_shape_manual(values = c(21, 24, 22))+
   xlim(-3, 5)+
   ylim(-2.5, 3.5)+
-  labs(shape="",
-       title = "PCA: WLE")+
+  labs(shape="", fill = "", color = "")+ 
   theme_kp()+
   theme(legend.position = "top", legend.box = "vertical")+
+  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
   NULL
@@
            ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
-  geom_point(size=3,stroke=1, alpha = 1,
+  geom_point(size=5,stroke=1, alpha = 1,
              aes(shape = pca_overall_cb$grp$site,
-                 color = groups))+ 
+                 fill = groups), 
+             stroke = 1)+ 
   scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                      values = pal_transect)+
-  xlim(-5.5, 3)+
-  ylim(-3, 3.5)+
-  labs(shape="",
-       title = "PCA: CB")+
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "wc"), 
+                    values = pal_transect)+
+  scale_shape_manual(values = c(21, 24, 22))+
+  xlim(-5, 3)+
+  ylim(-2.5, 3.5)+
+  labs(shape="", fill = "", color = "")+ 
   theme_kp()+
   theme(legend.position = "top", legend.box = "vertical")+
+  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
   NULL
 
-gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1)
+gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1,
+                                    labels = c("A: WLE", "B: CB"))
 gg_pca_regions
@@
        color = "cluster",
-       shape = "region",
-       title = "Overall PCA, both regions",
-       subtitle = "Surface horizons only")+
-  theme_kp()+
+       shape = "region")+
+  theme_kp()+
+  theme(legend.position = c(0.8, 0.75))+
+  guides(fill = guide_legend(override.aes = list(shape = 21)))+
   NULL

### 2024-11-11 3-reports/manuscript_figures.Rmd
updated scaled plot

@@
 
+```{r v2-scaled_gg2_split-NEW, fig.height = 10, fig.width = 14}
+library(ggh4x) # for nested facets
+# https://teunbrand.github.io/ggh4x/reference/facet_nested.html
+
+data_scaled_NEW2 = 
+  data_scaled_NEW %>% 
+  filter(name %in% c("WEOC", "TS", "Na", "Chloride", "Fe", "P")) %>% 
+    mutate(name = factor(name, levels = c("TC", "TS", "WEOC",
+                                        "Ca", "Na", "Chloride",
+                                        "NH4N", "Fe", "P")))
+
+data_scaled_NEW2 %>% 
+  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
+  geom_line(linewidth = 1) + geom_point(size = 4, aes(shape = site, fill = site), color = "black")+
+#  facet_wrap( ~ name)+
+  facet_nested_wrap(~ name + region, ncol = 4,
+                    nest_line = element_line(colour = "grey"))+ 
+  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c("#d73027", "#fc8d59", "#fdbb84", 
+                                "#74a9cf", "#2b8cbe", "#045a8d"))+
+    scale_fill_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c("#d73027", "#fc8d59", "#fdbb84", 
+                                "#74a9cf", "#2b8cbe", "#045a8d"))+
+  scale_shape_manual(breaks = c("CRC", "PTR", "OWC",
+                                "GCW", "MSM", "GWI"),
+                     values = c(21, 23, 24, 21, 23, 24))+
+  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  labs(x = "Transect", y = "Normalized concentration")+
+  theme_kp()+
+  theme(axis.text.y = element_blank(),
+        panel.grid = element_blank())
+
+```
+
 ### GWI only

### 2024-11-13 2-code/1-initial_processing.R
recode regions (Erie & Chesapeake) and wc (wetland)

@@ -23,7 +23,9 @@ sample_key_subset =
   dplyr::select(sample_label, region, site, transect, tree_number, horizon) %>% 
   filter(!is.na(sample_label)) %>% 
   mutate(transect = tolower(transect),
-         site = recode(site, "PR" = "PTR", "CC" = "CRC", "GCREW" = "GCW"))
+         transect = recode(transect, "wc" = "wetland"),
+         site = recode(site, "PR" = "PTR", "CC" = "CRC", "GCREW" = "GCW"),
+         region = recode(region, "WLE" = "Erie", "CB" = "Chesapeake"))
 sample_key_subset[sample_key_subset == "NULL"] <- NA
 
 ## export
@@ -51,9 +53,4 @@ analysis_key[analysis_key == "NULL"] <- NA
 ## export
 analysis_key %>% write.csv("1-data/analysis_key.csv", row.names = F, na = "")
 
-
 #
-
-# download data files -----------------------------------------------------
-
-

### 2024-11-13 2-code/a-functions_processing.R
recode regions (Erie, Chesapeake) and wc (wetland)

@@ -20,7 +20,7 @@ reorder_horizon = function(dat){
 }
 reorder_transect = function(dat){
   dat %>% 
-    mutate(transect = factor(transect, levels = c("upland", "transition", "wte", "wc", "wetland")))
+    mutate(transect = factor(transect, levels = c("upland", "transition", "wte", "wetland")))
 }
 reorder_site = function(dat){
   dat %>% 
@@ -65,14 +65,14 @@ process_loi = function(loi_data){
   
   loi_data %>%
     janitor::clean_names() %>% 
-    mutate(region = case_when(grepl("WLE", source) ~ "WLE",
-                              grepl("CB", source) ~ "CB")) %>% 
+    mutate(region = case_when(grepl("WLE", source) ~ "Erie",
+                              grepl("CB", source) ~ "Chesapeake")) %>% 
     mutate(label = case_when((customer_id >= 129 & customer_id <= 214) ~ paste0("COMPASS_May22_", customer_id),
                              customer_id >= 215 ~ paste0("COMPASS_Aug22_", customer_id)),
-           sample_label = case_when(region == "WLE" ~ sample_label,
-                                    region == "CB" ~ label)) %>% 
-    mutate(percentOM = case_when(region == "CB" ~ percent_om,
-                                  region == "WLE" ~ 100*(wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g)),
+           sample_label = case_when(region == "Erie" ~ sample_label,
+                                    region == "Chesapeake" ~ label)) %>% 
+    mutate(percentOM = case_when(region == "Chesapeake" ~ percent_om,
+                                  region == "Erie" ~ 100*(wt_tray_drysoil_g - wt_tray_combustedsoil_g)/(wt_tray_drysoil_g - wt_tray_g)),
            percentOM = round(percentOM, 2)) %>%
     dplyr::select(sample_label, percentOM) %>% 
     mutate(analysis = "LOI") %>% 
@@ -390,8 +390,8 @@ process_icp = function(icp_data, analysis_key, moisture_processed, subsampling){
 import_iron = function(FILEPATH){
   
   # import map
[…]

### 2024-11-13 2-code/b-functions_analysis.R

@@ -398,8 +398,8 @@ compute_overall_pca = function(data_combined, sample_key){
   
   ## PCA input files ----
   pca_overall = fit_pca_function(data_combined_wide_NO_DIC) 
-  pca_overall_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "WLE"))
-  pca_overall_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "CB"))
+  pca_overall_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "Erie"))
+  pca_overall_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "Chesapeake"))
   
   
   ## PCA plots overall ----
@@ -429,14 +429,14 @@ compute_overall_pca = function(data_combined, sample_key){
     geom_point(size=3,stroke=1, alpha = 1,
                aes(shape = pca_overall_wle$grp$site,
                    color = groups))+ 
-    scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+    scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                        values = pal_transect)+
     #scale_shape_manual(values = c(21, 19))+
     #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
     xlim(-6,6)+
     ylim(-4.5,4.5)+
     labs(shape="",
-         title = "PCA: WLE",
+         title = "PCA: Erie",
          subtitle = "surface horizons")+
     theme_kp()+
     theme(legend.position = "top", legend.box = "vertical")+
@@ -449,14 +449,14 @@ compute_overall_pca = function(data_combined, sample_key){
     geom_point(size=3,stroke=1, alpha = 1,
                aes(shape = pca_overall_cb$grp$site,
                    color = groups))+ 
-    scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+    scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                        values = pal_transect)+
     #scale_shape_manual(values = c(21, 19))+
     #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
     xlim(-5.5,7.5)+
     ylim(-5,5)+
     labs(shape="",
[…]

### 2024-11-13 3-reports/manuscript_figures.Rmd
recode region and transect

@@
 corr_all = fit_correlations_function(dat = data_wide_PCA)
-#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "WLE"), TITLE = "WLE" )
-#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "CB"), TITLE = "CB")
+#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "Erie"), TITLE = "Erie" )
+#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "Chesapeake"), TITLE = "Chesapeake")
 
@@
 pca_overall = fit_pca_function(data_wide_PCA) 
-pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
-pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
+pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "Erie"))
+pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "Chesapeake"))
 ```
@@
                  color = groups))+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                      values = pal_transect)+
@@
   labs(shape="",
-       title = "PCA: WLE",
+       title = "PCA: Erie",
        subtitle = "surface horizons")+
@@
   labs(shape="",
-       title = "PCA: CB",
+       title = "PCA: Chesapeake",
        subtitle = "surface horizons")+
@@
 pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
-pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
-pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")
+pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "Erie")) %>% mutate(region = "Erie")
+pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")
 
@@
          SO4 = Sulfate) %>% 
-  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
   pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
@@
   data_combined_wide %>% 
-  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
   pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
@@
 pca_overall = fit_pca_function(data_wide_PCA2) 
-pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
-pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
+pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "Erie"))
+pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "Chesapeake"))
 ```
@@
              stroke = 1)+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                      values = pal_transect)+
-  scale_fill_manual(breaks = c("upland", "transition", "wte", "wc"), 
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
@@
 gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1,
-                                    labels = c("A: WLE", "B: CB"))
+                                    labels = c("A: Erie", "B: Chesapeake"))
 gg_pca_regions
@@
 pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
-pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "WLE")) %>% mutate(region = "WLE")
-pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "CB")) %>% mutate(region = "CB")
+pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "Erie")) %>% mutate(region = "Erie")
+pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")
 
@@
   dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
-  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
   pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
@@
 data_combined_wide %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
   ggplot(aes(x = transect, y = Fe))+
@@
 data_combined_wide %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
   ggplot(aes(x = transect, y = P))+
@@
 
-  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
+  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
     filter(!is.na(value))
@@
 gg_spc =
-  data_combined_subset %>% filter(grepl("spCond", name)) %>% filter(region == "CB") %>% 
+  data_combined_subset %>% filter(grepl("spCond", name)) %>% filter(region == "Chesapeake") %>% 
   ggplot(aes(x = transect, y = value)) +
@@
                        values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
-  #  facet_wrap(~region, labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
     theme_kp()+
@@
 gg_cl = 
-  data_combined_subset %>% filter(grepl("Chloride", name)) %>% filter(region == "CB") %>% 
+  data_combined_subset %>% filter(grepl("Chloride", name)) %>% filter(region == "Chesapeake") %>% 
   ggplot(aes(x = transect, y = value)) +

### 2025-02-18 3-reports/manuscript_figures.Rmd
figures for poster

@@
 
+```{r map-poster, fig.height = 9.5, fig.width = 13}
+
+library(sf) # for map
+library(ggspatial) # for north arrow
+
+## Set CRS
+common_crs <- 4326
+
+## Set map size and point size
+point_size <- 2
+map_width = 9
+map_height = 6
+
+## Set regional and WLE/CB (inset) bounding boxes
+us_bbox <- c(xmin = -125, xmax = -60, ymin = 20, ymax = 50)
+region_bbox <- c(xmin = -98, xmax = -60, ymin = 32, ymax = 50)
+cb_bbox <- c(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5)
+wle_bbox <- c(xmin = -84, xmax = -82, ymin = 41, ymax = 42)
+
+## Make US states map cropped to GL/CB region
+us <- 
+  read_sf("cb_2018_us_state_5m/cb_2018_us_state_5m.shp") %>% 
+  st_transform(., crs = common_crs) %>% 
+  st_crop(., y = us_bbox)
+
+region <- st_crop(us, y = region_bbox)
+
+## Further crop states to WLE/CB region
+cb_states <- st_crop(region, y = cb_bbox)
+wle_states <- st_crop(region, y = wle_bbox)
+
+## Get state labels
+st_labels = st_centroid(us) %>% 
+  filter(!STUSPS %in% c("TN", "NC", "DC")) # remove state labels that mess with the graph
+
+## create a df with site coordinates and labels
+sites = data.frame(
+  y = c(41.61524, 41.50148, 41.37618,
+        38.43085, 37.21927, 38.874445),
+  x = c(-83.22889, -83.04611, -82.50685,
+        -76.22622, -76.40851, -76.551667),
+  region = c("Erie", "Erie", "Erie",
+             "Chesapeake", "Chesapeake", "Chesapeake"),
+  site = c("CRC", "PTR", "OWC",
+           "MSM", "GWI", "GCW"),
+  label = c("1", "2", "3", 
+            "5", "6", "4"))
+
+## Make the base map
+base_plot <- 
+  ggplot() + 
+  geom_sf(data = region) + 
+  #   geom_sf_text(data = st_labels, aes(label = STUSPS))+
+  geom_point(
+    data = sites, aes(x, y),
+    size = 2, color = "black")+
+  # CB inset rectangle
+  geom_rect(aes(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5), 
+            fill = NA, color = "black", lwd = 0.75) +
+  geom_segment(aes(x = -74.5, xend = -70, y = 38, yend = 38), 
+               color = "black", lwd = 0.75) + 
+  # WLE inset rectangle
+  geom_rect(aes(xmin = -84, xmax = -82, ymin = 41, ymax = 42), 
+            fill = NA, color = "black", lwd = 0.75) +
+  geom_segment(aes(x = -83, xend = -83, y = 42, yend = 44), 
+               color = "black", lwd = 0.75) + 
+  annotation_north_arrow(location = "tr", which_north = "true", 
+                         pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
+                         style = north_arrow_fancy_orienteering)+
+  xlim(-92, -67)+
+  ylim(35, 49)+
+#  scale_x_continuous(labels = NULL)+
+#  scale_y_continuous(labels = NULL)+
+  theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
+        legend.key = element_rect(fill = "transparent"), 
+        legend.position = c(0.85, 0.1),
+        axis.text = element_blank()) + 
+  labs(x = "", y = "")+
+  theme_void()+
+  NULL
+
+## Make the inset map with just WLE sites
+inset_plot_wle <- 
+  ggplot() + 
+  geom_sf(data = wle_states) + 
+  geom_point(
+    data = sites %>% filter(region == "Erie"), aes(x, y),
+    size = 5, color = "black")+
+  geom_text(data = sites %>% filter(region == "Erie"), aes(x, y, label = label), nudge_x = 0, nudge_y = -0.10, size = 7)+
+  scale_x_continuous(expand = c(0,0))+
+  scale_y_continuous(expand = c(0,0))+
+  cowplot::theme_map() + 
+  theme(legend.position = "none",
+        panel.background = element_rect(fill = "#56cfe1"),
+        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
+        axis.text = element_blank())
+
+## Make the inset map with just CB sites
+inset_plot_cb <- 
+  ggplot() + 
+  geom_sf(data = cb_states) + 
+  geom_point(
+    data = sites %>% filter(region == "Chesapeake"), aes(x, y),
+    size = 5, color = "black")+
+  geom_text(data = sites %>% filter(region == "Chesapeake"), aes(x, y, label = label), nudge_x = -0.25, nudge_y = -0.00, size = 7)+
+  scale_x_continuous(expand = c(0,0))+
+  scale_y_continuous(expand = c(0,0))+
+  cowplot::theme_map() + 
+  theme(legend.position = "none",
+        panel.background = element_rect(fill = "#48bfe3"),
+        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
+        axis.text = element_blank())
+
+## Combine into single figure
+base_plot + 
+  annotation_custom(
+    ggplotGrob(inset_plot_wle), 
+    xmin = -90, xmax = -80, ymin = 42, ymax = 50)+
+  annotation_custom(
+    ggplotGrob(inset_plot_cb), 
+    xmin = -75, xmax = -65, ymin = 35, ymax = 43)
+
+# save figure 1300 x 950
+```
+
 ---
@@
 
+```{r v2-pca_gg_regions-poster, fig.width=7, fig.height=7}
+
+gg_pca_wle = 
+  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
+           groups = as.character(pca_overall_wle$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=5,stroke=1, alpha = 1,
+             aes(shape = pca_overall_wle$grp$site,
+                 fill = groups), 
+             stroke = 1)+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                    values = pal_transect)+
+  scale_shape_manual(values = c(21, 24, 22))+
+  xlim(-3, 5)+
+  ylim(-2.5, 3.5)+
+  labs(shape="", fill = "", color = "", title = "Lake Erie")+ 
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical",
+        plot.title = element_text(face = "bold", size = 18))+
+  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
+  NULL
+
+gg_pca_cb = 
+  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
+           groups = as.character(pca_overall_cb$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=5,stroke=1, alpha = 1,
+             aes(shape = pca_overall_cb$grp$site,
+                 fill = groups), 
+             stroke = 1)+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                    values = pal_transect)+
+  scale_shape_manual(values = c(21, 24, 22))+
+  xlim(-5, 3)+
+  ylim(-2.5, 3.5)+
+  labs(shape="", fill = "", color = "", title = "Chesapeake Bay")+ 
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical",
+        plot.title = element_text(face = "bold", size = 18))+
+  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
+  NULL
+
+gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 2,
+                                    labels = c("Lake Erie", "Chesapeake Bay"))
+#gg_pca_regions
+
+gg_pca_wle
+gg_pca_cb
+
+```
+
 ### Drivers and loadings
@@
 
-```{r v2-scaled_gg2_split-NEW, fig.height = 10, fig.width = 14}
+```{r v2-scaled_gg2_split-NEW, fig.height = 10, fig.width = 12}
 library(ggh4x) # for nested facets

### 2025-02-18 2-code/0-packages.R
update theme_kp

@@ -16,7 +16,7 @@ theme_kp <- function() {  # this for all the elements common across plots
           panel.border = element_rect(color="black",linewidth=1, fill = NA),
           
           plot.title = element_text(hjust = 0, size = 14),
-          axis.text = element_text(size = 12, color = "black"),
+          axis.text = element_text(size = 14, color = "black"),
           axis.title = element_text(size = 14, face = "bold", color = "black"),
           
           # formatting for facets
@@ -24,8 +24,8 @@ theme_kp <- function() {  # this for all the elements common across plots
           strip.background = element_rect(colour="white", fill="white"), #facet formatting
           panel.spacing.x = unit(1, "lines"), #facet spacing for x axis
           panel.spacing.y = unit(1.5, "lines"), #facet spacing for x axis
-          strip.text.x = element_text(size=12, face="bold"), #facet labels
-          strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels
+          strip.text.x = element_text(size=14, face="bold"), #facet labels
+          strip.text.y = element_text(size=14, face="bold", angle = 270) #facet labels
     )
 }
 theme_set(theme_kp())

### 2025-02-18 _targets.R
clean up targets

@@ -1,7 +1,5 @@
 # Created by use_targets().
-# Follow the comments below to fill in this target script.
-# Then follow the manual to check and run the pipeline:
-#   https://books.ropensci.org/targets/walkthrough.html#inspect-the-pipeline # nolint
+# https://books.ropensci.org/targets/walkthrough.html#inspect-the-pipeline # nolint
 
 # Load packages required to define the pipeline:
 library(targets)
@@ -11,13 +9,7 @@ library(tarchetypes) # Load other packages as needed. # nolint
 tar_option_set(
   packages = c("tibble", "tidyverse"), # packages that your targets need to run
   format = "rds" # default storage format
-  # Set other options as needed.
 )
-
-# tar_make_clustermq() configuration (okay to leave alone):
-
-# tar_make_future() configuration (okay to leave alone):
-# Install packages {{future}}, {{future.callr}}, and {{future.batchtools}} to allow use_targets() to configure tar_make_future() options.
 
 # Load the R scripts with your custom functions:
 source("2-code/0-packages.R")
@@ -25,7 +17,6 @@ source("2-code/0-packages.R")
 source("2-code/a-functions_processing.R")
 source("2-code/b-functions_analysis.R")
 source("2-code/c-functions_fticrrr.R")
-# source("other_functions.R") # Source other scripts as needed. # nolint
 
 # Replace the target list below with your own:
 list(
@@ -52,52 +43,30 @@ list(
   tar_target(weoc_processed, process_weoc(weoc_data, analysis_key, moisture_processed, subsampling)),
   tar_target(dic_data, import_dic_data(FILEPATH = "1-data/raw/dic")),
   tar_target(dic_processed, process_dic(dic_data, analysis_key, moisture_processed, subsampling)),
-  #tar_target(din_data, readxl::read_xlsx("1-data/raw/din/223013 Patel Final Report.xlsx", sheet = "NO3-N and NH4-N data")),
   tar_target(din_data, import_din_data("1-data/raw/din")),
   tar_target(din_processed, process_din(din_data, analysis_key, moisture_processed, subsampling)),
   tar_target(icp_data, import_icp_data(FILEPATH = "1-data/raw/icp")),
[…]

### 2025-02-18 2-code/a-functions_processing.R
clean up processing script

@@ -1,4 +1,14 @@
+# SOIL CHARACTERIZATION FOR WLE AND CB SITES
 
+## a-functions_processing.R
+## Use these functions to import and process data files.
+## Some files are on Google Drive, others are local, as .csv/.xlsx.
+## These functions are called in the _targets.R file.
+
+## KFP
+
+######################## ####
+######################## ####
 
 
 # import function(s) ------------------------------------------------------
@@ -8,9 +18,7 @@ import_gsheet = function(dat){
   googlesheets4::read_sheet(dat) %>% mutate_all(as.character)
 }
 
-
 #
-
 # refactor/reorder functions ----------------------------------------------
 ## functions to set the order of factors
 
@@ -27,7 +35,7 @@ reorder_site = function(dat){
     mutate(site = factor(site, levels = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI")))
 }
 
-
+#
 # Chemistry data ----------------------------------------------------------
 
 ## Moisture
@@ -45,9 +53,6 @@ process_moisture = function(moisture_dat){
            gwc_perc = round(gwc_perc, 2)) %>% 
     dplyr::select(sample_label, gwc_perc) %>% 
     mutate(analysis = "GWC")
-  
-  
-  #
 }
 
 
@@ -106,8 +111,8 @@ import_tctn_data = function(FILEPATH){
              TN_perc = `N  [%]`) %>% 
       ## this is a ONE-TIME CORRECTION because of the way the sample names were entered
       mutate(Name = str_replace(Name, "TCTN_CMPS_KFP_0", "COMPASS_Dec2021_")) %>% 
-      #mutate(source = rep(path, nrow(.))) %>% 
       force()
+    
     df}))
   
 }
@@ -120,15 +125,10 @@ import_ts_data = function(FILEPATH){
       dplyr::select(Name, `S  [%]`, Memo) %>% 
       rename(TS_perc = `S  [%]`) %>% 
       mutate(Name = str_replace(Name, "DEC", "Dec")) %>% 
-      #mutate(source = rep(path, nrow(.))) %>% 
       force()
     df}))
   
 }
-
-# tctn_data = import_tctn_data(FILEPATH = "1-data/tctnts")
-# ts_data = import_ts_data(FILEPATH = "1-data/tctnts")
-
 process_tctnts = function(tctn_data, ts_data){
   
   tctn_samples = 
[…]

### 2025-02-21 2-code/a-functions_processing.R
clean

@@ -1273,7 +1273,8 @@ make_summary_tables <- function(data_combined,
     pivot_wider(names_from = "transect", values_from = "summary")
   
   summary_XRD = 
-    xrd_processed %>% subset_surface_horizons(.) %>% 
+    xrd_processed %>% 
+    subset_surface_horizons(.) %>% 
     dplyr::select(-c(tree_number)) %>% 
     pivot_longer(-c(sample_label, region, site, transect, horizon),
                  names_to = "mineral",
@@ -1290,7 +1291,8 @@ make_summary_tables <- function(data_combined,
     reorder_transect() %>% 
     reorder_site() %>% 
     arrange(region, site, transect) %>% 
-    pivot_wider(names_from = "transect", values_from = "summary")
+    filter(transect != "wte") %>% 
+    pivot_wider(names_from = "mineral", values_from = "summary")
   
   summary_wrc = wrc_parameters
   
@@ -1300,12 +1302,12 @@ make_summary_tables <- function(data_combined,
        summary_NUTRIENTS = summary_NUTRIENTS,
        summary_DOC_DIC = summary_DOC_DIC,
        summary_texture = summary_texture,
-       summary_PD = summary_PD,
+#       summary_PD = summary_PD,
        summary_wrc = summary_wrc,
        summary_XRD = summary_XRD)
   
 }
-# summary_table = make_summary_tables(data_combined)
+# summary_tables = make_summary_tables(data_combined)
 # openxlsx::write.xlsx(summary_tables, "summary_tables.xlsx")
 
 

### 2025-02-21 2-code/a-functions_processing.R
summary tables

@@ -1307,7 +1307,7 @@ make_summary_tables <- function(data_combined,
        summary_XRD = summary_XRD)
   
 }
-# summary_tables = make_summary_tables(data_combined)
+# targets::tar_load(summary_tables)
 # openxlsx::write.xlsx(summary_tables, "summary_tables.xlsx")
 
 

### 2025-02-21 _targets.R
delete old targets. add rmarkdown target

@@ -75,17 +75,6 @@ list(
                                          sample_key)),
   tar_target(data_combined, subset_surface_horizons(data_combined_all_horizons)),
   tar_target(data_combined_wide, make_data_wide_processing(data_combined, sample_key)),
-
-##  tar_target(analysis_completion_matrix, compute_analysis_matrix(data_combined)),
-##  tar_target(gg_pca_all, compute_overall_pca(data_combined, sample_key)),
-##  tar_target(gg_correlations, compute_correlations(data_combined, sample_key)),
-##  
-##  tar_target(gg_by_transect_colorsites, make_graphs_by_transect_SITE_AS_COLOR(data_combined)),
-##  tar_target(gg_by_transect, make_graphs_by_transect(data_combined)),
-##  tar_target(gg_by_site_oa, make_graphs_by_site(data_combined %>% filter(horizon != "B"))),
-##  tar_target(gg_by_site_oab, make_graphs_by_site(data_combined_all_horizons)),
-##  tar_target(gg_xrd, plot_xrd(xrd_processed)),
-##  
   tar_target(summary_tables, make_summary_tables(data_combined, pd_processed, xrd_processed, texture_processed, wrc_parameters)),
   
   # export
@@ -97,10 +86,11 @@ list(
 # PD
 # BD
 # WRC
-  }, format = "file")
+  }, format = "file"),
    
    
    # report  
 #   tar_render(report, path = "3-reports/characterization_report.Rmd")
-  
+   tar_render(report, path = "3-reports/manuscript_figures.Rmd")
+
 )

### 2025-02-21 3-reports/manuscript_figures.Rmd
add elevation and hyprop to rmd

@@
 
+## Water retention curves
+
+```{r gg_hyprop, fig.height=8, fig.width=10}
+library(ggh4x)
+
+tar_load(wrc_processed)
+
+wrc_processed %>% 
+  reorder_site() %>% 
+    filter(kpa_fit >= 0 | kpa_eval >= 0) %>% 
+    mutate(region = factor(region, levels = c("Erie", "Chesapeake"))) %>% 
+    ggplot(aes(y = water_content_vol_percent, color = transect))+
+    geom_line(aes(x = kpa_fit))+
+    geom_point(aes(x = kpa_eval), shape = 1)+
+    scale_x_log10(labels = scales::comma)+
+  scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
+  labs(x = "Water potential, kPa",
+       y = "Water content, % v/v",
+       color = "")+
+  #    facet_wrap(~region + site)+
+  facet_nested_wrap(region ~ site)+
+  theme_kp()
+```
+
+## Elevation
+
+```{r, gg_elevation, fig.height=8, fig.width=10}
+
+elevation = read.csv("synoptic-distances.csv", na = "")
+elevation_processed = 
+  elevation %>% 
+  mutate(transect = factor(transect, levels = c("water edge", "wetland", "transition", "upland"))) %>% 
+  reorder_site()
+
+gg_wle = 
+  elevation_processed %>% 
+  filter(region == "Erie") %>% 
+  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect))+
+  geom_line(color = "black")+
+  geom_point(size = 4)+
+  scale_color_manual(values = c("grey", "#2E5894", "#FFB200", "#96001B"),
+                     breaks = c("water edge", "wetland", "transition", "upland"))+
+  facet_nested_wrap(region ~ site)+
+  labs(x = "",
+       y = "Elevation, m
+       ",
+       color = "")+
+#  theme(legend.position = c(0.1,0.61))+
+  NULL
+
+gg_cb = 
+  elevation_processed %>% 
+  filter(region == "Chesapeake") %>%
+  filter(site != "Sweet Hall Marsh") %>% 
+  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect))+
+  geom_line(color = "black")+
+  geom_point(size = 4)+
+  scale_color_manual(values = c("grey", "#2E5894", "#FFB200", "#96001B"),
+                     breaks = c("water edge", "wetland", "transition", "upland"))+
+  facet_nested_wrap(region ~ site)+
+  labs(x = "Distance to water, m",
+       y = "Elevation, m
+       ",
+       color = "")
+
+ggpubr::ggarrange(gg_wle, gg_cb, common.legend = T, nrow = 2)
+
+```
+
 ---

### 2025-02-21 3-reports/manuscript_figures.Rmd
update file path

@@
 
-elevation = read.csv("synoptic-distances.csv", na = "")
+elevation = read.csv("1-data/synoptic-elevations.csv", na = "")
 elevation_processed = 

### 2025-02-21 3-reports/manuscript_figures.Rmd
update figures

@@
 
-```{r map-poster, fig.height = 9.5, fig.width = 13}
+```{r map-poster, fig.height = 9.5, fig.width = 13, eval=FALSE}
 
@@
 
-```{r pca_gg, fig.width=10, fig.height=6}
-## PCA plots overall
-#gg_pca_overall = 
-ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
-         groups = as.character(pca_overall$grp$region), 
-         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
-  geom_point(size=4,stroke=1, alpha = 1,
-             aes(shape = pca_overall$grp$region,
-                 fill = groups))+ 
-  scale_color_manual(values = c("#16879C", "#BB281E"))+
-  scale_fill_manual(values = c("#16879C", "#BB281E"))+
-  scale_shape_manual(values = c(23, 21))+
-  labs(shape="",
-       title = "Overall PCA, both regions",
-       subtitle = "Surface horizons only")+
-  theme_kp()+
-  guides(fill = "none")+
-  NULL
-
-```
-
-```{r pca_gg_regions, fig.width=10, fig.height=6}
-
-gg_pca_wle = 
-  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
-           groups = as.character(pca_overall_wle$grp$transect), 
-           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
-  geom_point(size=3,stroke=1, alpha = 1,
-             aes(shape = pca_overall_wle$grp$site,
-                 color = groups))+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
-                     values = pal_transect)+
-  xlim(-6,6)+
-  ylim(-4.5,4.5)+
-  labs(shape="",
-       title = "PCA: Erie",
-       subtitle = "surface horizons")+
-  theme_kp()+
-  theme(legend.position = "top", legend.box = "vertical")+
-  NULL
-
-gg_pca_cb = 
-  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
-           groups = as.character(pca_overall_cb$grp$transect), 
-           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
-  geom_point(size=3,stroke=1, alpha = 1,
-             aes(shape = pca_overall_cb$grp$site,
-                 color = groups))+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
-                     values = pal_transect)+
-  xlim(-5.5,7.5)+
-  ylim(-5,5)+
-  labs(shape="",
-       title = "PCA: Chesapeake",
-       subtitle = "surface horizons")+
-  theme_kp()+
-  theme(legend.position = "top", legend.box = "vertical")+
-  NULL
-
-gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
-gg_pca_regions
-
-```
-
-### Drivers and loadings
-
-```{r pc_loadings}
-get_pc_values = function(dat){
-
-  pca = fit_pca_function(dat)
-
-  x = pca$pca_int$rotation
-  x2 = 
-    x %>% 
-    as.data.frame() %>% 
-    dplyr::select(PC1) %>% 
-    arrange(desc(PC1)) %>% 
-    rownames_to_column("variable") %>% 
-    mutate(variable = fct_reorder(variable, (PC1)))
-
-  x2
-}
-
-pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
-pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "Erie")) %>% mutate(region = "Erie")
-pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")
-
-```
-
-```{r pc_gg, fig.height=8, fig.width=6}
-
-pc_all = 
-  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
-  mutate(variable = fct_reorder(variable, (PC1)))
-
-overall = 
-  pc_all %>% 
-  filter(region == "combined") %>% 
-  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
-  ggplot(aes(x = abs(PC1), y = variable))+
-  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
-  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
-  scale_shape_manual(values = c(23, 21))+
-
-  xlim(-0.4, 0.4)+
-  geom_vline(xintercept = 0)+
-  labs(y = "")+
-  theme_kp()
-
-non_absolute = 
-  pc_all %>% 
-  mutate(variable = fct_reorder(variable, (PC1))) %>% 
-  filter(region != "combined") %>% 
-  ggplot(aes(x = PC1, y = variable))+
-  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
-  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
-  scale_shape_manual(values = c(23, 21))+
-
-  xlim(-0.4, 0.4)+
-  geom_vline(xintercept = 0)+
-  labs(y = "")+
-  theme_kp()
-
-pc_all %>% 
-  filter(region != "combined") %>% 
-  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
-  ggplot(aes(x = abs(PC1), y = variable))+
-  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
-  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
-  scale_shape_manual(values = c(23, 21))+
-  labs(y = "",
-       x = "|PC1|")+
-  theme_kp()
-
-```
-
-### Clustering
+### PCA with clusters
 
@@
 
-```{r cluster_gg}
+```{r cluster_gg, eval=FALSE}
 df_cl %>%
@@
 ```
-
-### PCA with clusters
 
@@
 
+```{r pca_gg, fig.width=10, fig.height=6, eval=FALSE}
+## PCA plots overall
+#gg_pca_overall = 
+ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
+         groups = as.character(pca_overall$grp$region), 
+         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=4,stroke=1, alpha = 1,
+             aes(shape = pca_overall$grp$region,
+                 fill = groups))+ 
+  scale_color_manual(values = c("#16879C", "#BB281E"))+
+  scale_fill_manual(values = c("#16879C", "#BB281E"))+
+  scale_shape_manual(values = c(23, 21))+
+  labs(shape="",
+       title = "Overall PCA, both regions",
+       subtitle = "Surface horizons only")+
+  theme_kp()+
+  guides(fill = "none")+
+  NULL
+
+```
+
+```{r pca_gg_regions, fig.width=10, fig.height=6}
+
+gg_pca_wle = 
+  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall_wle$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall_wle$grp$site,
+                 color = groups))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  xlim(-6,6)+
+  ylim(-4.5,4.5)+
+  labs(shape="",
+       title = "PCA: Erie",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+gg_pca_cb = 
+  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
+           groups = as.character(pca_overall_cb$grp$transect), 
+           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_point(size=3,stroke=1, alpha = 1,
+             aes(shape = pca_overall_cb$grp$site,
+                 color = groups))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  xlim(-5.5,7.5)+
+  ylim(-5,5)+
+  labs(shape="",
+       title = "PCA: Chesapeake",
+       subtitle = "surface horizons")+
+  theme_kp()+
+  theme(legend.position = "top", legend.box = "vertical")+
+  NULL
+
+gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
+gg_pca_regions
+
+```
+
+### Drivers and loadings
+
+```{r pc_loadings}
+get_pc_values = function(dat){
+
+  pca = fit_pca_function(dat)
+
+  x = pca$pca_int$rotation
+  x2 = 
+    x %>% 
+    as.data.frame() %>% 
+    dplyr::select(PC1) %>% 
+    arrange(desc(PC1)) %>% 
+    rownames_to_column("variable") %>% 
+    mutate(variable = fct_reorder(variable, (PC1)))
+
+  x2
+}
+
+pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
+pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "Erie")) %>% mutate(region = "Erie")
+pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")
+
+```
+
+```{r pc_gg, fig.height=8, fig.width=6}
+
+pc_all = 
+  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
+  mutate(variable = fct_reorder(variable, (PC1)))
+
+overall = 
+  pc_all %>% 
+  filter(region == "combined") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
+non_absolute = 
+  pc_all %>% 
+  mutate(variable = fct_reorder(variable, (PC1))) %>% 
+  filter(region != "combined") %>% 
+  ggplot(aes(x = PC1, y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+
+  xlim(-0.4, 0.4)+
+  geom_vline(xintercept = 0)+
+  labs(y = "")+
+  theme_kp()
+
+pc_all %>% 
+  filter(region != "combined") %>% 
+  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
+  ggplot(aes(x = abs(PC1), y = variable))+
+  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
+  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
+  scale_shape_manual(values = c(23, 21))+
+  labs(y = "",
+       x = "|PC1|")+
+  theme_kp()
+
+```
+
 ---
@@
 
-### Normalized values - v1
-
-```{r scaled_gg1, fig.height = 8, fig.width = 7}
+```{r scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
 data_long = 
@@
 
-```{r v2-pca_gg, fig.width=10, fig.height=6}
-## PCA plots overall
-#gg_pca_overall = 
-ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
-         groups = as.character(pca_overall$grp$region), 
-         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
+### PCA with clusters
+
+```{r v2-cluster_pca, fig.width=10, fig.height=6}
+library(ggConvexHull)
+
+pca_cluster = fit_pca_function(df_cl)
+
+ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
+                   groups = as.character(pca_cluster$grp$cluster), 
+                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
+  geom_convexhull(aes(group = pca_cluster$grp$cluster,
+                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
   geom_point(size=4,stroke=1, alpha = 1,
-             aes(shape = pca_overall$grp$region,
+             aes(shape = pca_cluster$grp$region,
                  fill = groups))+ 
-  scale_color_manual(values = c("#16879C", "#BB281E"))+
-  scale_fill_manual(values = c("#16879C", "#BB281E"))+
   scale_shape_manual(values = c(23, 21))+
-  labs(shape="",
-       title = "Overall PCA, both regions",
-       subtitle = "Surface horizons only")+
-  theme_kp()+
-  guides(fill = "none")+
+  scale_color_manual(values = pnw_palette("Sailboat", 3))+
+  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
+  labs(fill="cluster",
+       color = "cluster",
+       shape = "region")+
+  theme_kp()+
+  theme(legend.position = c(0.8, 0.75))+
+  guides(fill = guide_legend(override.aes = list(shape = 21)))+
   NULL
@@
 ```
+
+```{r v2-cluster_calc, include=FALSE}
+
+#library(cluster)
+library(factoextra)
+
+df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+```
+
+```{r v2-cluster_gg, include=FALSE}
+df_cl %>%
+  ggplot(aes(y = cluster, x = site))+
+  geom_point(aes(color = transect, shape = transect),
+             size = 4, stroke = 1,
+             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
+                                             dodge.width = 0.4))+
+  scale_color_manual(values = pal_transect)
+
+```
+
+---
+
+## PCA by region
 
@@
 
-```{r v2-pca_gg_regions-poster, fig.width=7, fig.height=7}
+```{r v2-pca_gg_regions-poster, fig.width=7, fig.height=7, eval=FALSE}
 
@@
 
-### Clustering
-
-```{r v2-cluster_calc, include=FALSE}
-
-#library(cluster)
-library(factoextra)
-
-df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
-grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
-
-fviz_nbclust(df, FUN = hcut, method = "wss")
-fviz_nbclust(df, FUN = hcut, method = "silhouette")
-# three clusters
-
-d <- dist(df, method = "euclidean")
-hc1 <- hclust(d, method = "complete" )
-plot(hc1, cex = 0.6, hang = -1)
-hc5 <- hclust(d, method = "ward.D2" )
-sub_grp <- cutree(hc5, k = 3)
-table(sub_grp)
-plot(hc5, cex = 0.6)
-rect.hclust(hc5, k = 3, border = 2:5)
-
-df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
-count(df_cl,cluster)
-
-```
-
-```{r v2-cluster_gg}
-df_cl %>%
-  ggplot(aes(y = cluster, x = site))+
-  geom_point(aes(color = transect, shape = transect),
-             size = 4, stroke = 1,
-             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
-                                             dodge.width = 0.4))+
-  scale_color_manual(values = pal_transect)
-
-```
-
-### PCA with clusters
-
-```{r v2-cluster_pca, fig.width=10, fig.height=6}
-library(ggConvexHull)
-
-pca_cluster = fit_pca_function(df_cl)
-
-ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
-                   groups = as.character(pca_cluster$grp$cluster), 
-                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
-  geom_convexhull(aes(group = pca_cluster$grp$cluster,
-                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
-  geom_point(size=4,stroke=1, alpha = 1,
-             aes(shape = pca_cluster$grp$region,
-                 fill = groups))+ 
-  scale_shape_manual(values = c(23, 21))+
-  scale_color_manual(values = pnw_palette("Sailboat", 3))+
-  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
-  labs(fill="cluster",
-       color = "cluster",
-       shape = "region")+
-  theme_kp()+
-  theme(legend.position = c(0.8, 0.75))+
-  guides(fill = guide_legend(override.aes = list(shape = 21)))+
-  NULL
-
-```
-
----
-
 ## 1b. PERMANOVA
@@
 ## 2. Each analyte
-
-### Normalized values - v1
 
@@
 
-### Normalized values - v2
-
-```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
+### Normalized values
+
+```{r v2-scaled_gg2, fig.height = 8, fig.width = 8, eval=FALSE}
 data_long_all = 
@@
 
-```{r v2-scaled_gg2_split, fig.height = 10, fig.width = 14}
+```{r v2-scaled_gg2_split, fig.height = 10, fig.width = 14, eval=FALSE}
 library(ggh4x) # for nested facets
@@
 ```
+
+### only select analytes
 
@@
 
-### GWI only
-
-```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8}
+```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8, eval=FALSE}
 data_scaled_NEW %>% 
@@
 
-```{r analytesx-4, fig.height = 15, fig.width = 8}
+```{r analytesx-4, fig.height = 15, fig.width = 8, eval=FALSE}
 cowplot::plot_grid(gg_tc, gg_fe, gg_p, ncol = 1)

### 2025-07-28 2-code/250618_data_for_kp_synoptic_wl_vwc.R
synoptic water level

@@ -1,0 +1,224 @@
+## This script pulls data from 2021-2023 for 6 (not SWH) synoptic sites, looking
+## for a full year (ish) of data across our sites, and calculating 
+## 1) the % of time that each site-location is under water, and 
+## 2) the average VWC at 5 cm
+##
+## Peter Regier
+## 2025-06-18
+##
+# ########### #
+# ########### #
+
+
+# 1. Setup ---------------------------------------------------------------------
+
+## Load packages with pacman::p_load
+require(pacman)
+p_load(tidyverse,
+       janitor,
+       parsedate,
+       lubridate,
+       cowplot, 
+       purrr,
+       readr, 
+       tictoc,
+       beepr,
+       googledrive, 
+       furrr,
+       PNWColors,
+       googlesheets4)
+
+# Set ggplot theme
+theme_set(theme_bw())
+
+common_tz = "Etc/GMT+5"
+
+
+sites_to_pull = "PTR|CRC|OWC|MSM|GCW|GWI"
+years_to_pull = "2022|2023"
+
+temp_storage = "data/raw_from_gdrive/synoptic_kaizad/"
+
+
+# 2. Pull L1 data --------------------------------------------------------------
+
+## We are pulling lots of data, let's figure out how to do that programmatically
+l1_folders <- drive_ls("https://drive.google.com/drive/folders/1sDB3Q3ZXGvXkniWw03CHY1E0KV5GdN8j") %>% 
+  filter(grepl(sites_to_pull, name)) %>% 
+  filter(grepl(years_to_pull, name)) #%>% 
+  #slice(3:n())
+
+
+## Make functions to find and download needed csvs
+download_l1_csvs <- function(i){
+  
+  folder <- l1_folders %>% slice(i)
+  
+  folder_name <- folder$name
+  
+  message(paste("downloading data from", folder_name))
+  
+  folder_id <- folder$id
+  
+  y <- googledrive::drive_ls(folder_id, pattern = ".csv$") %>% 
+    filter(!(grepl("_OW_", name)))
+  
+  # for(i in 1:length(y$id)){
+  # #for(i in 1:1){
+  #   googledrive::drive_download(y$id[[i]],
+  #                               overwrite = TRUE,
+  #                               path = paste0(temp_storage, "raw_files/", y$name[[i]]))
+  # }
+  
+  # Download files only if they aren't already present
+  for (j in seq_along(y$id)) {
[…]

### 2025-07-28 3-reports/manuscript_figures.Rmd
add NH4N to figures

@@
 
-```{r v2-scaled_gg2_split-NEW, fig.height = 10, fig.width = 12}
+```{r v2-scaled_gg2_split-NEW, fig.height = 13, fig.width = 12}
 library(ggh4x) # for nested facets
@@
   data_scaled_NEW %>% 
-  filter(name %in% c("WEOC", "TS", "Na", "Chloride", "Fe", "P")) %>% 
+  filter(name %in% c("WEOC", "TS", "Na", "Chloride", "Fe", "P", "NH4N")) %>% 
     mutate(name = factor(name, levels = c("TC", "TS", "WEOC",
                                         "Ca", "Na", "Chloride",
-                                        "NH4N", "Fe", "P")))
+                                        "Fe", "P", "NH4N")))
 

### 2025-08-12 3-reports/manuscript_figures.Rmd
fix figure sizes

@@
 library(ggbiplot)
+library(targets)
 
@@
   theme_kp()+
-  theme(legend.position = "top", legend.box = "vertical")+
+  theme(legend.position = "top", legend.box = "vertical",
+        axis.text = element_text(size = 16, color = "black"),
+        axis.title = element_text(size = 16, face = "bold", color = "black"),
+  )+
   guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
@@
   scale_x_discrete(labels = c("U", "T" , "W" ))+
-  labs(x = "Transect", y = "Normalized concentration")+
-  theme_kp()+
-  theme(axis.text.y = element_blank(),
-        panel.grid = element_blank())
+  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
+  labs(x = "Transect", y = "Scaled concentration
+       ")+
+  theme_kp()+
+  theme(#axis.text.y = element_blank(),
+        panel.grid = element_blank()
+        )
 
@@
   scale_x_discrete(labels = c("U", "T" , "W" ))+
-  labs(x = "Transect", y = "Normalized concentration")+
+  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
+  labs(x = "Transect", y = "Scaled concentration
+       ")+
   theme(axis.text.y = element_blank(),

### 2025-08-12 2-code/250618_data_for_kp_synoptic_wl_vwc.R
soil moisture sensors -- functions

@@ -37,7 +37,7 @@ common_tz = "Etc/GMT+5"
 sites_to_pull = "PTR|CRC|OWC|MSM|GCW|GWI"
 years_to_pull = "2022|2023"
 
-temp_storage = "data/raw_from_gdrive/synoptic_kaizad/"
+temp_storage = "TEMP/"
 
 
 # 2. Pull L1 data --------------------------------------------------------------
@@ -221,4 +221,46 @@ percent_flooded <- troll %>%
 
 write_csv(percent_flooded, paste0(temp_storage, "/250620_kaizad_synoptic_troll_percent_flooded.csv"))
 
+
+troll_outlier = 
+  troll %>% 
+  arrange(site, plot, datetime_est) %>% 
+  group_by(site, plot) %>% 
+  mutate(rollmean = zoo::rollmean(wl_below_surface_m, 70, fill = NA),
+         lag = - wl_below_surface_m + lag(wl_below_surface_m),
+         outlier = case_when(abs(lag) >= 0.3 ~ "outlier"),
+         outlier2 = case_when(abs(wl_below_surface_m - rollmean) > 1 ~ "outlier"),
+         plot = factor(plot, levels = c("UP", "TR", "W")),
+         region = case_when(site %in% c("GCW", "MSM", "GWI") ~ "Chesapeake",
+                            site %in% c("CRC", "PTR", "OWC") ~ "Erie"))
+  
+##  group_by(site, plot) %>% 
+##  dplyr::summarise(mean_wl = mean(wl_below_surface_m),
+##                   median_wl = median(wl_below_surface_m),
+##                   min_wl = min(wl_below_surface_m),
+##                   max_wl = max(wl_below_surface_m))
+
+
+troll_outlier %>% 
+  filter(!(site == "GCW" & plot == "UP")) %>% 
+  ggplot(aes(x = datetime_est, y = wl_below_surface_m))+
+  geom_point(aes(color = outlier2))+
+  geom_line(aes(y = rollmean))+
+  facet_grid(site ~ plot)
+
+library(ggh4x) # for nested facets
+troll_outlier %>% 
+  #filter(!site == "GCW") %>%
+  mutate(rollmean = case_when(site == "GCW" ~ NA, .default = rollmean)) %>% 
+  filter(is.na(outlier2)) %>% 
+  ggplot(aes(x = datetime_est, y = wl_below_surface_m, color = plot))+
[…]

### 2025-08-12 2-code/c-functions_soil_moisture_sensors.R

@@ -1,0 +1,209 @@
+
+
+require(pacman)
+p_load(tidyverse,
+       janitor,
+       parsedate,
+       lubridate,
+       cowplot, 
+       purrr,
+       readr, 
+       tictoc,
+       beepr,
+       googledrive, 
+       furrr,
+       PNWColors,
+       googlesheets4)
+
+
+
+SENSOR_PATH = "TEMP/csvs_to_process"
+
+## read in TEROS/TROLL data. gives df_trim
+process_teros = function(SENSOR_PATH){
+  
+  to_read <- list.files(SENSOR_PATH, full.names = T)
+  
+  df_raw <- to_read %>% 
+    map(read_csv) %>% 
+    bind_rows() %>% 
+    dplyr::select(-instrument_id)
+  
+  ## Summary of start dates gives 3/22 to 4/22 coinciding with synoptic installs
+  ## I'm setting 4/22 - 4/23 as a consistent(ish) time-frame for all sites
+  df_trim <- df_raw %>% 
+    filter(datetime_est > "2022-04-01") %>% 
+    filter(datetime_est < "2023-04-01")
+  
+  ## Monster 68M dataset, free up memory
+  rm(df_raw)
+  
+  df_trim %>% 
+    rename(transect = plot) %>% 
+    mutate(region = case_when(site %in% c("GCW", "MSM", "GWI") ~ "Chesapeake",
+                       site %in% c("CRC", "PTR", "OWC") ~ "Erie"),
+           transect = case_match(transect, "UP" ~ "upland", "TR" ~ "transition", "W" ~ "wetland"))
+}
+df_trim = process_teros(SENSOR_PATH)
+
+
+
+# summarize VWC means
+get_vwc_data = function(df_trim){
+  
+  tic("bin teros") # only 5s!
+  vwc_data <- df_trim %>% 
+    filter(research_name == "soil_vwc_10cm") %>% 
+    group_by(datetime_est, region, site, transect) %>% 
+    summarize(value = mean(value, na.rm = T),
+              value = value * 100) %>% 
+    drop_na() 
+    
+  toc()
+  vwc_data
+}
+vwc_data = get_vwc_data(df_trim)
+
+summarize_vwc = function(vwc_data){
+
+  vwc_means <- 
+    vwc_data %>% 
+    ungroup() %>% 
+    group_by(region, site, transect) %>% 
+    summarize(mean_vwc = round(mean(value, na.rm = T), 2),
[…]

### 2025-08-12 3-reports/manuscript_figures.Rmd
add soil moisture sensors to targets and Rmd

@@
 library(targets)
+library(ggh4x)
 
@@
 targets::tar_load(c(data_combined, sample_key))
+targets::tar_load_everything()
 ```
@@
 
+## SOIL MOISTURE SENSORS
+
+Processed data are available here in this repository. Raw data are available on Google Drive (access needed).
+
+### Soil water content for 2022-23
+
+```{r gg_vwc, fig.height=8, fig.width=10}
+# plot vwc
+
+vwc_data %>% 
+  filter(value > 0) %>% 
+  reorder_site() %>% 
+  reorder_transect() %>% 
+  ggplot(aes(x = datetime_est, y = value, color = transect))+
+  geom_line()+
+  #  facet_wrap(~ site + transect)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey")) +
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  labs(x = "", 
+       y = "Water content (% v/v)",
+       color = "")+
+  theme_kp()+
+  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
+        panel.grid = element_blank())
+
+```
+
+### Summarized soil water content
+
+```{r}
+vwc_means %>% knitr::kable()
+```
+
+### Water level in soil
+
+When water level is positive (above surface), it means the soil is flooded.
+
+```{r gg_water_level, fig.height=8, fig.width=10}
+
+troll %>% 
+  ggplot(aes(x = datetime_est, y = wl_below_surface_m, color = transect))+
+  geom_hline(yintercept = 0, alpha = 0.8)+
+  geom_line(aes(y = rollmean), linewidth = 1)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey"))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+                     values = pal_transect)+
+  labs(x = "",
+       y = "Water level below surface (m)",
+       color = "")+
+  theme_kp()+
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(size = 14, angle = 45, hjust = 1))   ##
+
+```
+
+### Water table summary
+
+```{r}
+water_table %>% knitr::kable()
+```
+
+### Calculating time flooded (% of the year)
+
+```{r}
+percent_flooded %>% knitr::kable()
+```
+
+---
+
 ## Session Info

### 2025-08-12 2-code/c-functions_soil_moisture_sensors.R

@@ -17,9 +17,7 @@ p_load(tidyverse,
 
 
 
-SENSOR_PATH = "TEMP/csvs_to_process"
-
-## read in TEROS/TROLL data. gives df_trim
+## read in TEROS/TROLL data. gives sensor_df_trim
 process_teros = function(SENSOR_PATH){
   
   to_read <- list.files(SENSOR_PATH, full.names = T)
@@ -31,28 +29,25 @@ process_teros = function(SENSOR_PATH){
   
   ## Summary of start dates gives 3/22 to 4/22 coinciding with synoptic installs
   ## I'm setting 4/22 - 4/23 as a consistent(ish) time-frame for all sites
-  df_trim <- df_raw %>% 
+  sensor_df_trim <- df_raw %>% 
     filter(datetime_est > "2022-04-01") %>% 
     filter(datetime_est < "2023-04-01")
   
   ## Monster 68M dataset, free up memory
   rm(df_raw)
   
-  df_trim %>% 
+  sensor_df_trim %>% 
     rename(transect = plot) %>% 
     mutate(region = case_when(site %in% c("GCW", "MSM", "GWI") ~ "Chesapeake",
                        site %in% c("CRC", "PTR", "OWC") ~ "Erie"),
            transect = case_match(transect, "UP" ~ "upland", "TR" ~ "transition", "W" ~ "wetland"))
 }
-df_trim = process_teros(SENSOR_PATH)
-
-
 
 # summarize VWC means
-get_vwc_data = function(df_trim){
+get_vwc_data = function(sensor_df_trim){
   
   tic("bin teros") # only 5s!
-  vwc_data <- df_trim %>% 
+  vwc_data <- sensor_df_trim %>% 
     filter(research_name == "soil_vwc_10cm") %>% 
     group_by(datetime_est, region, site, transect) %>% 
     summarize(value = mean(value, na.rm = T),
@@ -62,7 +57,6 @@ get_vwc_data = function(df_trim){
   toc()
   vwc_data
 }
-vwc_data = get_vwc_data(df_trim)
 
 summarize_vwc = function(vwc_data){
 
@@ -71,41 +65,20 @@ summarize_vwc = function(vwc_data){
     ungroup() %>% 
     group_by(region, site, transect) %>% 
     summarize(mean_vwc = round(mean(value, na.rm = T), 2),
-              sd_vwc = round(sd(value, na.rm = T), 2)) %>% 
[…]

### 2025-08-12 _targets.R

@@ -15,8 +15,9 @@ tar_option_set(
 source("2-code/0-packages.R")
 #source("2-code/1-initial_processing.R")
 source("2-code/a-functions_processing.R")
-source("2-code/b-functions_analysis.R")
-source("2-code/c-functions_fticrrr.R")
+#source("2-code/b-functions_analysis.R")
+#source("2-code/c-functions_fticrrr.R")
+source("2-code/c-functions_soil_moisture_sensors.R")
 
 # Replace the target list below with your own:
 list(
@@ -82,13 +83,21 @@ list(
     write.csv(data_combined, "1-data/processed/chemistry_combined_surface_horizon.csv", row.names = FALSE)
     write.csv(data_combined_all_horizons, "1-data/processed/chemistry_combined_all_horizons.csv", row.names = FALSE)
     write.csv(data_combined_wide, "1-data/processed/chemistry_combined_wide.csv", row.names = FALSE)
-# texture
-# PD
-# BD
-# WRC
+    # texture
+    # PD
+    # BD
+    # WRC
   }, format = "file"),
-   
-   
+  
+  
+  # soil moisture sensors
+  tar_target(sensor_df_trim, process_teros(SENSOR_PATH = "TEMP/csvs_to_process")),
+  tar_target(vwc_data, get_vwc_data(sensor_df_trim)),
+  tar_target(vwc_means, summarize_vwc(vwc_data)),
+  tar_target(troll, calculate_water_levels(sensor_df_trim)),
+  tar_target(percent_flooded, calculate_percent_flooded(troll)),
+  tar_target(water_table, summarize_water_table(troll)),
+
    # report  
 #   tar_render(report, path = "3-reports/characterization_report.Rmd")
    tar_render(report, path = "3-reports/manuscript_figures.Rmd")

### 2025-08-13 2-code/x-old/pr_ml/240804_initial_RF.Rmd
reorganize old script files

--- 2024-08-06 pr_ml/240804_initial_RF.Rmd
@@
 2.  I'm a little suspicious that the models performed "too" good... I was expecting them to get maybe half of the values right, 100% makes me suspicious.
+
+#### 240808 - adding on: how do models for transect location vary between regions? {.tabset}
+
+```{r set up wrapper plotting function}
+
+make_rf_model_by_subset <- function(var, data){
+
+  # 1) Dynamically create formula
+  formula <- as.formula(paste(var, "~ ."))
+
+  # 2) Tune mtry and ntree hyperparameters
+  rf_spec <- rand_forest() %>%
+    set_engine("ranger", importance = "impurity") %>% #using the ranger pkg instead of rf; importance needed for FI
+    set_mode("classification")
+
+  # 3) Define a recipe
+  recipe <- recipe(formula, data = data) %>%
+    step_integer(all_nominal_predictors()) %>% # integerize all dbl
+    step_corr(all_predictors()) %>% # check correlations between predictors and remove any strong correlations
+    step_normalize(all_predictors(), -all_outcomes()) # normalize non-normally distributed predictors
+
+  # 4) Define a workflow
+  wf <- workflow() %>%
+    add_model(rf_spec) %>%
+    add_recipe(recipe)
+
+  # 5) Run the  model
+  rf_model <- wf %>%
+    update_model(rf_spec) %>%
+    fit(data = data)
+
+  ## Extract predictions
+  predictions <- predict(rf_model, data) %>% rename("predicted" = .pred_class)
+
+  ### Feature Importance
+  ## Set vectors for renaming stuff
+  var_names <- rf_model$fit$fit$fit$variable.importance
+
+  ## Convert feature importance to a tibble with variables as a column
+  fi0 <- as.data.frame(var_names) %>%
+    tibble::rownames_to_column() %>%
+    as_tibble() %>% 
+    rename("predictor" = rowname, 
+           "raw_fi" = var_names) %>% 
+    mutate(fi_n = raw_fi / sum(raw_fi))
+
+  ## Bind predictions to data
+  prediction_df <- bind_cols(data, predictions)
+
+  ## Return two things: how good the model did and which variables are most important
+  return(list(predictions = prediction_df,
+                feature_importance = fi0))
+}
+
+# Wrapper function for plotting
+plot_rf_results_by_subset <- function(var, data) {
+
+  # Call the make_rf_model function
+  model_results <- make_rf_model_by_subset(var, data)
+
+  ## Plot for classification equivalent to goodness-of-fit
+  p1 <- ggplot(model_results$predictions, aes_string(x = var, y = "predicted")) + 
+    geom_jitter(width = 0.2, 
+                height = 0.2,
+                alpha = 0.5) + 
+    ggtitle(paste("Predictions:", var))
+
+  ## Set up a color palette
+  var_colors <- PNWColors::pnw_palette("Bay", n = length(unique(model_results$feature_importance$predictor)))
+
+  ## Plot of feature importance
+  p2 <- ggplot(model_results$feature_importance, aes(x = fi_n * 100, 
+                                                     y = reorder(predictor, fi_n), 
+                                                     fill = predictor)) + 
+    geom_col(alpha = 0.8, show.legend = F, width = 0.7) + 
+    scale_fill_manual(values = var_colors) + 
+    labs(x = "Feature Importance (%)", 
+         y = "", fill = "") + 
+    ggtitle(paste("Feature Imp.:", var))
+
+  ## Combine into a plot
+  plot_grid(p1, p2, nrow = 1)
+}
+
+```
+
+##### Transect location for CB
+
+```{r}
+plot_rf_results_by_subset("transect", df %>% filter(region == "CB"))
+```
+
+##### Transect location for WLE
+
+```{r}
+plot_rf_results_by_subset("transect", df %>% filter(region == "WLE"))
+```
+
+```{r}
+
+both_fi_by_region <- bind_rows(make_rf_model_by_subset("transect", df %>% filter(region == "CB"))$feature_importance %>% mutate(region = "CB"), 
+          make_rf_model_by_subset("transect", df %>% filter(region == "WLE"))$feature_importance %>% mutate(region = "WLE"))
+
+both_fi_by_region_trim <- both_fi_by_region %>% 
+  group_by(region) %>%
+  filter(fi_n > (fi_n[predictor == "white_noise"])) %>%
+  ungroup()
+
+var_colors <- PNWColors::pnw_palette("Bay", n = length(unique(both_fi_by_region_trim$predictor)))
+
+both_fi_by_region_trim %>% 
+  mutate(fi = fi_n * 100) %>% 
+  ggplot(aes(x = fi_n * 100, 
+             y = reorder(predictor, fi_n), 
+             fill = predictor)) + 
+  geom_col(alpha = 0.8, show.legend = F, width = 0.7) + 
+  scale_fill_manual(values = var_colors) + 
+  facet_wrap(~region, nrow = 1) +
+  labs(x = "Feature Importance (%)", 
+       y = "", fill = "") 
+```

### 2025-08-13 2-code/3a-functions_soil_moisture_sensors.R

@@ -68,7 +68,8 @@ summarize_vwc = function(vwc_data){
               sd_vwc = round(sd(value, na.rm = T), 2),
               median_vwc = round(median(value, na.rm = T), 2)) %>% 
     reorder_site() %>% 
-    reorder_transect()
+    reorder_transect() %>% 
+    arrange(site, transect)
   
 } 
 
@@ -154,6 +155,7 @@ summarize_water_table = function(troll){
   troll %>% 
       group_by(region, site, transect) %>% 
       dplyr::summarise(mean_wl = mean(wl_below_surface_m),
+                       sd_wl = sd(wl_below_surface_m),
                        median_wl = median(wl_below_surface_m),
                        min_wl = min(wl_below_surface_m),
                        max_wl = max(wl_below_surface_m)) %>% 

### 2025-08-13 _targets.R

@@ -14,10 +14,8 @@ tar_option_set(
 # Load the R scripts with your custom functions:
 source("2-code/0-packages.R")
 #source("2-code/1-initial_processing.R")
-source("2-code/a-functions_processing.R")
-#source("2-code/b-functions_analysis.R")
-#source("2-code/c-functions_fticrrr.R")
-source("2-code/c-functions_soil_moisture_sensors.R")
+source("2-code/2-functions_processing.R")
+source("2-code/3a-functions_soil_moisture_sensors.R")
 
 # Replace the target list below with your own:
 list(

### 2025-08-13 3-reports/manuscript_figures.Rmd
update elevation script

@@
   filter(region == "Erie") %>% 
-  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect))+
-  geom_line(color = "black")+
+  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect, linetype = site))+
+  geom_line(color = "black", show.legend = F)+
   geom_point(size = 4)+
@@
                      breaks = c("water edge", "wetland", "transition", "upland"))+
+  scale_linetype_manual(values = c("dashed", "solid", "solid"),
+                     breaks = c("CRC", "PTR", "OWC"))+
   facet_nested_wrap(region ~ site)+

### 2025-08-13 _targets.R
update file path for sensor data

@@ -89,7 +89,7 @@ list(
   
   
   # soil moisture sensors
-  tar_target(sensor_df_trim, process_teros(SENSOR_PATH = "TEMP/csvs_to_process")),
+  tar_target(sensor_df_trim, process_teros(SENSOR_PATH = "1-data/soil_moisture_sensors/csvs_to_process")),
   tar_target(vwc_data, get_vwc_data(sensor_df_trim)),
   tar_target(vwc_means, summarize_vwc(vwc_data)),
   tar_target(troll, calculate_water_levels(sensor_df_trim)),

### 2025-09-08 3-reports/manuscript_figures.Rmd
update PCA
relabel wetland --> marsh

--- 2025-08-12 3-reports/manuscript_figures.Rmd
@@
 targets::tar_load(c(data_combined, sample_key))
+targets::tar_load_everything()
+```
+
+```{r}
+reorder_transect = function(dat){
+  dat %>% 
+    mutate(transect = case_match(transect, "wetland" ~ "marsh", .default = transect),
+           transect = factor(transect, levels = c("upland", "transition", "wte", "marsh")))
+}
+
+# update transect names in sample_key
+sample_key = 
+  sample_key %>% 
+  reorder_transect()
+
 ```
@@
     data_combined %>% 
+    reorder_transect() %>% 
     filter(!transect %in% "wte") %>% 
@@
     dat %>% 
-    dplyr::select(where(is.character)) %>% 
+    dplyr::select(where(is.character) | where(is.factor)) %>% 
     dplyr::mutate(row = row_number()) %>% 
@@
 ```{r pca_input}
-pca_overall = fit_pca_function(data_wide_PCA) 
+pca_overall = fit_pca_function(dat = data_wide_PCA) 
 pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "Erie"))
@@
                  color = groups))+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                      values = pal_transect)+
@@
   pivot_wider(names_from = "transect", values_from = "scaled") %>% 
-  mutate(correction = case_when(upland < wetland ~ 1,
-                                upland > wetland ~ -1)) %>% 
-  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
+  mutate(correction = case_when(upland < marsh ~ 1,
+                                upland > marsh ~ -1)) %>% 
+  pivot_longer(cols = c(upland, marsh, transition), names_to = "transect", values_to = "scaled") %>% 
   mutate(scaled = scaled * correction,
          scaled = case_when(transect == "upland" ~ 0,
-                            transect == "wetland" ~ 1,
+                            transect == "marsh" ~ 1,
                             TRUE ~ scaled),
@@
   annotate("text", label = "U", x = 0, y = 21)+
-  annotate("text", label = "W", x = 1, y = 21)+
+  annotate("text", label = "M", x = 1, y = 21)+
   annotate("text", label = "", x = 1, y = 21.5)+
@@
                                 "#74a9cf", "#2b8cbe", "#045a8d"))+
-  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  scale_x_discrete(labels = c("U", "T" , "M" ))+
   labs(x = "Transect", y = "Normalized concentration")+
@@
 ### PCA with clusters
+
+```{r v2-cluster_calc, include=FALSE}
+
+#library(cluster)
+library(factoextra)
+
+df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
+grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
+
+fviz_nbclust(df, FUN = hcut, method = "wss")
+fviz_nbclust(df, FUN = hcut, method = "silhouette")
+# three clusters
+
+d <- dist(df, method = "euclidean")
+hc1 <- hclust(d, method = "complete" )
+plot(hc1, cex = 0.6, hang = -1)
+hc5 <- hclust(d, method = "ward.D2" )
+sub_grp <- cutree(hc5, k = 3)
+table(sub_grp)
+plot(hc5, cex = 0.6)
+rect.hclust(hc5, k = 3, border = 2:5)
+
+df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
+count(df_cl,cluster)
+
+```
+
+```{r v2-cluster_gg, include=FALSE}
+df_cl %>%
+  ggplot(aes(y = cluster, x = site))+
+  geom_point(aes(color = transect, shape = transect),
+             size = 4, stroke = 1,
+             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
+                                             dodge.width = 0.4))+
+  scale_color_manual(values = pal_transect)
+
+```
 
@@
 
-```{r v2-cluster_calc, include=FALSE}
-
-#library(cluster)
-library(factoextra)
-
-df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
-grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))
-
-fviz_nbclust(df, FUN = hcut, method = "wss")
-fviz_nbclust(df, FUN = hcut, method = "silhouette")
-# three clusters
-
-d <- dist(df, method = "euclidean")
-hc1 <- hclust(d, method = "complete" )
-plot(hc1, cex = 0.6, hang = -1)
-hc5 <- hclust(d, method = "ward.D2" )
-sub_grp <- cutree(hc5, k = 3)
-table(sub_grp)
-plot(hc5, cex = 0.6)
-rect.hclust(hc5, k = 3, border = 2:5)
-
-df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
-count(df_cl,cluster)
-
-```
-
-```{r v2-cluster_gg, include=FALSE}
-df_cl %>%
-  ggplot(aes(y = cluster, x = site))+
-  geom_point(aes(color = transect, shape = transect),
-             size = 4, stroke = 1,
-             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
-                                             dodge.width = 0.4))+
-  scale_color_manual(values = pal_transect)
-
-```
-
 ---
@@
              stroke = 1)+ 
-  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                      values = pal_transect)+
-  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
+  scale_fill_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
@@
                      values = c(21, 23, 24, 21, 23, 24))+
-  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  scale_x_discrete(labels = c("U", "T" , "M" ))+
   scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
@@
                                 "#74a9cf", "#2b8cbe", "#045a8d"))+
-  scale_x_discrete(labels = c("U", "T" , "W" ))+
+  scale_x_discrete(labels = c("U", "T" , "M" ))+
   scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
@@
 data_combined_wide %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
   ggplot(aes(x = transect, y = Fe))+
@@
 data_combined_wide %>% 
-  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
+  mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
   ggplot(aes(x = transect, y = P))+
@@
 
-  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
+  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
     filter(!is.na(value))
@@
   reorder_site() %>% 
+  reorder_transect() %>% 
     filter(kpa_fit >= 0 | kpa_eval >= 0) %>% 
@@
   elevation %>% 
-  mutate(transect = factor(transect, levels = c("water edge", "wetland", "transition", "upland"))) %>% 
+  mutate(transect = case_match(transect, "wetland" ~ "marsh", .default = transect),
+         transect = factor(transect, levels = c("water edge", "marsh", "transition", "upland"))) %>% 
   reorder_site()
@@
   filter(region == "Erie") %>% 
-  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect))+
-  geom_line(color = "black")+
+  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect, linetype = site))+
+  geom_line(color = "black", show.legend = F)+
   geom_point(size = 4)+
   scale_color_manual(values = c("grey", "#2E5894", "#FFB200", "#96001B"),
-                     breaks = c("water edge", "wetland", "transition", "upland"))+
+                     breaks = c("water edge", "marsh", "transition", "upland"))+
+  scale_linetype_manual(values = c("dashed", "solid", "solid"),
+                     breaks = c("CRC", "PTR", "OWC"))+
   facet_nested_wrap(region ~ site)+
@@
   scale_color_manual(values = c("grey", "#2E5894", "#FFB200", "#96001B"),
-                     breaks = c("water edge", "wetland", "transition", "upland"))+
+                     breaks = c("water edge", "marsh", "transition", "upland"))+
   facet_nested_wrap(region ~ site)+
@@
 
+## SOIL MOISTURE SENSORS
+
+Processed data are available here in this repository. Raw data are available on Google Drive (access needed).
+
+### Soil water content for 2022-23
+
+```{r gg_vwc, fig.height=8, fig.width=10}
+# plot vwc
+
+vwc_data %>% 
+  filter(value > 0) %>% 
+  reorder_site() %>% 
+  reorder_transect() %>% 
+  ggplot(aes(x = datetime_est, y = value, color = transect))+
+  geom_line()+
+  #  facet_wrap(~ site + transect)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey")) +
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
+                     values = pal_transect)+
+  labs(x = "", 
+       y = "Water content (% v/v)",
+       color = "")+
+  theme_kp()+
+  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
+        panel.grid = element_blank())
+
+```
+
+### Summarized soil water content
+
+```{r}
+vwc_means %>% reorder_transect() %>% knitr::kable()
+```
+
+### Water level in soil
+
+When water level is positive (above surface), it means the soil is flooded.
+
+```{r gg_water_level, fig.height=8, fig.width=10}
+
+troll %>% 
+  reorder_transect() %>% 
+  ggplot(aes(x = datetime_est, y = wl_below_surface_m, color = transect))+
+  geom_hline(yintercept = 0, alpha = 0.8)+
+  geom_line(aes(y = rollmean), linewidth = 1)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey"))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
+                     values = pal_transect)+
+  labs(x = "",
+       y = "Water level below surface (m)",
+       color = "")+
+  theme_kp()+
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(size = 14, angle = 45, hjust = 1))   ##
+
+```
+
+### Water table summary
+
+```{r}
+water_table %>% reorder_transect() %>% knitr::kable()
+```
+
+### Calculating time flooded (% of the year)
+
+```{r}
+percent_flooded %>% reorder_transect() %>% knitr::kable()
+```
+
+---
+
 ## Session Info

### 2025-09-14 2-code/3a-functions_soil_moisture_sensors.R
clean up sensor data script. download L2 data, change date ranges.

@@ -29,18 +29,21 @@ process_teros = function(SENSOR_PATH){
   
   ## Summary of start dates gives 3/22 to 4/22 coinciding with synoptic installs
   ## I'm setting 4/22 - 4/23 as a consistent(ish) time-frame for all sites
-  sensor_df_trim <- df_raw %>% 
-    filter(datetime_est > "2022-04-01") %>% 
-    filter(datetime_est < "2023-04-01")
+  sensor_df_trim1 <- df_raw %>% 
+    rename(transect = plot) %>% 
+  #  filter(datetime_est > "2023-01-01") %>% 
+  #  filter(datetime_est < "2023-12-31") %>% 
+    force()
   
   ## Monster 68M dataset, free up memory
-  rm(df_raw)
+#  rm(df_raw)
   
-  sensor_df_trim %>% 
-    rename(transect = plot) %>% 
+  sensor_df_trim = sensor_df_trim1 %>% 
     mutate(region = case_when(site %in% c("GCW", "MSM", "GWI") ~ "Chesapeake",
                        site %in% c("CRC", "PTR", "OWC") ~ "Erie"),
-           transect = case_match(transect, "UP" ~ "upland", "TR" ~ "transition", "W" ~ "wetland"))
+           transect = case_match(transect, "UP" ~ "upland", "TR" ~ "transition", "W" ~ "wetland")) %>% 
+    filter(!is.na(transect)) %>% 
+    filter(!is.na(value))
 }
 
 # summarize VWC means
@@ -48,7 +51,7 @@ get_vwc_data = function(sensor_df_trim){
   
   tic("bin teros") # only 5s!
   vwc_data <- sensor_df_trim %>% 
-    filter(research_name == "soil_vwc_10cm") %>% 
+    filter(research_name == "soil-vwc-10cm") %>% 
     group_by(datetime_est, region, site, transect) %>% 
     summarize(value = mean(value, na.rm = T),
               value = value * 100) %>% 
@@ -62,6 +65,8 @@ summarize_vwc = function(vwc_data){
 
   vwc_means <- 
     vwc_data %>% 
+    filter(datetime_est > "2023-01-01") %>% 
+    filter(datetime_est < "2023-12-31") %>% 
     ungroup() %>% 
     group_by(region, site, transect) %>% 
     summarize(mean_vwc = round(mean(value, na.rm = T), 2),
[…]

### 2025-09-14 2-code/3b-250618_data_for_kp_synoptic_wl_vwc.R

@@ -1,266 +1,0 @@
-## This script pulls data from 2021-2023 for 6 (not SWH) synoptic sites, looking
-## for a full year (ish) of data across our sites, and calculating 
-## 1) the % of time that each site-location is under water, and 
-## 2) the average VWC at 5 cm
-##
-## Peter Regier
-## 2025-06-18
-##
-# ########### #
-# ########### #
-
-
-# 1. Setup ---------------------------------------------------------------------
-
-## Load packages with pacman::p_load
-require(pacman)
-p_load(tidyverse,
-       janitor,
-       parsedate,
-       lubridate,
-       cowplot, 
-       purrr,
-       readr, 
-       tictoc,
-       beepr,
-       googledrive, 
-       furrr,
-       PNWColors,
-       googlesheets4)
-
-# Set ggplot theme
-theme_set(theme_bw())
-
-common_tz = "Etc/GMT+5"
-
-
-sites_to_pull = "PTR|CRC|OWC|MSM|GCW|GWI"
-years_to_pull = "2022|2023"
-
-temp_storage = "TEMP/"
-
-
-# 2. Pull L1 data --------------------------------------------------------------
-
-## We are pulling lots of data, let's figure out how to do that programmatically
-l1_folders <- drive_ls("https://drive.google.com/drive/folders/1sDB3Q3ZXGvXkniWw03CHY1E0KV5GdN8j") %>% 
-  filter(grepl(sites_to_pull, name)) %>% 
-  filter(grepl(years_to_pull, name)) #%>% 
-  #slice(3:n())
-
-
-## Make functions to find and download needed csvs
-download_l1_csvs <- function(i){
-  
-  folder <- l1_folders %>% slice(i)
-  
-  folder_name <- folder$name
-  
-  message(paste("downloading data from", folder_name))
-  
-  folder_id <- folder$id
-  
-  y <- googledrive::drive_ls(folder_id, pattern = ".csv$") %>% 
-    filter(!(grepl("_OW_", name)))
-  
-  # for(i in 1:length(y$id)){
-  # #for(i in 1:1){
-  #   googledrive::drive_download(y$id[[i]],
-  #                               overwrite = TRUE,
-  #                               path = paste0(temp_storage, "raw_files/", y$name[[i]]))
-  # }
-  
-  # Download files only if they aren't already present
-  for (j in seq_along(y$id)) {
[…]

### 2025-09-14 2-code/3b-download_soil_moisture_sensor_data.R

@@ -1,0 +1,154 @@
+
+
+## This script pulls data from 2021-2023 for 6 (not SWH) synoptic sites, looking
+## for a full year (ish) of data across our sites, and calculating 
+## 1) the % of time that each site-location is under water, and 
+## 2) the average VWC at 5 cm
+##
+## Peter Regier
+## 2025-06-18
+##
+# ########### #
+# ########### #
+
+
+# 1. Setup ---------------------------------------------------------------------
+
+## Load packages with pacman::p_load
+require(pacman)
+p_load(tidyverse,
+       janitor,
+       parsedate,
+       lubridate,
+       cowplot, 
+       purrr,
+       readr, 
+       tictoc,
+       beepr,
+       googledrive, 
+       furrr,
+       PNWColors,
+       googlesheets4)
+
+# Set ggplot theme
+theme_set(theme_bw())
+
+common_tz = "Etc/GMT+5"
+
+
+sites_to_pull = "GWI|GCW|MSM|PTR|CRC|OWC"
+years_to_pull = "2022|2023|2024"
+
+temp_storage = "TEMP/"
+
+
+# 2. Pull L1 data --------------------------------------------------------------
+
+## We are pulling lots of data, let's figure out how to do that programmatically
+l2_folders <- drive_ls("https://drive.google.com/drive/folders/14F8IH0UBC9Yn4YsHhLixtnacjUvcKtta") %>% 
+  filter(grepl(sites_to_pull, name)) %>% 
+  filter(grepl(years_to_pull, name)) #%>% 
+#slice(3:n())
+
+
+## Make functions to find and download needed csvs
+download_l2_csvs <- function(i){
+  
+  folder <- l2_folders %>% slice(i)
+  
+  folder_name <- folder$name
+  
+  message(paste("downloading data from", folder_name))
+  
+  folder_id <- folder$id
+  
+  y <- googledrive::drive_ls(folder_id, pattern = ".parquet$") %>% 
+    filter(!(grepl("_OW_", name))) %>% 
+    filter(grepl("soil-vwc-10cm|gw-pressure|gw-density", name))
+  
+  # Download files only if they aren't already present
+  for (j in seq_along(y$id)) {
+    file_path <- paste0(temp_storage, "raw_files/", y$name[[j]])
+    
[…]

### 2025-09-14 2-code/x-old/3b-250618_data_for_kp_synoptic_wl_vwc.R

@@ -1,0 +1,633 @@
+
+
+## This script pulls data from 2021-2023 for 6 (not SWH) synoptic sites, looking
+## for a full year (ish) of data across our sites, and calculating 
+## 1) the % of time that each site-location is under water, and 
+## 2) the average VWC at 5 cm
+##
+## Peter Regier
+## 2025-06-18
+##
+# ########### #
+# ########### #
+
+
+# 1. Setup ---------------------------------------------------------------------
+
+## Load packages with pacman::p_load
+require(pacman)
+p_load(tidyverse,
+       janitor,
+       parsedate,
+       lubridate,
+       cowplot, 
+       purrr,
+       readr, 
+       tictoc,
+       beepr,
+       googledrive, 
+       furrr,
+       PNWColors,
+       googlesheets4)
+
+# Set ggplot theme
+theme_set(theme_bw())
+
+common_tz = "Etc/GMT+5"
+
+
+sites_to_pull = "GWI|GCW|MSM|PTR|CRC|OWC"
+years_to_pull = "2022|2023|2024"
+
+temp_storage = "TEMP/"
+
+
+# 2. Pull L1 data --------------------------------------------------------------
+
+## We are pulling lots of data, let's figure out how to do that programmatically
+l2_folders <- drive_ls("https://drive.google.com/drive/folders/14F8IH0UBC9Yn4YsHhLixtnacjUvcKtta") %>% 
+  filter(grepl(sites_to_pull, name)) %>% 
+  filter(grepl(years_to_pull, name)) #%>% 
+#slice(3:n())
+
+
+## Make functions to find and download needed csvs
+download_l2_csvs <- function(i){
+  
+  folder <- l2_folders %>% slice(i)
+  
+  folder_name <- folder$name
+  
+  message(paste("downloading data from", folder_name))
+  
+  folder_id <- folder$id
+  
+  y <- googledrive::drive_ls(folder_id, pattern = ".parquet$") %>% 
+    filter(!(grepl("_OW_", name))) %>% 
+    filter(grepl("soil-vwc-10cm|gw-pressure|gw-density", name))
+  
+  # Download files only if they aren't already present
+  for (j in seq_along(y$id)) {
+    file_path <- paste0(temp_storage, "raw_files/", y$name[[j]])
+    
+    if (!file.exists(file_path)) {
+      message(paste("Downloading", y$name[[j]], "to", file_path))
+      googledrive::drive_download(y$id[[j]],
[…]

### 2025-09-14 3-reports/manuscript_figures.Rmd
update sensor graphs

@@
 
+print("2022-24")
+
 vwc_data %>% 
@@
 
+print("Jan-Dec 2023")
+
+vwc_data %>% 
+  filter(value > 0) %>% 
+  filter(datetime_est > "2023-01-01") %>% 
+  filter(datetime_est < "2023-12-31") %>% 
+  reorder_site() %>% 
+  reorder_transect() %>% 
+  ggplot(aes(x = datetime_est, y = value, color = transect))+
+  geom_line()+
+  #  facet_wrap(~ site + transect)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey")) +
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
+                     values = pal_transect)+
+  labs(x = "", 
+       y = "Water content (% v/v)",
+       color = "")+
+  theme_kp()+
+  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
+        panel.grid = element_blank())
+
 ```
@@
 ```{r gg_water_level, fig.height=8, fig.width=10}
+
+print("2022-24")
 
@@
   geom_hline(yintercept = 0, alpha = 0.8)+
-  geom_line(aes(y = rollmean), linewidth = 1)+
+  geom_line(aes(y = wl_below_surface_m), linewidth = 1)+
   facet_nested_wrap(~region + site,
@@
   theme(panel.grid = element_blank(),
-        axis.text.x = element_text(size = 14, angle = 45, hjust = 1))   ##
+        axis.text.x = element_text(size = 14, angle = 45, hjust = 1))
+
+print("Jan-Dec 2023")
+
+troll %>% 
+  filter(datetime_est > "2023-01-01") %>% 
+  filter(datetime_est < "2023-12-31") %>% 
+  reorder_transect() %>% 
+  ggplot(aes(x = datetime_est, y = wl_below_surface_m, color = transect))+
+  geom_hline(yintercept = 0, alpha = 0.8)+
+  geom_line(aes(y = wl_below_surface_m), linewidth = 1)+
+  facet_nested_wrap(~region + site,
+                    nest_line = element_line(colour = "grey"))+ 
+  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
+                     values = pal_transect)+
+  labs(x = "",
+       y = "Water level below surface (m)",
+       color = "")+
+  theme_kp()+
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(size = 14, angle = 45, hjust = 1))
 
@@
 ```{r}
+print("Jan-Dec 2023")
+
 water_table %>% reorder_transect() %>% knitr::kable()
@@
 ```{r}
+print("Jan-Dec 2023")
+
 percent_flooded %>% reorder_transect() %>% knitr::kable()

### 2025-09-26 _targets.R
update targets. export physics data

@@ -13,7 +13,7 @@ tar_option_set(
 
 # Load the R scripts with your custom functions:
 source("2-code/0-packages.R")
-#source("2-code/1-initial_processing.R")
+#source("2-code/1-initial_processing.R") # use this only once, to pull the necessary sample weights, etc.
 source("2-code/2-functions_processing.R")
 source("2-code/3a-functions_soil_moisture_sensors.R")
 
@@ -28,7 +28,7 @@ list(
   tar_target(analysis_key_data, "1-data/analysis_key.csv", format = "file"),
   tar_target(analysis_key, read.csv(analysis_key_data)),
   
-  # data files
+  # data files - chemistry
   tar_target(moisture_data, import_gsheet("1nYzExPmpv01IYalqo2L3pmDsBDvhjdvHYMJo31N5_MA")),
   tar_target(moisture_processed, process_moisture(moisture_data)),
   tar_target(loi_data, import_loi(FILEPATH = "1-data/raw/loi")),
@@ -63,30 +63,19 @@ list(
   tar_target(wrc_processed, process_wrc(wrc_data)),
   tar_target(wrc_parameters, import_wrc_parameters(FILEPATH = "1-data/raw/wrc")),
   tar_target(texture_processed, compute_texture(FILEPATH = "https://docs.google.com/spreadsheets/d/13yOYC7vVzVzatXgJUaey2RYdmsJnfnmLmnVWVt4mOXo/", sample_key)),
-  # bulk density
+  tar_target(bd_data_file, "1-data/raw/bulk_density.csv", format = "file"),
+  tar_target(bd_data, read.csv(bd_data_file)),
   tar_target(pd_data, read_sheet("https://docs.google.com/spreadsheets/d/1wHmiOvxa2SQhiGLIiUkNqiDpJQk__XjfrmA1fyuVYEA/")),
   tar_target(pd_processed, process_pd(pd_data)),
-
+  
   # combined data
   tar_target(data_combined_all_horizons, combine_data(moisture_processed, pH_processed, tctnts_data_samples, loi_processed,
-                                         weoc_processed, dic_processed, din_processed, icp_processed,
-                                         ferrozine_processed, mehlich_processed, ions_processed_meq, 
-                                         sample_key)),
+                                                      weoc_processed, dic_processed, din_processed, icp_processed,
+                                                      ferrozine_processed, mehlich_processed, ions_processed_meq, 
[…]

### 2025-09-26 2-code/0-packages.R
update packages

@@ -4,6 +4,7 @@ library(PNWColors)
 library(soilpalettes)
 library(lubridate)
 library(googlesheets4)
+library(multcomp)
 
 theme_kp <- function() {  # this for all the elements common across plots
   theme_bw() %+replace%

### 2025-09-26 3-reports/manuscript_figures.Rmd
minor cleaning

@@
   ylim(35, 49)+
-#  scale_x_continuous(labels = NULL)+
-#  scale_y_continuous(labels = NULL)+
   theme(legend.background = element_rect(fill = alpha("white", 0.0)), 

### 2025-09-26 2-code/2-functions_processing.R

@@ -967,7 +967,7 @@ process_xrd = function(xrd_data, sample_key){
 
 
 ## bulk density
-
+## importing calculated directly
 
 ## particle density
 process_pd = function(pd_data){
@@ -1015,7 +1015,6 @@ combine_data = function(moisture_processed, pH_processed, tctnts_data_samples, l
   data_combined_all_horizons
 
 }
-#data_combined = combine_data()  
 
 subset_surface_horizons = function(data_combined_all_horizons){
   
@@ -1307,7 +1306,3 @@ make_summary_tables <- function(data_combined,
        summary_XRD = summary_XRD)
   
 }
-# targets::tar_load(summary_tables)
-# openxlsx::write.xlsx(summary_tables, "summary_tables.xlsx")
-
-
